---
title: "20221214_MIC_MBC_oxygen"
output: html_document
date: '2022-12-14'
---

```{r setup, echo =FALSE}
.libPaths('C:\\Users\\zuendj\\Documents\\R4Libs')
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/zuendj/Desktop/03_data R", cho = TRUE, warning = FALSE, message = FALSE)

# Install Packages:
# # install.packages("rlang")
# install.packages("tidyverse")
# # install.packages("ggbreak")

# library(plyr)
#  library(rstatix)
#  library(tidyverse)
#  library(ggplot2)
# 
# # # install.packages("devtools")
# # # devtools::install_github("slowkow/ggrepel")
# library(ggrepel)
# library(ggbreak)
# # # install.packages("openxlsx")
# library(openxlsx)
library(ggh4x)
```

```{r}

o2_vector <- c(
  "Faecalibacterium prausnitzii",
  "Fusicatenibacter saccharivorans",
  "Lachnospira eligens",
  "Agathobacter rectalis",
  "Anaerotignum lactatifermentans",
  "Bacteroides fragilis",
  "Bifidobacterium adolescentis",
  "Blautia hydrogenotrophica",
  "Dorea longicatena",
  "Eubacterium limosum",
  "Faecalibacterium duncaniae",
  "Roseburia intestinalis",
  "Segatella copri",
  "Anaerostipes hadrus",
  "Bacteroides uniformis",
  "Bacteroides xylanisolvens",
  "Blautia hansenii",
  "Butyricicoccus pullicaecorum",
  "Collinsella aerofaciens",
  "Mediterraneibacter gnavus",
  "Phocaeicola dorei",
  "Phocaeicola vulgatus",
  "Bacteroides intestinalis",
  "Blautia obeum",
  "Blautia producta",
  "Enterocloster aldenensis",
  "Roseburia faecis",
  "Anaerostipes caccae",
  "Clostridium sporogenes",
  "Bacteroides caccae",
  "Bacteroides cellulosilyticus",
  "Bacteroides faecis",
  "Bacteroides ovatus",
  "Bacteroides thetaiotaomicron",
  "Parabacteroides distasonis",
  "Parabacteroides merdae",
  "Phascolarctobacterium faecium",
  "Bifidobacterium longum",
  "Enterococcus faecalis",
  "Escherichia coli",
  "Lactiplantibacillus plantarum"
)

```



# 1. Load the Data - aerobic
```{r}
#########################################
#Load the data
########################################

setwd("P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/01_OD_pure/01_O2/01_import")
  # bacteria <- list.files(path = "P:/Shared_documents/Janina Z?nd/08_phd/09_student thesis/Marina/03_experimentation/03_MIC_MBC/01_OD-measurements/single_bacteria_for_plot")

list.files(path = "P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/01_OD_pure/01_O2/01_import")->bacteria


options(scipen=999)
#########################################
#load the data
########################################

#define empty matrices, no values added so far 
MIC_O2_values<- matrix(0, length(bacteria)*6, 12)
MIC_O2_average<- matrix(0, length(bacteria), 12)

MBC_O2_values<- matrix(0, length(bacteria)*6, 12)
MBC_O2_average<- matrix(0, length(bacteria), 12)



#MIC_O2, fill values into MIC_O2, here OD corrected values, but not corrected with agar 
row = 0
for (l in 1:length(bacteria)) {
  #generate MIC_O2wide format
  MIC_O2wide = read.csv(bacteria[l], sep=";", header=F, check.names =F)
  #look for A indicating the area in the excel file with the data

  for (r in 35:42) {
    if (MIC_O2wide[r,1]=="C") {
      zeile = r
    }
  }
#########################################
  for(i in 1:12) {
    #x indicated the colum->'A' is in column 1 and the data next to it from column 2 on
    x = 1+i
    zeile_end = zeile+6
    d = as.numeric(unlist(data.frame(MIC_O2wide[zeile:zeile_end, x])))
    for(j in 1:6){
      y = d[j]
      MIC_O2_values[row+j, i] = y
      }
  }
  
  #########################################
  #generate averages 
  for(i in 1:12) {
    r1= row+1
    r2= row+6
    MIC_O2_average[l, i] = as.numeric(mean(MIC_O2_values[r1:r2, i],  na.rm = TRUE))
  }
  
  row=row+6
  
}  

#MBC_O2, fill values into MBC_O2, here OD corrected values, but not corrected with agar 
row = 0
for (l in 1:length(bacteria)) {
  #generate MIC_O2_O2wide format
  MIC_O2wide = read.csv(bacteria[l], sep=";", header=F, check.names =F)
  #look for A indicating the area in the excel file with the data
  for (r in 49:56) {
    if (MIC_O2wide[r,1]=="C") {
      zeile = r
    }
  }
  #########################################
  for(i in 1:12) {
    #x indicated the colum->'A' is in column 1 and the data next to it from column 2 on
    x = 1+i
    zeile_end = zeile+6
    d = as.numeric(unlist(data.frame(MIC_O2wide[zeile:zeile_end, x])))
    for(j in 1:6){
      y = d[j]
        MBC_O2_values[row+j, i] = y
    }
  }
  
  #########################################
  #generate averages 
  for(i in 1:12) {
    r1= row+1
    r2= row+6
    MBC_O2_average[l, i] = as.numeric(mean(MBC_O2_values[r1:r2, i],  na.rm = TRUE))
  }
  
  row=row+6
  
}  


#########################################
```
#Load the anaerobic data

```{r}
#########################################
#Load the data
########################################

setwd("P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/01_OD_pure/01_O2/01_import")

options(scipen=999)
#########################################
#load the data
########################################


####  anaerobic
MIC_O2_values_ana<- matrix(0, length(bacteria)*6, 12)
MIC_O2_average_ana<- matrix(0, length(bacteria), 12)

MBC_O2_values_ana<- matrix(0, length(bacteria)*6, 12)
MBC_O2_average_ana<- matrix(0, length(bacteria), 12)

#MIC_O2, fill values into MIC_O2, here OD corrected values, but not corrected with agar 
row = 0
for (l in 1:length(bacteria)) {
  #generate MIC_O2wide format
  MIC_O2wide = read.csv(bacteria[l], sep=";", header=F, check.names =F)
  #look for A indicating the area in the excel file with the data

  for (r in 35:42) {
    if (MIC_O2wide[r,19]=="C") {
      zeile = r
    }
  }
#########################################
  for(i in 1:12) {
    #x indicated the colum->'A' is in column 1 and the data next to it from column 2 on
    x = 19+i
    zeile_end = zeile+6
    d = as.numeric(unlist(data.frame(MIC_O2wide[zeile:zeile_end, x])))
    for(j in 1:6){
      y = d[j]
      MIC_O2_values_ana[row+j, i] = y
      }
  }
  
  #########################################
  #generate averages 
  for(i in 1:12) {
    r1= row+1
    r2= row+6
    MIC_O2_average_ana[l, i] = as.numeric(mean(MIC_O2_values_ana[r1:r2, i],  na.rm = TRUE))
  }
  
  row=row+6
  
}  

#MBC_O2, fill values into MBC_O2, here OD corrected values, but not corrected with agar 
row = 0
for (l in 1:length(bacteria)) {
  #generate MIC_O2_O2wide format
  MIC_O2wide = read.csv(bacteria[l], sep=";", header=F, check.names =F)
  #look for A indicating the area in the excel file with the data
  for (r in 49:56) {
    if (MIC_O2wide[r,19]=="C") {
      zeile = r
    }
  }
  #########################################
  for(i in 1:12) {
    #x indicated the colum->'A' is in column 1 and the data next to it from column 2 on
    x = 19+i
    zeile_end = zeile+6
    d = as.numeric(unlist(data.frame(MIC_O2wide[zeile:zeile_end, x])))
    for(j in 1:6){
      y = d[j]
        MBC_O2_values_ana[row+j, i] = y
    }
  }
  
  #########################################
  #generate averages 
  for(i in 1:12) {
    r1= row+1
    r2= row+6
    MBC_O2_average_ana[l, i] = as.numeric(mean(MBC_O2_values_ana[r1:r2, i],  na.rm = TRUE))
  }
  
  row=row+6
  
}  


#########################################
```


# Row names
```{r}
#########################################
#define row names for the blank corrected 
bacteria <- strsplit(bacteria, ".csv")
bacteria <- as.character(bacteria)



row_names<-matrix(0,length(bacteria)*6,1)
position = 0
for (i in 1:length(bacteria)) {
  for (j in 1:6) {
    row_names[position+j,1] = bacteria[i]
  }
  position = position+6
}

MIC_O2_replicates <-data.frame(row_names, MIC_O2_values)
colnames(MIC_O2_replicates)<-list("bacteria", 1, 2, 3, 4, 5, 6, 7, 8, 9, 10 , 11 , 12 )

MBC_O2_replicates <-data.frame(row_names, MBC_O2_values)
colnames(MBC_O2_replicates)<-list("bacteria", 1, 2, 3, 4, 5, 6, 7, 8, 9, 10 , 11 , 12 )

MIC_O2_replicates_ana <-data.frame(row_names, MIC_O2_values_ana)
colnames(MIC_O2_replicates_ana)<-list("bacteria", 1, 2, 3, 4, 5, 6, 7, 8, 9, 10 , 11 , 12 )

MBC_O2_replicates_ana <-data.frame(row_names, MBC_O2_values_ana)
colnames(MBC_O2_replicates_ana)<-list("bacteria", 1, 2, 3, 4, 5, 6, 7, 8, 9, 10 , 11 , 12 )
```

# 2. Export blanck corrected OD values for outlier correction
```{r}
MBC_O2_replicates$replicate = rep(c(1,1,2,2,3,3), times=length(unique(bacteria)))
MBC_O2_replicates$technical_rep = rep(c("a","b"), times = 3* length(unique(bacteria)))
MIC_O2_replicates$replicate = rep(c(1,1,2,2,3,3), times=length(unique(bacteria)))
MIC_O2_replicates$technical_rep = rep(c("a","b"), times = 3* length(unique(bacteria)))
```

##Outlier correction
```{r}
calculate_iqr <- function(x) {
  q3 <- quantile(x, 0.75 , na.rm =TRUE)
  q1 <- quantile(x, 0.25, na.rm =TRUE)
  iqr_value <- q3 - q1
  return(iqr_value)}
q3_calc <- function(x) {
  q3 <- quantile(x, 0.75, na.rm =TRUE)
  return(q3)}

q1_calc <- function(x) {
  q1 <- quantile(x, 0.25, na.rm =TRUE)
  return(q1)}

z=1.5

#####MBC

MBC_O2_replicates%>%
  gather(condition, value, `1`: `12`)->MBC_O2_replicates_long



left_join(MBC_O2_replicates_long, setNames(aggregate(MBC_O2_replicates_long$value, by=list(MBC_O2_replicates_long$bacteria, MBC_O2_replicates_long$condition), FUN=calculate_iqr), c("bacteria", "condition","IQR")))%>%
  left_join(., setNames(aggregate(MBC_O2_replicates_long$value, by=list(MBC_O2_replicates_long$bacteria, MBC_O2_replicates_long$condition), FUN=q3_calc), c("bacteria", "condition","Q3")))%>%
  left_join(., setNames(aggregate(MBC_O2_replicates_long$value, by=list(MBC_O2_replicates_long$bacteria, MBC_O2_replicates_long$condition), FUN=q1_calc), c("bacteria", "condition","Q1")))%>%
mutate(value = ifelse(value - (Q1 - (z*IQR)) < 0, NA, value))%>% #lowerbound
  mutate(value = ifelse((Q3 + (z*IQR))- value < 0, NA, value))-> outlier_corr

outlier_corr[, c("bacteria", "condition", "value", "replicate","technical_rep")]%>%
  pivot_wider(values_from = value, names_from = condition)->MBC_O2_replicates_outlier_corr

#######MIC

MIC_O2_replicates%>%
  gather(condition, value, `1`: `12`)->MIC_O2_replicates_long

left_join(MIC_O2_replicates_long, setNames(aggregate(MIC_O2_replicates_long$value, by=list(MIC_O2_replicates_long$bacteria, MIC_O2_replicates_long$condition), FUN=calculate_iqr), c("bacteria", "condition","IQR")))%>%
  left_join(., setNames(aggregate(MIC_O2_replicates_long$value, by=list(MIC_O2_replicates_long$bacteria, MIC_O2_replicates_long$condition), FUN=q3_calc), c("bacteria", "condition","Q3")))%>%
  left_join(., setNames(aggregate(MIC_O2_replicates_long$value, by=list(MIC_O2_replicates_long$bacteria, MIC_O2_replicates_long$condition), FUN=q1_calc), c("bacteria", "condition","Q1")))%>%
mutate(value = ifelse(value - (Q1 - (z*IQR)) < 0, NA, value))%>% #lowerbound
  mutate(value = ifelse((Q3 + (z*IQR))- value < 0, NA, value)) -> outlier_corr

outlier_corr[, c("bacteria", "condition", "value", "replicate","technical_rep")]%>%
  pivot_wider(values_from = value, names_from = condition)->MIC_O2_replicates_outlier_corr
  

#####MBC ana

cbind(MBC_O2_replicates_ana, MBC_O2_replicates[,c( "replicate", "technical_rep")])%>%
  gather(condition, value, `1`: `12`)->MBC_O2_replicates_long



left_join(MBC_O2_replicates_long, setNames(aggregate(MBC_O2_replicates_long$value, by=list(MBC_O2_replicates_long$bacteria, MBC_O2_replicates_long$condition), FUN=calculate_iqr), c("bacteria", "condition","IQR")))%>%
  left_join(., setNames(aggregate(MBC_O2_replicates_long$value, by=list(MBC_O2_replicates_long$bacteria, MBC_O2_replicates_long$condition), FUN=q3_calc), c("bacteria", "condition","Q3")))%>%
  left_join(., setNames(aggregate(MBC_O2_replicates_long$value, by=list(MBC_O2_replicates_long$bacteria, MBC_O2_replicates_long$condition), FUN=q1_calc), c("bacteria", "condition","Q1")))%>%
mutate(value = ifelse(value - (Q1 - (z*IQR)) < 0, NA, value))%>% #lowerbound
  mutate(value = ifelse((Q3 + (z*IQR))- value < 0, NA, value))-> outlier_corr

outlier_corr[, c("bacteria", "condition", "value", "replicate","technical_rep")]%>%
  pivot_wider(values_from = value, names_from = condition)->MBC_O2_replicates_outlier_corr_ana

#######MIC ana
cbind(MIC_O2_replicates_ana, MIC_O2_replicates[,c( "replicate", "technical_rep")])%>%
  gather(condition, value, `1`: `12`)->MIC_O2_replicates_long

left_join(MIC_O2_replicates_long, setNames(aggregate(MIC_O2_replicates_long$value, by=list(MIC_O2_replicates_long$bacteria, MIC_O2_replicates_long$condition), FUN=calculate_iqr), c("bacteria", "condition","IQR")))%>%
  left_join(., setNames(aggregate(MIC_O2_replicates_long$value, by=list(MIC_O2_replicates_long$bacteria, MIC_O2_replicates_long$condition), FUN=q3_calc), c("bacteria", "condition","Q3")))%>%
  left_join(., setNames(aggregate(MIC_O2_replicates_long$value, by=list(MIC_O2_replicates_long$bacteria, MIC_O2_replicates_long$condition), FUN=q1_calc), c("bacteria", "condition","Q1")))%>%
mutate(value = ifelse(value - (Q1 - (z*IQR)) < 0, NA, value))%>% #lowerbound
  mutate(value = ifelse((Q3 + (z*IQR))- value < 0, NA, value)) -> outlier_corr

outlier_corr[, c("bacteria", "condition", "value", "replicate","technical_rep")]%>%
  pivot_wider(values_from = value, names_from = condition)->MIC_O2_replicates_outlier_corr_ana

```
```{r}
rbind(MIC_O2_replicates_outlier_corr%>%
        mutate(assay="MIC_aerobic"),MBC_O2_replicates_outlier_corr%>%
        mutate(assay="MBC_aerobic") )%>%
  rbind(., MIC_O2_replicates_outlier_corr_ana%>%
        mutate(assay="MIC_anaerobic"))%>%
   rbind(., MBC_O2_replicates_outlier_corr_ana%>%
        mutate(assay="MBC_anaerobic"))%>%
  write.xlsx(., "P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/01_OD_pure/01_O2/O2_growthdata_blank_outlier_corr.xlsx")
```

# 3. Average of triplicates & Find max values 
```{r}
#get long format
average_long_O2<-gather(MIC_O2_replicates_outlier_corr, condition, OD, "1":"12", factor_key = T )
#take averages of the technical replicate
average_long_tec_O2<-set_names(aggregate(average_long_O2$OD, list(average_long_O2$bacteria, average_long_O2$replicate, average_long_O2$condition), FUN=mean, na.rm=T), c("bacteria",  "replicate","condition", "OD"))
#get average per condition -> to get max
average_O2<-set_names(aggregate(average_long_tec_O2$OD, list(average_long_tec_O2$bacteria, average_long_tec_O2$condition), FUN=mean, na.rm=TRUE), c("bacteria", "condition" ,"average"))
# find max
max_mic_O2<-set_names(aggregate(average_O2$average, by = list(average_O2$bacteria), max, na.rm=TRUE), c("bacteria", "max"))

MIC_replicates_wide_O2<-pivot_wider(average_long_tec_O2, names_from = condition , values_from = OD)

average_long_O2<-gather(MBC_O2_replicates_outlier_corr, condition, OD, "1":"12", factor_key = T )
average_long_tec_O2<-set_names(aggregate(average_long_O2$OD, list(average_long_O2$bacteria, average_long_O2$replicate, average_long_O2$condition), FUN=mean, na.rm=TRUE), c("bacteria",  "replicate","condition", "OD"))
average_O2<-aggregate(average_long_tec_O2$OD, list(average_long_tec_O2$bacteria, average_long_tec_O2$condition), FUN=mean, na.rm=TRUE)
names(average_O2)<-c("bacteria", "condition" ,"average")
max_mbc_O2<-set_names(aggregate(average_O2$average, by = list(average_O2$bacteria), max, na.rm=TRUE), c("bacteria", "max"))

MBC_replicates_wide_O2<-pivot_wider(average_long_tec_O2, names_from = condition , values_from = OD)

MIC_mod_av_O2<-left_join(MIC_replicates_wide_O2, max_mic_O2)
MBC_mod_av_O2<-left_join(MBC_replicates_wide_O2, max_mbc_O2)
```

#compare the effect of agar
```{r}


left_join(MIC_O2_replicates_outlier_corr_ana,
          read.xlsx("P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/00_experimentation/01_pure_culture/20240207_Strain_panel.xlsx")[, c("bacteria", "Phylum", "Family", "Genus")])%>%
  mutate(control = `12`)%>%
  gather(condition, OD, "1":"12")%>%
  mutate(condition = factor(condition, levels = c("12","11", "10", "9", "8", "7", "6", "5", "4", "3", "2", "1")))%>%
  mutate(OD_rel = OD/control)->sub

setNames(aggregate(sub$OD_rel, by =list(sub$bacteria, sub$condition, sub$Genus), FUN="mean", na.rm=T), c("bacteria", "condition","Phylum", "mean"))%>%
  ggplot(aes(x=condition, y=mean))+
  geom_violin(aes())+
  geom_jitter(aes(colour=Phylum), alpha=0.6, width = 0.4)+
   stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               size=0.4,
                colour = "black")+
  theme(text = element_text(size  = 15),
        axis.text.x = element_text(angle=90, hjust=1, vjust=0.5),
        legend.text = element_text(vjust=1, hjust=0,face="italic"),
        strip.background = element_rect(fill="white"),
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), 
        panel.spacing = unit(0.31, "cm"),
      
        legend.text.align = 0,
        legend.key.height  = unit(0.5, 'cm'))+
  # scale_colour_manual(values=c("#a6cee3","indianred","cyan4","#b2df8a"))+
  scale_colour_manual(values=c("#1f78b4","#a6cee3", "#80b1d3","#8dd3c7","#33a02c", "pink3","#e31a1c", "#ff7f00",   "#fb9a99",  "#6a3d9a", "#cab2d6","#fdbf6f", "darkgoldenrod2", "#b15928",  "#fb8072", "#5c342c", "#b36966", "#A8AF51" , "#484939", "#CE6C4E", "#E6BAAC",  "#B1A7CD", "darksalmon", "lightgreen"), name="Genus")+
  scale_x_discrete(labels = c("1" = "0.15", 
                              "2"= "0.13", 
                              "3"="0.11", 
                              "4"="0.09",
                              "5" ="0.08", 
                             "6"= "0.07", 
                             "7" = "0.06", 
                             "8" = "0.05", 
                             "9" = "0.04", 
                             "10" = "0.035",  
                              "11" = "0.03",
                             "12" = "0"))+
   force_panelsizes(rows = unit(5, "cm"))+
  ylab("Relative OD600")+xlab("Agar [%]")


   
ggsave(filename = paste("agar_pure..jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final",
       width = 17,
       height = 15,
       units = c("cm"))

  
```


##Comparison of aerobic values and anaerobic values

```{r}
library(ggpubr)
#join the two datframes (anaerobic an aerobic values)

left_join(MIC_O2_replicates_outlier_corr %>%
  gather(condition, aerobe, `1`:`12`),
  MIC_O2_replicates_outlier_corr_ana %>%
  gather(condition, anaerobe, `1`:`12`))%>%
  gather(incubation, OD, "aerobe":"anaerobe")%>%
  mutate(condition = factor(condition, levels=c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")))-> sub


sub%>%
  group_by(condition, bacteria)%>%
  wilcox_test(OD~incubation)%>%
  add_significance()%>%
  add_xy_position(fun = "max", x="condition")->stats

left_join(setNames(aggregate(sub$OD, by=list(sub$condition, sub$incubation, sub$bacteria), FUN=sd, na.rm=T), c("condition", "incubation", "bacteria", "sd")),
setNames(aggregate(sub$OD, by=list(sub$condition, sub$incubation, sub$bacteria), FUN=mean, na.rm=T), c("condition", "incubation", "bacteria", "mean"))) %>%
left_join(., sub)%>%
  ggplot(aes(x= condition, y=mean))+
   geom_col(aes(fill=incubation, position =incubation), position = position_dodge(0.8), width=0.7)+
  geom_errorbar(aes(ymin=mean - sd, ymax=mean +sd, group=incubation), width=.2,
                 position=position_dodge(.8))+
  stat_pvalue_manual(stats, label="p.signif", y.position = "y.position", x.position="condition")+
  facet_grid(rows=vars(bacteria), scales="free")+
   theme(text = element_text(size=8), 
        axis.title.x = element_text(vjust=-0.8),
        axis.text.x = element_text(angle = 90, vjust= 0.5, hjust=1),
        axis.title.y= element_blank(),
        strip.text.y = element_text(angle=0,  hjust = 0, face="italic"),
        strip.background = element_blank(),
        axis.ticks.y = element_blank(),
        plot.title = element_text(hjust=0.25),
        panel.background = element_rect(fill="transparent"),
        panel.spacing = unit(0, "lines"))
  
  
ggsave(filename = "MIC_MBC_o2_growth_only.jpeg", 
       path = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/output_reanalysis", 
       width = 20,
       height = 40,
       units = c("cm"))
```


#heatplot with values relative to anaerobic controls
```{r}
#combine all data frames
left_join(MIC_O2_replicates_outlier_corr %>%
  gather(condition, aerobe, `1`:`12`),
  MIC_O2_replicates_outlier_corr_ana %>%
  gather(condition, anaerobe, `1`:`12`)%>%
    mutate(assay = "MIC")) ->MIC

  left_join(MBC_O2_replicates_outlier_corr %>%
  gather(condition, aerobe, `1`:`12`),
             MBC_O2_replicates_outlier_corr_ana %>%
  gather(condition, anaerobe, `1`:`12`)%>%
    mutate(assay = "MBC")) -> MBC
  
rbind(MIC, MBC)%>%
  gather(incubation, OD, "aerobe":"anaerobe")  -> joint
  
  
setNames(aggregate(joint$OD, by =list( joint$condition, joint$bacteria, joint$assay, joint$incubation), FUN="mean", na.rm=T), c( "condition", "bacteria", "assay", "incubation","OD_mean"))%>%
  pivot_wider(names_from = incubation, values_from = OD_mean)%>%
  mutate(ratio = aerobe / anaerobe)->OD_relative_to_anaerobe
      
```
# 4. get values of tolerated concentration
### two vectors are generated: one using a growth tresh hold of 10% of the max growth and one for 50 % of the may grwoth &&& an minimal growth of an od of 0.0344 is required (derived from the stdv. of absorbance of the blank)
```{r}
OD_relative_to_anaerobe[,c("condition", "bacteria", "assay", "ratio")]%>%
  # mutate(condition = as.factor(condition, levels=c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")))%>%
  pivot_wider(names_from = condition, values_from = c(ratio))->wide


wide [,c("assay", "bacteria","1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12" )]->wide

wide%>%
  subset(assay %in% "MIC")-> MIC_mod_av_O2

MIC_vector_O2 <- vector()
for (i in 1:nrow(MIC_mod_av_O2)) { #vertical 
  candidates <- vector()
     for(j in 3:14) {#horizontal 
       if (MIC_mod_av_O2[i,j]>0.2) {#-> MIC_value_absolut)

         candidates <- append(candidates, colnames(MIC_mod_av_O2)[j])
         candidates <-as.numeric(candidates)
         # print( max(candidates))
       }
     }
  MIC_vector_O2 <- append(MIC_vector_O2, max(candidates))
}

wide %>%
  subset(assay %in% "MBC")-> MBC_mod_av_O2

MBC_vector_O2 <- vector()
for (i in 1:nrow(MBC_mod_av_O2)) { #vertical 
  candidates <- vector()
     for(j in 3:14) {#horizontal 
       if ( MBC_mod_av_O2[i,j] > 0.2) {#-> MIC_value_absolut)
         
         

         candidates <- append(candidates, colnames(MBC_mod_av_O2)[j])
         candidates <-as.numeric(candidates)
        
       }
     }
  # print( max(candidates))
  MBC_vector_O2 <- append(MBC_vector_O2, max(candidates))
}


```


```{r}
#assign vectors with MIC/MBCs to df
df_final_replicates_O2 <- data.frame(bacteria = MBC_mod_av_O2$bacteria,
                                  MIC = as.numeric(MIC_vector_O2), 
                                  MBC = as.numeric(MBC_vector_O2))%>%
   mutate(bacteria = ifelse(bacteria == "Ruminococcus gnavus", "Mediterraneibacter gnavus", bacteria))%>%
   mutate(bacteria = ifelse(bacteria == "Prevotella copri", "Segatella copri", bacteria))

write.xlsx(df_final_replicates_O2, "P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/01_OD_pure/01_O2/O2_tolerances.xlsx" )


df_final_replicates_O2%>%
 mutate(MBC = ifelse(bacteria == "Escherichia coli" |
                      bacteria == "Lactiplantibacillus plantarum" |
                        bacteria == "Enterococcus faecalis" |
                        bacteria == "Bifidobacterium longum", 13, MBC))-> order_o2


df_final_replicates_O2%>%
  subset(MBC > MIC)
```


##plot - relative to anaerobe values
```{r}
# install.packages("ggnewscale")
 library(ggnewscale)

write.xlsx(OD_relative_to_anaerobe, "P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/01_OD_pure/01_O2/O2_relative_OD_values.xlsx")

read.xlsx("P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/02_hplc_pure/01_O2/sampling_O2.xlsx")[, c("assay", "well.column")]%>%
  set_names(., c("bacteria", "agar_condition"))
  

left_join(OD_relative_to_anaerobe%>%
           
  mutate(bacteria = ifelse(bacteria == "Prevotella copri", "Segatella copri", bacteria))%>%
  mutate(bacteria = ifelse(bacteria == "Ruminococcus gnavus", "Mediterraneibacter gnavus", bacteria))%>%
  mutate(bacteria = ifelse(bacteria == "Bacteroides dorei", "Phocaeicola dorei", bacteria))%>%
  mutate(bacteria = factor(bacteria)) , df_final_replicates_O2%>% 
    mutate(bacteria = ifelse(bacteria == "Bacteroides dorei", "Phocaeicola dorei", bacteria)))%>% 
  mutate(condition = as.numeric(paste(condition)))%>% 
  mutate(MBC = ifelse(bacteria == "Escherichia coli" |
                      bacteria == "Lactiplantibacillus plantarum" |
                        bacteria == "Enterococcus faecalis" |
                        bacteria == "Bifidobacterium longum", 13, MBC))%>%
  mutate(bacteria = reorder(bacteria, MBC))%>%
   mutate(ratio = ratio*100)%>%
  mutate(ratio = ifelse(ratio <0, 0, ratio))%>%
  mutate(ratio = ifelse(ratio >120, 120, ratio))%>%
  mutate(assay= ifelse(assay == "MIC", "a_MIC", assay))->o2_long


o2_long%>%
  mutate(MIC= MIC+1)%>%
   mutate(MBC= MBC+1)%>%
  ggplot(aes(x = condition, y = assay)) + 
  labs(x=c("Agar [%]"))+
  theme(text = element_text(size=10), 
        axis.title.x = element_text(vjust=-0.8),
        axis.text.x = element_text(angle = 90, vjust= 0.5, hjust=1),
        axis.title.y= element_blank(),
        strip.text.y = element_text(angle=0,  hjust = 0, face="italic"),
        strip.background = element_blank(),
        axis.ticks.y = element_blank(),
        panel.border=element_rect(fill=NA), 
          panel.background = element_rect(fill="white"),
          legend.text.align = 0, 
          panel.spacing = unit(0, "lines"))+
  scale_x_continuous(labels=c("0.15","0.13", "0.11", "0.09" , "0.08", "0.07", "0.06", "0.05", "0.04", "0.035",  "0.03","0"),
                     breaks=c(1, 2, 3, 4, 5, 6, 7,  8 ,9 ,10, 11, 12))+
   scale_y_discrete(labels = c('a_MIC' = 'Stress', 'MBC' = 'Post-stress'), limits=rev)+
  labs(fill = expression(OD~`[%]`), colour="")+
 
  
   geom_raster(aes(fill = ratio)) +
  scale_fill_steps(
     breaks = c( 20, 40,  60,80 , 100, 120),
    name = expression(OD~`[%]`),

    guide = guide_colorbar(title.position = "top", title.vjust = 0.5), low="#193854", high ="deepskyblue" ) +
  # geom_point(data = subset(o2_long, agar_condition == condition & assay == "a_MIC"), aes(x = condition, y = assay), shape=16, colour="seagreen3", alpha=0.7, size=2)+
  geom_point(data = subset(o2_long, condition == MIC & assay == "a_MIC"), aes(x = condition, y = assay), shape=0, colour="seagreen2", size=2.1)+
  geom_point(data = subset(o2_long, condition == MBC & assay == "MBC"), aes(x = condition, y = assay), shape=0, colour="seagreen2", size=2.1)+
  # geom_point(data=subset(sub, ratio > 20), colour="red", size=0.5)+
  facet_grid(rows =vars( bacteria), scales = "free", space = "free")->growth_o2
  
growth_o2

write.csv(sub, file="C:/Users/zuendj/Desktop/03_data R/oxidative stress/gowth_data_o2.csv")


# ggsave(filename = "MIC_MBC_o2.png", 
#        path = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/output_reanalysis", 
#        width =25,
#        height = 39.5,
#        units = c("cm"), 
#        bg="transparent")



ggsave(filename = "MIC_MBC_o2_rel_to_ana..jpeg", 
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final", 
       width = 12,
       height = 22.8,
       units = c("cm"))

```




```{r}
setdiff(
unique(o2_long$bacteria), unique(read.xlsx("P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/00_experimentation/01_pure_culture/20240207_Strain_panel_metabol.xlsx")$bacteria )
)



df_final_replicates_O2%>%
  mutate(mean = MBC + (MIC/5))%>%
  gather(assay, tolerance, "MIC":"MBC")%>%
  mutate(bacteria = ifelse(bacteria == "Bacteroides dorei", "Phocaeicola dorei", bacteria))%>%
  left_join(., 
            read.xlsx("P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/00_experimentation/01_pure_culture/20240207_Strain_panel_metabol.xlsx")[, c("bacteria", "Phylum", "Genus")])%>%
   mutate(bacteria = ifelse(bacteria == "Prevotella copri", "Segatella copri", bacteria))%>%
  mutate(bacteria = ifelse(bacteria == "Ruminococcus gnavus", "Mediterraneibacter gnavus", bacteria))%>%
   mutate(Genus = ifelse(bacteria ==  "Segatella copri", "Segatella", Genus))%>%
  mutate(Genus = ifelse(bacteria ==  "Mediterraneibacter gnavus", "Mediterraneibacter",  Genus))->phylogeny

rbind(phylogeny%>%
        mutate(type= "toler"), 
      phylogeny %>%
        mutate(assay = "a_Genus",
               type="a_Phylo"))%>%
  rbind(., phylogeny %>%
        mutate(assay = "b_Phylum",
               type="a_Phylo"))->sub

sub%>%
  mutate(assay = factor (assay, levels= c( "MIC", "MBC",  "a_Genus","b_Phylum")))%>%
  mutate(bacteria = fct_reorder(bacteria, mean))->sub.

sub. %>%
  # mutate(tolerance = tolerance +1)%>%
  # mutate(tolerance = ifelse( tolerance == 8, 7, tolerance))%>%
  # mutate(tolerance = ifelse( tolerance == 11, 8, tolerance))%>%
  # mutate(tolerance = ifelse( tolerance == 12, 9, tolerance))%>%
  ggplot(aes(x=bacteria, y=assay))+ 
  
  new_scale_fill() +
   geom_tile(data = subset(sub., assay %in% c("a_Genus")), aes(fill = Genus ), colour="black", aplpha=0.75) +
   scale_fill_manual(values=c("#1f78b4","#a6cee3", "#80b1d3","#8dd3c7","#33a02c", "pink3","#e31a1c", "#ff7f00",   "#fb9a99",  "#6a3d9a", "#cab2d6","#fdbf6f", "darkgoldenrod2", "#b15928",  "#fb8072", "#5c342c", "#b36966", "#A8AF51" , "#484939", "#CE6C4E", "#E6BAAC",  "#B1A7CD", "darksalmon", "lightgreen"))+
  guides(fill = guide_legend(nrow = 6))+
  # new_scale_fill() +
  #  geom_tile(data = subset(sub., assay %in% c("a_Genus")), aes(fill = Genus ), colour="black") +
  #  scale_fill_manual(values=c("#1f78b4","#a6cee3", "#80b1d3","#8dd3c7","#33a02c", "pink3","#e31a1c", "#ff7f00",   "#fb9a99",  "#6a3d9a", "#cab2d6","#fdbf6f", "darkgoldenrod2", "#b15928",  "#fb8072", "#5c342c", "#b36966", "#A8AF51" , "#484939", "#CE6C4E", "#E6BAAC",  "#B1A7CD"))+
  
  
  new_scale_fill() +
   geom_tile(data = subset(sub., assay %in% c("MIC", "MBC")), aes(fill = tolerance ), colour="black") +
   # scale_fill_gradientn(colours= c( "indianred","#b36966", "#5e5d5d","cyan4")  , na.value= "white",   )+
    # scale_fill_steps(n.breaks = length(unique(sub.$tolerance)) + 4)+
  
  facet_grid(rows=vars(type), scales="free", space="free")+
   theme(text= element_text(size=15),
        axis.title = element_blank(),
        axis.ticks.y = element_blank(),
        panel.border=element_rect(fill=NA), 
        panel.background = element_rect(fill="white"),
        rect = element_rect(fill = "transparent"),
        # legend.text.align = 0, 
        panel.spacing = unit(0, "lines"),
        axis.text.x = element_text(angle=60, hjust=1, face="italic"),
        strip.background = element_blank(), 
        strip.text = element_blank(),
         # legend.key.height  = unit(0.2, 'cm'),legend.key.width  = unit(0.4, 'cm'),
       legend.position = "top")+
scale_y_discrete(labels = c('MIC' = 'Stress', 'MBC' = 'Post-stress', "a_Genus"="Genus", "b_Phylum"="Phylum"))+
  force_panelsizes(cols=unit(23, "cm"))



ggsave(filename = "O2_species_rolerances.png",   
         path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final",
       device = "png", bg= "transparent", 
          width = 26,
       # width = 50,
         height = 13,
         units = c("cm"))


```
#
```{r}
setdiff(unique(O2$bacteria) ,unique(H2O2$bacteria))


```


### compare O2 and H2O2
```{r}
library(ggnewscale)
 # library(pheatmap)
H2O2<- read.xlsx("P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/01_OD_pure/02_H2O2/H2O2_tolerances.xlsx")%>%
  gather(assay, tolerance, "MIC":"MBC")%>%
  mutate(bacteria = ifelse(bacteria == "Ruminococcus gnavus", "Mediterraneibacter gnavus", bacteria))%>%
  
   mutate(bacteria = ifelse(bacteria == "Prevotella copri", "Segatella copri", bacteria))%>%
  mutate(stress = "H2O2")%>%
  mutate(tolerance = (tolerance - 7)/ (11 - 7))



O2<- read.xlsx("P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/01_OD_pure/01_O2/O2_tolerances.xlsx")%>%
  gather(assay, tolerance, "MIC":"MBC")%>%
  mutate(bacteria = ifelse(bacteria == "Bacteroides dorei", "Phocaeicola dorei", bacteria))%>%
  mutate(stress = "O2")%>%
   mutate(bacteria = ifelse(bacteria == "Ruminococcus gnavus", "Mediterraneibacter gnavus", bacteria))%>%
   mutate(bacteria = ifelse(bacteria == "Prevotella copri", "Segatella copri", bacteria))%>%
    mutate(tolerance = ifelse(tolerance == 8, 7, tolerance)) %>%
  mutate(tolerance = ifelse(tolerance == 11, 8, tolerance)) %>%
  mutate(tolerance = ifelse(tolerance == 12, 9, tolerance)) %>%
  mutate(tolerance = (tolerance - 2)/ (7))
  




names<-list("H2O2"= expression(H[2]*O[2]),"O2"= expression(O[2]))

order<- c("Faecalibacterium duncaniae",
          "Faecalibacterium prausnitzii",
          "Fusicatenibacter saccharivorans",
          "Lachnospira eligens",
          "Agathobacter rectalis" , 
          "Anaerotignum lactatifermentans",
          "Blautia hydrogenotrophica",
          "Dorea longicatena"  ,
          "Segatella copri" ,
          "Roseburia intestinalis" ,  
          "Blautia obeum" , 
          "Butyricicoccus pullicaecorum",
          "Mediterraneibacter gnavus",
          "Roseburia faecis",
          "Bacteroides fragilis", 
          "Anaerostipes hadrus" , 
          "Blautia hansenii"   , 
          "Clostridium sporogenes" , 
          "Bifidobacterium adolescentis" ,
          "Eubacterium limosum" ,
           "Blautia producta"   , 
          "Enterocloster aldenensis",
             "Bacteroides uniformis"  , 
          "Bacteroides xylanisolvens",  "Collinsella aerofaciens"  , 
          "Phocaeicola dorei",
           "Phocaeicola vulgatus"   ,
          "Bacteroides intestinalis"    ,    
          "Anaerostipes caccae"    ,
          "Bacteroides cellulosilyticus",
         "Parabacteroides merdae" , "Bacteroides caccae","Bacteroides faecis" , "Bacteroides ovatus"   , 
          "Bacteroides thetaiotaomicron"  ,            "Bifidobacterium longum"  ,     "Parabacteroides distasonis"  ,              "Phascolarctobacterium faecium" ,         "Enterococcus faecalis"  , "Escherichia coli"  ,   "Lactiplantibacillus plantarum"  )

rbind(H2O2 %>%  subset(assay %in% "MBC"), O2 %>%  subset(assay %in% "MBC"))->both

setNames(as.data.frame(unique(both[, c("bacteria")])), c("bacteria"))%>%
   left_join(., 
            read.xlsx("P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/00_experimentation/01_pure_culture/20240207_Strain_panel_metabol.xlsx")[, c("bacteria",  "Genus")])%>%
 mutate(bacteria = ifelse(bacteria == "Ruminococcus gnavus", "Mediterraneibacter gnavus", bacteria))%>%
   mutate(Genus = ifelse(bacteria == "Mediterraneibacter gnavus", "Mediterraneibacter", Genus))%>%
   mutate(bacteria = ifelse(bacteria == "Prevotella copri", "Segatella copri", bacteria))%>%
   mutate(Genus = ifelse(bacteria == "Segatella copri", "Segatella", Genus))-> sub_phyl

sub_phyl$stress <-"phlo"
setNames(sub_phyl, c("bacteria", "tolerance", "stress"))%>%
  rbind(., both[, c("bacteria", "tolerance", "stress")])%>%
  mutate(stress = factor (stress, levels = c("phlo", "H2O2", "O2")))->sub


sub%>%

  ggplot(aes(y=bacteria, x=stress))+
   new_scale_fill() +
  
   geom_tile(data = subset(sub, stress %in% c("phlo"))%>%
               mutate(bacteria = factor (bacteria, levels = c(order))), aes(fill = tolerance ), colour="black", aplpha=0.75) +
   scale_fill_manual(values=c("#1f78b4","#a6cee3", "#80b1d3","#8dd3c7","#33a02c", "pink3","#e31a1c", "#ff7f00",   "#fb9a99",  "#6a3d9a", "#cab2d6","#fdbf6f", "darkgoldenrod2", "#b15928",  "#fb8072", "#5c342c", "#b36966", "#A8AF51" , "#484939", "#CE6C4E", "#E6BAAC",  "#B1A7CD", "darksalmon", "lightgreen"), name="Genus")+
  guides(fill = guide_legend(nrow = 4))+
  
   new_scale_fill() +
   geom_tile(data = sub%>%
               subset(., stress %in% c("O2"))%>%
               mutate(tolerance = as.numeric(tolerance))%>%
               mutate(bacteria = factor (bacteria, levels = c(order))), aes(fill = tolerance ), colour= "black") +
  scale_fill_continuous(
    name = expression(O[2]),
    guide = guide_colorbar(title.position = "top", title.vjust = 0.5))+
  
  new_scale_fill() +
   geom_tile(data = sub%>%
               subset(., stress %in% c("H2O2"))%>%
               mutate(tolerance = as.numeric(tolerance))%>%
               mutate(bacteria = factor (bacteria, levels = c(order))), colour= "black",aes(fill = tolerance )) +
  scale_fill_continuous(

    name = "Scaled tolerance",
    guide = guide_colorbar(title.position = "top", title.vjust = 0.5))+
  

  facet_grid(cols=vars(stress), scales="free", space = "free")+
  theme(text= element_text(size=15),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        panel.border=element_rect(fill=NA), 
        panel.background = element_rect(fill="white"),
        rect = element_rect(fill = "transparent"),
        legend.text.align = 0, 
        panel.spacing = unit(0, "lines"),
        axis.text.x = element_blank(),
        axis.text.y = element_text(face="italic"),
        strip.background = element_blank(), 
        strip.text.x = element_text(angle=0))

ggsave(filename = "species_rolerances..png",   
         path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final",
       device = "png", bg= "transparent", 
         width = 50,
         height = 19,
         units = c("cm"))



aggregate(both$tolerance, by =list(both$bacteria), FUN="mean")%>%
  subset(x < 0.25)

```

```{r}
 library(pheatmap)
MBC_H2O2<-setNames(tolerances_h2o2[, c("bacteria", "MBC")], c("bacteria", "H2O2"))
MBC_O2<-setNames(df_final_replicates_O2[, c("bacteria", "MBC")], c("bacteria", "O2"))
mat<-left_join( MBC_H2O2, MBC_O2)

unique(mat$O2)
mat %>%
  mutate(O2 = ifelse(O2 == 8, 7, O2)) %>%
  mutate(O2 = ifelse(O2 == 11, 8, O2)) %>%
  mutate(O2 = ifelse(O2 == 12, 9, O2)) %>%
  mutate(O2 = (O2 - 3) / 6) %>%
  mutate(H2O2 = (H2O2 - 7) / 4) %>%
  column_to_rownames("bacteria") %>%
  as.matrix() %>%
  pheatmap(
    cluster_cols = FALSE,
    legend_breaks = c(0, 0.25, 0.5, 0.75, 1),
    method = "ward",
    cellwidth = 10,
    cellheight = 10,
    angle = 90,  # Rotate y-axis text
    fontsize_row = 8,  # Adjust font size
    fontsize_col = 10,
     font.family = "italic" # Adjust font size
  )

setwd("C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final")
ggsave("heatmap..png", p, device = "png", bg= "transparent", width=5, heigh =10)
  
```



```{r}
df_final_replicates_O2$toll <- NA

df_final_replicates_O2%>%
  mutate(toll = ifelse(MIC ==MBC, "yes", toll))%>%
  subset(!toll %in% "yes")%>%
  subset(!MBC %in% 12)

mean(sub$MBC)
```



