---
title: "Relative metabolites O2"
output: html_document
date: '2022-12-29'
---
```{r setup, echo =FALSE}
.libPaths('C:\\Users\\zuendj\\Documents\\R4Libs')
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/zuendj/Desktop/03_data R/oxidative stress", cho = TRUE, warning = FALSE, message = FALSE)

# Install Packages:
# install.packages("rlang")
# install.packages("tidyr")
# install.packages("ggbreak")
# 
# library(ggpmisc)
 # library(tidyverse)
# library(ggplot2)
# library(tidyr)
# # install.packages("devtools")
# # devtools::install_github("slowkow/ggrepel")
# library(ggrepel)
# library(ggbreak)
 # library(openxlsx)
# library(readxl)
# library(ggh4x)
# library(ggpubr)
# library(ggpubr)
 # library(rstatix)
# library(dplyr)
# library(scales)
library(readxl)
```

# 1. Read all the data into one df
```{r}
setwd("P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/02_hplc_pure/01_O2/01_import")
bacteria <- list.files(path = "P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/02_hplc_pure/01_O2/01_import")
```
## aerobic and anaerobic values go next to each other in order to do correlation: blank correction
```{r}
setwd("P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/02_hplc_pure/01_O2/01_import")

#########################################
#blank correction
########################################

#define empty matrices, no values added so far
HPLC_values_blank_corrected<- matrix(0, 0, 5)
HPLC_values_wide<- matrix(0, 0, 13)

#room for bacteria, phyla, agar condition,growth condition, aerobic/anaerobic and 10 metabolites, each species has 21 data point which are 7 triplicates
#HPLC_values_relative_mean<-matrix(0, 6, 15)#same matrix as above but with mean of the triplicates

#column names for matrix
col_names_HPLC <- list("sample_ID", "condition_agar", "condition", "succinate", "lactate", "formate", "acetate","propionate", "isobutyrate", "butyrate", "ethanol", "isovalerate", "valerate")

colnames(HPLC_values_wide)<- col_names_HPLC

for (l in 1:length(bacteria)) {
  #generate MICwide format

  HPLCwide = read_excel(bacteria[l])
  # print(bacteria[l])
  aerobe<-as.data.frame(HPLCwide[c(2:25), c(7,3,5,17:26)])
  colnames(aerobe)<- col_names_HPLC
  # aerobe[is.na(aerobe)] <- 0

aerobe <-cbind(aerobe, HPLCwide[c(2:25),c(13)] )  
##blank correction all (aerobe from 1 to 15 and anaerobe form 15 23)
for (i in 1:3) {
  for (z in 4:12) {
    aerobe[i, z] = as.numeric(aerobe[i, z]) - as.numeric(aerobe[4, z])
  }}
  for (i in 5:7) {
    for (z in 4:12) {
      aerobe[i, z] = as.numeric(aerobe[i, z]) - as.numeric(aerobe[8, z])
    }}
  for (i in 9:11) {
    for (z in 4:12) {
      aerobe[i, z] = as.numeric(aerobe[i, z]) - as.numeric(aerobe[12, z])
    }}

  for (i in 13:15) {
    for (z in 4:12) {
      aerobe[i, z] = as.numeric(aerobe[i, z]) - as.numeric(aerobe[16, z])
    }}
  for (i in 17:19) {
    for (z in 4:12) {
      aerobe[i, z] = as.numeric(aerobe[i, z]) - as.numeric(aerobe[20, z])
    }}
  for (i in 21:23) {
    for (z in 4:12) {
      aerobe[i, z] = as.numeric(aerobe[i, z]) - as.numeric(aerobe[24, z])
    }}

  aerobe<-aerobe[-c(4,8,12,16,20,24),]
  
  

  aerobe$assay <- gsub("_", " ", as.character(strsplit(bacteria[l], ".xls")))


  HPLC_values_wide<-rbind(HPLC_values_wide, aerobe, stringsAsFactors = FALSE)
  HPLC_values_wide$condition_agar[is.na(HPLC_values_wide$condition_agar)] <- "agar with highest growth"

  
}


```
#conditions to analyze
```{r}
highest_agar<-("Bacteroides fragilis")
highest<-c("Lachnospira eligens","Agathobacter rectalis","Bacteroides intestinalis","Prevotella copri", "Bacteroides xylanisolvens" , "Bacteroides faecis", "Bacteroides uniformis","Bacteroides ovatus" ,
"Phocaeicola vulgatus", "Dorea longicatena")
lowest<- c("Anaerostipes caccae", 
"Anaerostipes hadrus", 
"Anaerotignum lactatifermentans" ,    
"Bacteroides caccae",  
"Bacteroides thetaiotaomicron" ,
         
 "Bifidobacterium adolescentis",
"Bifidobacterium longum",
"Blautia hansenii", 
"Blautia hydrogenotrophica"  ,        
 "Blautia obeum",  
"Blautia producta", 
"Clostridium sporogenes"  ,
"Collinsella aerofaciens" ,           
"Enterococcus faecalis",
"Escherichia coli",   
"Eubacterium limosum", 
"Faecalibacterium duncaniae",
"Lactiplantibacillus plantarum" ,
"Parabacteroides merdae" ,            
"Phascolarctobacterium faecium",
"Roseburia intestinalis" , 
"Ruminococcus gnavus"  )
```

### get row names & recode variables
```{r}
#export wide file
write.xlsx(HPLC_values_wide, file = "P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/02_hplc_pure/01_O2/O2_hplc_blank_corrected.xlsx")
#export file with wel columns of the conditions analyzed
write.xlsx(HPLC_values_wide[, c("assay","condition_agar", "well column", "condition")]%>%
             subset((assay %in% c(lowest) & condition_agar %in% "lowest" )|
          (assay %in% c(highest) & condition_agar %in% "agar with highest growth" )|
           (assay %in% c(highest_agar) & condition_agar %in% "highest" ))%>%
            subset(condition %in% "aerobe"), file = "P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/02_hplc_pure/01_O2/sampling_O2.xlsx")
```


# 4. Calculate relative values: mM and total C
```{r}
##values have to be numeric
numeric_columns <- c("succinate", "lactate", "formate", "propionate", "butyrate", "acetate", "valerate", "isobutyrate", "isovalerate")
HPLC_values_wide[, numeric_columns] <- sapply(HPLC_values_wide[, numeric_columns], as.numeric)

#drop ethanol etc.
HPLC_values <- HPLC_values_wide[, !colnames(HPLC_values_wide) %in% c("isobutyrate","ethanol", "isovalerate","valerate")]

#define new df for relative values
HPLC_values_relative_wide <- HPLC_values

#calculate relative values; first calculate total SCFA
HPLC_values$total_SCFA<-0

scfas<-c("succinate", "lactate" , "formate", "acetate","propionate","butyrate")
scfas_rel<-paste0(scfas, "_rel_mM")
scfas_rel_C<-paste0(scfas, "_rel_C")
number_c <- c(4, 3, 1, 2, 3, 4) #nuber of c atoms for each organic acid

###calculate the relative values based on mM
for (i in 1:nrow(HPLC_values_relative_wide)){
  x<-HPLC_values_relative_wide[i,c(scfas)]
  HPLC_values_relative_wide[i,c("total_mM")]<-  sum(x[which(x>0)])
  
  for (j in 1:length(scfas)){
    HPLC_values_relative_wide[i,scfas_rel[j]] = HPLC_values_relative_wide[i,scfas[j]] / HPLC_values_relative_wide[i, c("total_mM")]
  }
}

## calculate the relative values based in total C
for (i in 1:nrow(HPLC_values_relative_wide)){
  vec<-c(HPLC_values_relative_wide[i,c("lactate")]*3,
         HPLC_values_relative_wide[i,c("formate")]*1,
         HPLC_values_relative_wide[i,c("succinate")]*4,
         HPLC_values_relative_wide[i,c("butyrate")]*4,
         HPLC_values_relative_wide[i,c("propionate")]*3,
         HPLC_values_relative_wide[i,c("acetate")]*2)
  HPLC_values_relative_wide[i,c("total_C")]<-  sum(vec[which(vec>0)])
  
  for (j in 1:length(scfas)){
    HPLC_values_relative_wide[i,scfas_rel_C[j]] = HPLC_values_relative_wide[i,scfas[j]]*number_c[j]/ HPLC_values_relative_wide[i, c("total_C")]
  }
}


```

```{r}
# HPLC_values_relative_wide %>%
#   gather(metabolite, rel, "succinate_rel_C":"butyrate_rel_C")%>%
#   ggplot(aes(x=interaction(sample_ID), y=rel))+
#   geom_bar(position="stack", stat="identity", aes(fill=metabolite))
```


# Select the conditions to analyze
```{r}
library(ggpubr)
#calculate mean and sd for bar plot
HPLC_values_relative_wide%>%
 subset((assay %in% c(lowest) & condition_agar %in% "lowest" )|
          (assay %in% c(highest) & condition_agar %in% "agar with highest growth" )|
           (assay %in% c(highest_agar) & condition_agar %in% "highest" ))->HPLC_relative_select
``` 

# Test statistical differences in total metabolites
```{r}
#get means and sd
left_join(setNames(aggregate(HPLC_relative_select$total_C, by=list(HPLC_relative_select$condition, HPLC_relative_select$condition_agar, HPLC_relative_select$assay), FUN=sd, na.rm=T), c("condition", "condition_agar", "assay", "sd")),
setNames(aggregate(HPLC_relative_select$total_C, by=list(HPLC_relative_select$condition, HPLC_relative_select$condition_agar, HPLC_relative_select$assay), FUN=mean, na.rm=T), c("condition", "condition_agar", "assay", "mean")))->mean

#test for normality
HPLC_relative_select[, c("assay", "condition","condition_agar", "total_C")]%>%
  group_by(condition_agar, assay, condition)%>%
  shapiro_test(total_C)%>%
  subset(p<0.05)->shapi

shapi[, c("assay", "condition_agar")]->shapi
shapi$sig <- "non_normal"

#get the MBC tolerance to order the plot - add 1 to the faculative anaerobes so that they are in the bottom row
order_o2<- read.xlsx("P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/01_OD_pure/01_O2/O2_tolerances.xlsx")%>%
  mutate(MBC = ifelse(bacteria == "Escherichia coli" |
                      bacteria == "Lactiplantibacillus plantarum" |
                        bacteria == "Enterococcus faecalis" |
                        bacteria == "Bifidobacterium longum", 13, MBC))


#t test
HPLC_relative_select[, c("assay", "condition","condition_agar", "total_C")]%>%
  left_join(., setNames(order_o2, c("assay", "MIC", "MBC")))%>%
  mutate(assay = factor(assay))%>%
  mutate(assay = reorder(assay, MBC))%>%
  group_by(condition_agar, assay)%>%
  t_test(total_C ~ condition)%>%
  add_significance()%>%
  add_xy_position(fun="max")%>%
  mutate(xmin=0.5,
         xmax=1.5)%>%
  mutate(p.signif = ifelse(assay == shapi$assay | condition_agar == shapi$condition_agar, "ns", p.signif))->stats


mean%>%
  left_join(., setNames(order_o2, c("assay", "MIC", "MBC")))%>%
  mutate(assay = factor(assay))%>%
  mutate(assay = reorder(assay, MBC))-> sub

sub%>%
  ggplot(aes(y=mean, x=assay))+
  theme(text=element_text(size=15),
          strip.text.x = element_text(angle=90))+
  # geom_point()+
  geom_col(aes(fill=condition, position = condition), position = position_dodge(0.8), width=0.7)+
  geom_errorbar(aes(ymin=mean - sd, ymax=mean +sd, group=condition), width=.7,position = position_dodge(0.8))+
  stat_pvalue_manual(stats, label="p.signif", coord.flip=T, hide.ns = T )+
  facet_grid(rows=vars(assay), scales = "free")+
  labs(fill="Condition")+
  coord_flip()-> total_C_plot

total_C_plot


ggsave(filename = paste( "o2_pure cult_tot.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/metabolites",
       width = 20,
       height = 20,
       units = c("cm"))

```


# 6. Perform T test & Shapiro Wilk test to check for normality

```{r}
#transform into long - use values taht are relative to total C
HPLC_relative_long <-  gather(HPLC_relative_select, metabolite, concentration, "succinate_rel_C":"butyrate_rel_C", factor_key = T ) 




# stats<-as.data.frame(matrix(0,0,5))
# names(stats)<-c("condition_agar", "metabolite", "p", "p.signif", "Bacteria")

stats<-matrix(0,0,5)
stats_shapi<-matrix(0,0,5)
bacteria_names<- unique(HPLC_relative_long$assay)
metabol<- unique(HPLC_relative_long$metabolite)
# cond<- unique(HPLC_relative_long$condition_agar)
ana<- unique(HPLC_relative_long$condition)

wil<-"init"
out<-"init"
shapi_out<-"init"
shapi<-"init"

HPLC_relative_long$metabolite<-as.character(HPLC_relative_long$metabolite)

##filter out samples with no growth
HPLC_relative_long%>%
  subset(total_C >1)->HPLC_relative_long


HPLC_relative_long[,-c(1)]%>% #remove first col
  filter(!is.na(concentration))->HPLC_relative_long #get only teh metabolites that are actually produced

#loop for plots for all bacteria
for (i in 1:length(bacteria_names)) {
##create subset
  subset(HPLC_relative_long, assay == bacteria_names[i])->bacteria_subset
  #get the metabolites that are produced as a vector to loop over
  metabol<-unique(bacteria_subset$metabolite)
  
  for (j in 1:length(metabol)){
     subset(bacteria_subset, metabolite == metabol[j])->metabol_subset
    
    #get the agar conditions
   cond<-unique(metabol_subset$condition_agar) 
    #loop over the agar conditions to create the final subset to bet tested
   
    # for (x in 1:length(cond)){
    #  subset(metabol_subset, condition_agar == cond[x])->subset

#remove all df taht contain results
rm(wil) 
rm(out)
rm(shapi_out)
rm(shapi)

#do t test - try() to avoid interruption of loop in case t test cannot be performed (for zero value sor so) 
try(
  metabol_subset %>%
  t_test(data= .,
  formula = concentration ~ condition, ref.group = 
    "anaerobe")%>%
  add_significance()%>% 
  add_xy_position(x="metabolite") -> wil
  , 
  silent = T)
    
    
try(wil$Bacteria<- bacteria_names[i], silent=T)
try(wil$agar_condition<- unique(metabol_subset$condition_agar), silent=T)
try(wil$metabolite<- metabol[j], silent=T)
try(out<-wil[, c("agar_condition", "metabolite", "p", "p.signif", "Bacteria")], silent=T)

# do shapiro tets for the anaerobic and aerobic samples individually
for (z in 1:length(ana)){
try(
  metabol_subset%>%
    subset(condition %in% ana[z])%>%
     shapiro_test(concentration) -> shapi_out
  , silent = T)
  
try(shapi<-shapi_out[, c("p")], silent=T)
try(shapi$Bacteria<- bacteria_names[i], silent=T)
try(shapi$agar_condition<- unique(metabol_subset$condition_agar), silent=T)
try(shapi$metabolite<- metabol[j], silent=T)
try(shapi$condition<- ana[z], silent=T)
try(stats_shapi<-rbind(stats_shapi, shapi), silent=T)}
try(stats<-rbind(stats, out), silent=T)

  }}
# }



#to match the formate, pivot wider shapiro
shapiro_wide<-setNames(pivot_wider(stats_shapi, values_from = p, names_from = condition), c( "Bacteria", "agar_condition", "metabolite", "p_shapiro_aerobe", "p_shapiro_anaerobe"))
```


# 4. Calculate means & SD 
```{r}
left_join(
  setNames(aggregate(HPLC_relative_long$concentration, list(HPLC_relative_long$assay, HPLC_relative_long$condition_agar, HPLC_relative_long$metabolite, HPLC_relative_long$condition), FUN=mean, na.rm=T), c("Bacteria",  "agar_condition", "metabolite", "condition", "mean")),
setNames(aggregate(HPLC_relative_long$concentration, list(HPLC_relative_long$assay, HPLC_relative_long$condition_agar, HPLC_relative_long$metabolite, HPLC_relative_long$condition), FUN=sd, na.rm=T),
         c("Bacteria",  "agar_condition", "metabolite", "condition" ,"sd")))->means_long

means_long%>%
  pivot_wider(names_from = condition, values_from = c(sd, mean)) ->means_corr

```

## Merge mean values with statistics

```{r}
#combine means, stats and shapiro stats
left_join(means_corr, stats)%>%
  left_join(., shapiro_wide) -> means_stats_shapiro

#make new col for shapiro value => get the lower of the two shapiro values
means_stats_shapiro$shapiro_min<-0

#get the minimal p value of the shapiro test (if one of the group has p > 0.05 we cannot use t test)
for (i in 1:nrow(means_stats_shapiro)){
x<-as.data.frame(c(means_stats_shapiro[i,"p_shapiro_aerobe"], means_stats_shapiro[i,"p_shapiro_anaerobe"]))
x<-x[!is.na(x)]
means_stats_shapiro[i, "shapiro_min"] <- min(x)} #if the smaller of the two p values is <0.05 => make p value of t test ns
means_stats_shapiro$p_corrected<-means_stats_shapiro$p.signif

for (i in 1:nrow(means_stats_shapiro)){
    if (means_stats_shapiro[i,c("shapiro_min")] <= 0.05 |
        is.na(means_stats_shapiro[i,c("shapiro_min")])){
    means_stats_shapiro[i,c("p_corrected")] ="ns"}}
```

```{r}
#get the ratios 
means_stats_shapiro%>%
  mutate(ratio = mean_aerobe - mean_anaerobe )%>%
  mutate(agar_condition = as.character(agar_condition))%>%
  mutate(p.signif = ifelse(p.signif =="****", "***", p.signif))->ratio_O2
```

# 4. Calculate means of absolute values: to filter for metabolites that are of low concentration
```{r}
HPLC_abs_long <-  gather(HPLC_relative_select, metabolite, concentration, "succinate":"butyrate", factor_key = T ) 


setNames(aggregate(HPLC_abs_long$concentration, list(HPLC_abs_long$assay, HPLC_abs_long$condition_agar, HPLC_abs_long$metabolite, HPLC_abs_long$condition), FUN=mean), c("Bacteria",  "agar_condition", "metabolite", "condition", "mean_conc"))%>%
  pivot_wider(names_from = condition, values_from = mean_conc)%>%
  setNames(c("Bacteria",  "agar_condition", "metabolite","mean_conc_aerobe_abs", "mean_conc_anaerobe_abs"))%>%
  mutate(metabolite = paste0(metabolite, "_rel_C"))%>%
  mutate(agar_condition = ifelse(agar_condition == "0.15% agar", "highest", agar_condition))%>%
   mutate(agar_condition = ifelse(agar_condition == "minimal agar (''MIC'')", "lowest", agar_condition))%>%
  subset(!is.na(mean_conc_aerobe_abs) & !is.na(mean_conc_anaerobe_abs))->means_absolute #machtch names of metabolites with ratio_O2 to perform left_join

 
```


```{r}
stats_absolute<-left_join(ratio_O2, means_absolute)

stats_absolute$p_corrected_low_conc<-stats_absolute$p_corrected


for (i in 1:nrow(stats_absolute)){
  if (!is.na(stats_absolute[i,"mean_conc_aerobe_abs"]) == T &&
      !is.na(stats_absolute[i,"mean_conc_anaerobe_abs"]) == T &&

      (stats_absolute[i,"mean_conc_aerobe_abs"] > -1 &&
        stats_absolute[i,"mean_conc_aerobe_abs"] <= 0 &&
        stats_absolute[i,"mean_conc_anaerobe_abs"] > -1 &&
        stats_absolute[i,"mean_conc_anaerobe_abs"] <= 0)
      ||
       (stats_absolute[i,"mean_conc_aerobe_abs"] >= 0 &&
        stats_absolute[i,"mean_conc_aerobe_abs"] < 1 &&
        stats_absolute[i,"mean_conc_anaerobe_abs"] >= 0 &&
        stats_absolute[i,"mean_conc_anaerobe_abs"] < 1)
      ||
       (stats_absolute[i,"mean_conc_aerobe_abs"] <= 0 &&
        stats_absolute[i,"mean_conc_aerobe_abs"] > -1 &&
        stats_absolute[i,"mean_conc_anaerobe_abs"] >= 0 &&
        stats_absolute[i,"mean_conc_anaerobe_abs"] < 1)
      ||

      (stats_absolute[i,"mean_conc_aerobe_abs"] >= 0 &&
        stats_absolute[i,"mean_conc_aerobe_abs"] < 1 &&
        stats_absolute[i,"mean_conc_anaerobe_abs"] <= 0 &&
        stats_absolute[i,"mean_conc_anaerobe_abs"] > -1)  ||

      (stats_absolute[i,"mean_conc_aerobe_abs"] <= 0 &&
        stats_absolute[i,"mean_conc_aerobe_abs"] > -1 &&
        stats_absolute[i,"mean_conc_anaerobe_abs"] <= 0 &&
        stats_absolute[i,"mean_conc_anaerobe_abs"] > -1)
      ||

      (stats_absolute[i,"mean_conc_aerobe_abs"] >= 0 &&
        stats_absolute[i,"mean_conc_aerobe_abs"] < 1 &&
        stats_absolute[i,"mean_conc_anaerobe_abs"] >= 0 &&
        stats_absolute[i,"mean_conc_anaerobe_abs"] < 1)){
   stats_absolute[i,"p_corrected_low_conc"] = "ns"
  }
}


```

## selection of the condition to analyze based on total metabolite production => go for the non-sig. different or the condition with the smallest difference
```{r}
selection<-stats_absolute
selection$p_corrected_low_conc[selection$p_corrected_low_conc=="ns"]<-NA
```

```{r}
# taxonomy = read.csv(file = "P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/Strain_panel.csv", sep=";", header=T, check.names =F)
# colnames(taxonomy)[colnames(taxonomy) == "bacteria"] <- "Bacteria"
# 
# tax<-left_join(selection[, -c(15)], taxonomy)
```

```{r}
 library(scales)

selection%>%
  # filter(Bacteria %in% filter_tax)%>%
  mutate(metabolite = factor(metabolite, levels= c("succinate_rel_C","lactate_rel_C", "formate_rel_C" ,"acetate_rel_C","butyrate_rel_C","propionate_rel_C")))%>%
  mutate(ratio = (mean_aerobe - mean_anaerobe))%>%
  mutate(ratio = ratio *100)->sub


max<-max(sub$ratio)
min<-min(sub$ratio)
# 
# unique(sub$Bacteria)

sub%>%
  left_join(., setNames(order_o2, c("Bacteria", "MIC", "MBC")))%>%
  mutate(Bacteria = factor(Bacteria))%>%
  mutate(Bacteria = reorder(Bacteria, MBC))%>%
  ggplot( aes(x=metabolite, y=Bacteria, fill=ratio))+
  geom_tile()+
  scale_fill_gradientn(colors = c( "coral", "white", "cyan4"),  na.value= "grey", values = rescale(c(-120, 0, 120)), limits=c(-120, 120))+
  facet_grid(rows=vars(Bacteria), scales="free")+
    theme(strip.text.y = element_text(angle=0,  face="italic"),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90,  vjust=0.5, hjust=1),
        axis.text.y = element_text( face="italic"), 
        panel.border=element_rect(fill=NA), 
        panel.background = element_rect(fill="white"),
        legend.text.align = 0, 
        strip.background = element_blank(),
        panel.spacing = unit(0, "lines"))+
  labs(fill="Delta [%]")+
  scale_x_discrete(labels=c("succinate_rel_C"="succinate",  "lactate_rel_C" ="lactate" ,  "formate_rel_C"  ="formate" , "acetate_rel_C"  = "acetate" , "propionate_rel_C"="propionate", "butyrate_rel_C" = "butyrate"))+
  # ggtitle(expression(Metabolic~changes~as~a~response~to~O[2]))+
  geom_text(aes(label=p_corrected_low_conc))-> diff_metabol_plot

ggsave(filename = "O2_heatplot_selection_lowconc.jpeg",   
         path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/metabolites", 
         width = 22,
         height = 40,
         units = c("cm"))

```


merge with total_C

```{r}
 library(cowplot)
total_C_plot %>%
  ggpubr::get_legend() %>%
  ggpubr::as_ggplot()-> totalC_legend 

diff_metabol_plot%>%
  ggpubr::get_legend() %>%
  ggpubr::as_ggplot()  -> metabol_legend 


plot_grid(totalC_legend, metabol_legend, ncol=1 )->legends


plot_grid(
  #plot 1: growth
  growth_o2+ 
    theme(legend.position = "none",
          text=element_text(size=15),
          plot.title = element_text(size=15),
          strip.text.y = element_blank())+
    ggtitle("Growth"), 
  #plot 2: metabol
  total_C_plot+
    theme(legend.position = "none", 
          plot.title = element_text(size=15),
          text=element_text(size=15),
          axis.text.y=element_blank(), 
          axis.title.y=element_blank(), 
          axis.ticks.y=element_blank(),
          strip.text.y = element_blank(), 
          panel.border=element_rect(fill=NA), 
          panel.background = element_rect(fill="white"),
          legend.text.align = 0, 
          strip.background = element_blank(),
          panel.spacing = unit(0, "lines"), 
          legend.text = element_blank())+
            ylab("Total C [mM]")+
            scale_fill_manual(values=c("indianred", "cyan4"))+
    ggtitle("Total C"), 
  #plot3: total c
  diff_metabol_plot+
    theme(  text=element_text(size=15),
            plot.title = element_text(size=15),
            legend.position = "none", 
          axis.text.y=element_blank(),
          strip.text.y = element_text(face="italic", hjust=0),
          axis.ticks = element_blank())+
    ggtitle("Individual metabolites"),
  ##define plot grid 
        align= "h", rel_heights = c(1,1,1), rel_widths=c(0.65,0.4, 1), nrow=1, axis = 'tblr')

ggsave(filename = "metabolt_totalc_O2.jpeg",   
         path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final", 
         width = 21,
         height = 30,
         units = c("cm"))
```






#Figure with Bacteroides
```{r}

HPLC_values_relative_wide%>%
  gather(metabolite, concentration, "succinate":"butyrate")->abs_long


setNames(aggregate(abs_long$concentration, list(abs_long$assay, abs_long$condition_agar, abs_long$metabolite, abs_long$condition), FUN=mean), c("Bacteria",  "agar_condition", "metabolite", "condition", "mean_conc"))%>%
  subset(Bacteria %in% c("Bacteroides caccae", "Bacteroides faecis", "Bacteroides fragilis",  "Bacteroides ovatus", "Bacteroides thetaiotaomicron", "Bacteroides uniformis", "Bacteroides xylanisolvens"))%>%
  mutate(condition = factor(condition, levels=c("anaerobe", "aerobe")))%>%
  ggplot(aes(x=interaction( agar_condition), y=mean_conc))+
  geom_bar(position="stack", stat="identity", aes(fill=metabolite, alpha=condition))+
  facet_grid(cols=vars(Bacteria), rows=vars(condition), scales = "free")+
  scale_alpha_manual(values=c(1, 1))+
  theme(strip.text.y = element_text(angle=0,  face="italic"),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90,  vjust=0.5, hjust=1),
        axis.text.y = element_text( face="italic"), 
        panel.border=element_rect(fill=NA), 
        panel.background = element_rect(fill="white"),
        legend.text.align = 0, 
        strip.background = element_blank(),
        panel.spacing = unit(0, "lines"))
  
```


#Figure with Bacteroides
```{r}
means_long%>%
  mutate(agar_condition = factor(agar_condition, levels=c("lowest", "agar with highest growth", "highest")))%>%
  mutate(metabolite = factor(metabolite, levels=c("lactate_rel_C", "succinate_rel_C", "formate_rel_C", "acetate_rel_C", "propionate_rel_C", "butyrate_rel_C")))%>%
  subset(Bacteria %in% c("Bacteroides_caccae", "Bacteroides_faecis", "Bacteroides_fragilis",  "Bacteroides_ovatus", "Bacteroides_thetaiotaomicron", "Bacteroides_uniformis", "Bacteroides_xylanisolvens"))%>%
  # mutate(condition = factor(condition, levels=c("anaerobe", "aerobe")))%>%
   group_by(condition, agar_condition) %>%
  mutate(new_y = cumsum(mean))%>%
  ggplot(aes(x=interaction(agar_condition, condition), y=mean))+
  geom_bar(position="stack", stat="identity", aes(fill=metabolite))+
  scale_alpha_manual(values=c(1, 1))+
  theme(strip.text.y = element_text(angle=0,  face="italic"),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90,  vjust=0.5, hjust=1),
        axis.text.y = element_text( face="italic"), 
        panel.border=element_rect(fill=NA), 
        panel.background = element_rect(fill="white"),
        legend.text.align = 0, 
        strip.background = element_blank(),
        panel.spacing = unit(0, "lines"))+
    geom_errorbar(aes( ymax = new_y + sd, ymin = new_y - sd, colour = metabolite), 
                width = 0.2)+
 facet_grid(cols=vars(Bacteria))
  
```

#odl code

# 5. Correlation-Plots with means
```{r, echo=FALSE, include=FALSE}
# library(scales)
# 
# for (i in 1:length(bacteria_names)) {
#   # for (i in 1:1) {
# ##create subset 
#     
#     
#     means_corr%>%
#       subset(Bacteria %in% c(bacteria_names[i]))->bacteria_subset
#   
#   plot <-ggplot(subset(bacteria_subset, metabolite %in% c("succinate", "lactate", "formate", "butyrate", "propionate", "acetate")), aes(x=mean_aerobe, y=mean_anaerobe))+
#     geom_abline(intercept=0, slope=1, color="darkgrey", size=0.75, alpha=0.5, linetype =3)+
#     stat_poly_line(color="darkgrey", b_0.constant =T) +
#     geom_point(aes(colour=metabolite, shape=agar_condition), size=3, alpha=0.5) +
#     geom_errorbar(aes(ymin=mean_anaerobe-sd_anaerobe, ymax=mean_anaerobe+sd_anaerobe, colour=metabolite))+
#     geom_errorbarh(aes(xmin=mean_aerobe-sd_aerobe, xmax=mean_aerobe+sd_aerobe, colour=metabolite))+
#     coord_fixed()+
#     #expand_limits(x=c(0.05, 1), y=c(0.05, 1))+
#     # geom_abline(intercept=0.0, slope=1.1111111, color="darkgrey", size=0.75, alpha=0.5, linetype =3)+
#     # geom_abline(intercept=-0.0, slope=0.9, color="darkgrey", size=0.75, alpha=0.5, linetype =3 )+
#     ggtitle(bacteria_names[i])+
#     labs(x="aerobic", y="anaerobic")+
#     stat_poly_eq(b_0.constant =T)+
#     scale_x_continuous( limits = c(-1, 1), labels = label_number(accuracy = 0.01)) +
#     scale_y_continuous( limits = c(-1, 1), labels = label_number(accuracy = 0.01))
#   show(plot)
#   
#   x<-paste0(bacteria[i], ".jpg")
#   
#   ggsave(filename = x,   path = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/output/correlations_O2", 
#          width = 18,
#          height = 15,
#          units = c("cm"))
#   }

```


# 5. relative plots & Statistics: t-test

```{r}
# bacteria_names<- unique(HPLC_relative_long$row_names_relative_2)
# metabol<- unique(HPLC_relative_long$metabolite)
# cond<- unique(HPLC_relative_long$condition_agar)
# 
# 
# HPLC_relative_long$condition<-factor(HPLC_relative_long$condition, levels=c("anaerobe", "aerobe"))
# 
# wil<-x
# 
# 
# #loop for plots for all bacteria
# for (i in 1:length(bacteria_names)) {
# ##create subset
#   subset(HPLC_relative_long, row_names_relative_2 == bacteria_names[i])->bacteria_subset
#   
# 
# rm(wil) 
#   
# try(
#   bacteria_subset %>%
#   filter(!is.na(concentration)) %>%
#   group_by(metabolite, condition_agar) %>%
#   t_test(data= .,
#   formula = concentration ~ condition, ref.group = 
#     "anaerobe")%>%
#   add_significance()%>% 
#   add_xy_position(x="metabolite") -> wil
#   
#   
#   , silent = T)
#     
#     
#   t<-paste0(bacteria_names[i])
#   
#   
#   p_individual<-ggplot(subset(bacteria_subset, metabolite %in% c("succinate", "lactate", "formate", "acetate", "propionate", "butyrate")), aes(x=metabolite, y=concentration)) +
#     stat_summary(fun = mean, geom = "bar", position = position_dodge(1), color = "black", aes(fill=condition)) +
#     stat_summary(fun.data = mean_se, geom = "errorbar" , width = .5, position = position_dodge(1), aes(group=condition)) +
#     #scale_fill_manual(values=c('darkgoldenrod1', "cyan4"))+
#     facet_grid( rows=vars(condition_agar), scales = "free", switch="y")+
#     #geom_hline(yintercept=0)+
#     theme(axis.title.x = element_blank(),
#           # axis.text.x = element_blank(),
#           axis.ticks.x = element_blank() )+
#     ggtitle(t)+
#     ylab("metabolites produced [mM]")
#   
#   try(show(p_individual+stat_pvalue_manual(wil, hide.ns = F, label = "p.signif", tip.length = 0)), silent=T)
#   
#   
#   x<-paste0(bacteria_names[i], ".jpg")
# 
#   ggsave(filename = x, path = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/output/relative_O2")
#   }
```


# 3. Plots: absolute data, loop over each taxa
```{r}
# #loop for plots for all bacteria
# for (i in 1:length(bacteria_names)) {
# ##create subset
#   subset<- HPLC_blank_corrected_O2$row_names == bacteria_names[i]
#   bacteria_subset<- HPLC_blank_corrected_O2[subset,]
# 
#   t<-paste0(bacteria_names[i])
# 
# ggplot(subset(bacteria_subset, metabolite %in% c("succinate", "lactate", "formate", "acetate", "propionate", "butyrate")), aes(x=metabolite, y=concentration, fill=condition)) +
#     stat_summary(fun = mean, geom = "bar", position = position_dodge(1), color = "black") +
#     stat_summary(fun.data = mean_se, geom = "errorbar" , width = .5, position = position_dodge(1)) +
#     #scale_fill_manual(values=c('darkgoldenrod1', "cyan4"))+
#     facet_grid( rows=vars(condition_agar), cols= vars(metabolite), scales = "free", switch="y")+
#     #geom_hline(yintercept=0)+
#     theme(axis.title.x = element_blank(),
#           axis.text.x = element_blank(),
#           axis.ticks.x = element_blank() )+
#     ggtitle(t)+
#     ylab("metabolites produced [mM]")->p
#   #scale_y_continuous(limits = c(-0.5, 1))+
# 
# 
# show(p)
# 
#   x<-paste0(bacteria[i], ".jpg")
# 
#   # ggsave(filename = x, path = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/output/absolute_O2")
# }
```
