---
title: "Relative metabolites H2O2"
output: html_document
date: '2022-12-29'
---
```{r setup, echo =FALSE}
.libPaths('C:\\Users\\zuendj\\Documents\\R4Libs')
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/zuendj/Desktop/03_data R/oxidative stress", cho = TRUE, warning = FALSE, message = FALSE)

# Install Packages:
# install.packages("rlang")
# install.packages("tidyr")
# install.packages("ggbreak")

library(ggpmisc)
library(tidyverse)
library(ggplot2)
library(tidyr)
# install.packages("devtools")
# devtools::install_github("slowkow/ggrepel")
library(ggrepel)
library(ggbreak)
library(openxlsx)
library(readxl)
library(ggh4x)
library(ggpubr)
library(ggpubr)
library(rstatix)
library(dplyr)
```

# 1. Read all the data into one df
```{r}
setwd("C:/Users/zuendj/Desktop/03_data R/oxidative stress/input/h2o2")
hplc_raw <- read.csv(file= "20221220_final mic_hplc data_corrected blank_only_knon_metabolites.csv", sep=";")
```

```{r}
hplc_raw%>%
  subset(remove_for_data_analysis != "yes")%>%
  subset(condition != "blank")->subset

# subset[is.na(subset[])] <- 0


subset<-subset[, c("assay", "condition", "Succinate", "Lactate", "Formate", "Acetate", "Propionate", "Butyrate")]

names(subset)<-c("assay", "condition", "succinate", "lactate", "formate", "acetate", "propionate", "butyrate")
```


# 2. Calculate relative values
```{r}

#define new df for relative values
HPLC_values_relative_wide <- subset

HPLC_values<- subset
#calculate relative values; first calculate total SCFA
HPLC_values$total_SCFA<-0

scfas<-c("succinate", "lactate", "formate", "acetate", "propionate", "butyrate")

for (i in 1:nrow(HPLC_values)){
  x<-HPLC_values[i,c("succinate", "lactate", "formate", "acetate", "propionate", "butyrate")]
  HPLC_values[i,c("total_SCFA")]<- sum(x[!is.na(x) & x > 0], na.rm=T)
  
  for (j in 1:length(scfas)){
    HPLC_values_relative_wide[i,scfas[j]] = HPLC_values[i,scfas[j]] / HPLC_values[i, c("total_SCFA")]
  }
}
```

```{r}
#transform into long
HPLC_relative_long <-  gather(HPLC_values_relative_wide, metabolite, concentration, "succinate":"butyrate", factor_key = T ) 
```


# 3. Calculate means & SD -> replicates do not correspont to each other -> better display mean

```{r}
#calculate means & sd
mean_metabolites <- aggregate(HPLC_relative_long$concentration, list(HPLC_relative_long$assay, HPLC_relative_long$condition, HPLC_relative_long$metabolite), FUN=mean)
sdv <- aggregate(HPLC_relative_long$concentration, list(HPLC_relative_long$assay, HPLC_relative_long$condition, HPLC_relative_long$metabolite), FUN=sd)

names(mean_metabolites)<-c("Bacteria",  "condition", "metabolite",  "mean_conc")
names(sdv)<-c("Bacteria",  "condition", "metabolite",  "sd")

#long table with all means
total_means<-left_join(mean_metabolites, sdv)
```

## wide table for correlation plots
```{r}
#separate table for all the conditions-> both NIC and NIC-1 have to be linked with control,...
total_means%>%
  subset(condition %in% c("control"))->control_means

names(control_means)[names(control_means) == "sd"] <- "sd_control"
names(control_means)[names(control_means) == "mean_conc"] <- "mean_control"
control_means <- control_means[, !colnames(control_means) %in% c("condition")]


total_means%>%
  subset(condition %in% c("NIC"))->NIC_means


names(NIC_means)[names(NIC_means) == "sd"] <- "sd_h2o2"
names(NIC_means)[names(NIC_means) == "mean_conc"] <- "mean_h2o2"
NIC_means <- NIC_means[, !colnames(NIC_means) %in% c("condition")]


total_means%>%
  subset(condition %in% c("NIC-1"))->NIC_1_means

names(NIC_1_means)[names(NIC_1_means) == "sd"] <- "sd_h2o2"
names(NIC_1_means)[names(NIC_1_means) == "mean_conc"] <- "mean_h2o2"
NIC_1_means <- NIC_1_means[, !colnames(NIC_1_means) %in% c("condition")]

#####
#this generates a wide table with the sd and means of controls and h2o2 next to each other... controls are duplicated in the dataframe
means_corr_nic<-inner_join(control_means, NIC_means)
means_corr_nic$condition<-"NIC"


means_corr_nic_1<-inner_join(control_means, NIC_1_means)
means_corr_nic_1$condition<-"NIC-1"


all_means<-rbind(means_corr_nic, means_corr_nic_1)
```

# 4. Plots correlation of means

```{r}
setwd("C:/Users/zuendj/Desktop/03_data R/oxidative stress/output")
library(scales)

bacteria_names<-unique(all_means$Bacteria)
# 
# 
# for (i in 1:length(bacteria_names)) {
#   # for (i in 1:1) {
# ##create subset 
#     
#     
#     all_means%>%
#       subset(Bacteria %in% c(bacteria_names[i]))->bacteria_subset
#   
#   plot <-ggplot(bacteria_subset, aes(y=mean_control, x=mean_h2o2))+
#     geom_abline(intercept=0, slope=1, color="darkgrey", size=0.75, alpha=0.5, linetype =3)+
#     stat_poly_line(color="darkgrey", b_0.constant =T) +
#     geom_point(aes(colour=metabolite, shape=condition), size=3, alpha=0.5) +
#     geom_errorbar(aes(ymin=(mean_control - sd_control), ymax=(mean_control + sd_control), colour=metabolite, group=condition))+
#     geom_errorbarh(aes(xmin=mean_h2o2 - sd_h2o2, xmax=mean_h2o2 + sd_h2o2, colour=metabolite, group=condition))+
#     coord_fixed()+
#     #expand_limits(x=c(0.05, 1), y=c(0.05, 1))+
#     # geom_abline(intercept=0.0, slope=1.1111111, color="darkgrey", size=0.75, alpha=0.5, linetype =3)+
#     # geom_abline(intercept=-0.0, slope=0.9, color="darkgrey", size=0.75, alpha=0.5, linetype =3 )+
#     ggtitle(bacteria_names[i])+
#     labs(x=expression(H[2]*O[2]), y="control")+
#     stat_poly_eq(b_0.constant =T)+
#     scale_x_continuous( limits = c(-1, 1.1), labels = label_number(accuracy = 0.01)) +
#     scale_y_continuous( limits = c(-1, 1.1), labels = label_number(accuracy = 0.01))
#   show(plot)
#   
#   x<-paste0(bacteria_names[i], ".jpg")
#   
  # ggsave(filename = x,   
  #        path = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/output/corr_H2O2", 
  #        width = 18,
  #        height = 15,
  #        units = c("cm"))
  # 
  # }
```

# 5. display absolute values

```{r}
#transform into long
HPLC_absolute_long <-  gather(HPLC_values, metabolite, concentration, "succinate":"butyrate", factor_key = T ) 
```

```{r}
#loop for plots for all bacteria
for (i in 1:length(bacteria_names)) {
##create subset
HPLC_absolute_long%>%
    subset(assay %in% bacteria_names[i])->bacteria_subset

  t<-paste0(bacteria_names[i])

  ggplot(bacteria_subset, aes(x=metabolite, y=concentration, fill=condition)) +
    stat_summary(fun = mean, geom = "bar", position = position_dodge(1), color = "black") +
    stat_summary(fun.data = mean_se, geom = "errorbar" , width = .5, position = position_dodge(1)) +
    #scale_fill_manual(values=c('darkgoldenrod1', "cyan4"))+
    facet_grid( cols= vars(metabolite), scales = "free", switch="y")+
    #geom_hline(yintercept=0)+
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank() )+
    ggtitle(t)+
    ylab("metabolites produced [mM]")
  #scale_y_continuous(limits = c(-0.5, 1))+


  x<-paste0(bacteria_names[i], ".jpg")

  # ggsave(filename = x, path = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/output/absolute_H2O2",
  #        width = 14,
  #        height = 12,
  #        units = c("cm"))
  }
```

#6. Assign taxonomy to bacteria name


```{r}

#########################################
#Assign phyla to species name
########################################
bacteria_pyla<-as.data.frame(unique(all_means[,1]))

bacteria_pyla$class<-0
bacteria_pyla<-as.data.frame(bacteria_pyla)


names(bacteria_pyla)<-c("Bacteria", "class")

###

pho<-grep("Phocaei", bacteria_pyla$Bacteria)
bacteria_pyla[pho, c(2)] ="Bacteroidia"

bacteroidetes<-grep("Bac", bacteria_pyla$Bacteria)
bacteria_pyla[bacteroidetes, c(2)] ="Bacteroidia"

pre<-grep("Prevo", bacteria_pyla$Bacteria)
bacteria_pyla[pre,2]="Bacteroidia"

parabac<-grep("Parabac", bacteria_pyla$Bacteria)

bacteria_pyla[parabac,2]="Bacteroidia"

blaut<-grep("Blaut", bacteria_pyla$Bacteria)
bacteria_pyla[blaut,2]="Clostridia"

rumi<-grep("Rumi", bacteria_pyla$Bacteria)
bacteria_pyla[rumi,2]="Clostridia"

rose<-grep("Rose", bacteria_pyla$Bacteria)
bacteria_pyla[rose,2]="Clostridia"

clo<-grep("Clo", bacteria_pyla$Bacteria)
bacteria_pyla[clo,2]="Clostridia"

Faec<-grep("Faecalibacterium", bacteria_pyla$Bacteria)

bacteria_pyla[Faec,2]="Clostridia"

flavi<-grep("Flavi", bacteria_pyla$Bacteria)
bacteria_pyla[flavi,2]="Clostridia"

copr<-grep("Copro", bacteria_pyla$Bacteria)
bacteria_pyla[copr,2]="Clostridia"

eu<-grep("Eu", bacteria_pyla$Bacteria)
bacteria_pyla[eu,2]="Clostridia"

flavi<-grep("Flavo", bacteria_pyla$Bacteria)
bacteria_pyla[flavi,2]="Clostridia"

lacto<-grep("Lacti", bacteria_pyla$Bacteria)
bacteria_pyla[lacto,2]="Bacilli"

subdo<-grep("Sub", bacteria_pyla$Bacteria)
bacteria_pyla[subdo,2]="Clostridia"

bif<-grep("Bif", bacteria_pyla$Bacteria)
bacteria_pyla[bif,2]="Actinomycetia"

colin<-grep("Collin", bacteria_pyla$Bacteria)
bacteria_pyla[colin,2]="Coriobacteriia"

anaero<-grep("Anaero", bacteria_pyla$Bacteria)
bacteria_pyla[anaero,2]="Clostridia"

anaro<-grep("Anaro", bacteria_pyla$Bacteria)
bacteria_pyla[anaro,2]="Clostridia"

marvin<-grep("Marvin", bacteria_pyla$Bacteria)
bacteria_pyla[marvin,2]="Clostridia"

phasc<-grep("Phasco", bacteria_pyla$Bacteria)
bacteria_pyla[phasc,2]="Negativicutes"

entero<-grep("Entero", bacteria_pyla$Bacteria)
bacteria_pyla[entero,2]="Bacilli"

esch<-grep("Esch", bacteria_pyla$Bacteria)
bacteria_pyla[esch,2]="Gammaproteobacteria"
#########################

names(bacteria_pyla)<-c("Bacteria", "Class")

```

# 7. Perform T test & Shapiro Wilk test to check for normality
```{r}

# stats<-as.data.frame(matrix(0,0,5))
# names(stats)<-c("condition_agar", "metabolite", "p", "p.signif", "Bacteria")

stats<-matrix(0,0,5)
stats_shapi<-matrix(0,0,5)


bacteria_names<- unique(HPLC_relative_long$assay)
metabol<- unique(HPLC_relative_long$metabolite)
cond<- c("NIC", "NIC-1")



wil<-"init"
out<-"init"
shapi_out<-"init"
shapi<-"init"

HPLC_relative_long%>%
  filter(!is.na(concentration))->lon_na_filtered
  

#loop for plots for all bacteria
for (i in 1:length(bacteria_names)) {
##create subset
  subset(HPLC_relative_long, assay == bacteria_names[i])->bacteria_subset
  
  for (j in 1:length(metabol)){
     subset(bacteria_subset, metabolite == metabol[j])->metabol_subset
    
    for (x in 1:length(cond)){
     subset(metabol_subset, condition %in% c(cond[x], "control"))->subset

rm(wil) 
rm(out)

rm(shapi_out)
rm(shapi)

try(
  subset %>%
  t_test(data= .,
  formula = concentration ~ condition, ref.group = 
    "control")%>%
  add_significance() -> wil
   , 
 silent = T)
    
    
try(wil$Bacteria<- bacteria_names[i], silent=T)
try(wil$condition<- cond[x], silent=T)
try(wil$metabolite<- metabol[j], silent=T)
try(out<-wil[, c("condition", "metabolite", "p", "p.signif", "Bacteria")], silent=T)


try(stats<-rbind(stats, out), silent=T)

ana<-c("control", cond[x])

for (z in 1:length(ana)){
try(
  subset%>%
    subset(condition %in% ana[z])%>%
     shapiro_test(concentration) -> shapi_out
  , silent = T)
  
try(shapi<-shapi_out[, c("p")], silent=T)
try(shapi$Bacteria<- bacteria_names[i], silent=T)
try(shapi$metabolite<- metabol[j], silent=T)
try(shapi$condition<- ana[z], silent=T)
try(stats_shapi<-rbind(stats_shapi, shapi), silent=T)}



}}}

stats
stats_shapi

# %>%
#   filter(p != "NaN")->stats

```


```{r}
stats_shapi_distinct<-distinct(stats_shapi)


shapiro_wide<-pivot_wider(stats_shapi_distinct, values_from = p, names_from = condition)
names(shapiro_wide)<-c( "Bacteria",  "metabolite", "p_shapiro_control", "p_shapiro_NIC", "p_shapiro_NIC_1")
```


## Merge mean values with statistics

```{r}
means_stats<-left_join(all_means, stats)
means_stats_shapiro<-left_join(means_stats, shapiro_wide)

means_stats_shapiro$shapiro_min<-0

for (i in 1:nrow(means_stats_shapiro)){
  
  
x<-c(means_stats_shapiro[i,"p_shapiro_control"], means_stats_shapiro[i,"p_shapiro_NIC"], means_stats_shapiro[i,"p_shapiro_NIC_1"])

x<-x[!is.na(x)]
  
means_stats_shapiro[i, "shapiro_min"] <- min(x)
}


means_stats_shapiro$p_corrected<-means_stats_shapiro$p.signif

# means_stats_shapiro$p_shapiro_aerobe[is.na(means_stats_shapiro$p_shapiro_aerobe)]<-0
# means_stats_shapiro$p_shapiro_anaerobe[is.na(means_stats_shapiro$p_shapiro_anaerobe)]<-0


for (i in 1:nrow(means_stats_shapiro)){
  
    if (!is.na(means_stats_shapiro[i,c("shapiro_min")]) <= 0.05){
    means_stats_shapiro[i,c("p_corrected")] ="ns"}
  }

```


# 8. calculate the difference between treatment & control

```{r}
ratio<-means_stats_shapiro

ratio<-left_join(ratio, bacteria_pyla, by= "Bacteria")

# ratio$mean_control[ratio$mean_control<0.001]<-0
# ratio$mean_h2o2[ratio$mean_h2o2<0.001]<-0

ratio$ratio <-  ratio$mean_h2o2 -ratio$mean_control 




ggplot(subset(ratio, condition %in% "NIC"), aes(x=metabolite, y=Bacteria, fill=ratio))+
  geom_tile()+
  scale_fill_gradientn(colors = c("coral", "white", "cyan4"), values = rescale(c(-0.5,0,0.48)))+
  facet_grid(rows=vars(Class), space="free", scale="free")+
  theme(strip.text.y = element_text(angle=0, size = 15),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, size=15, vjust=0.2),
        axis.text.y = element_text(size=15, face="italic"))+
  labs(fill="Δrel. metabolites")


 ggsave(filename = "H2O2_heatplot.jpeg",   
         path = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/output", 
         width = 22,
         height = 20,
         units = c("cm"))

```


```{r}
ratio%>%
  subset(Bacteria %in% c("Anaerostipes caccae", 
                         "Anaerostipes hadrus", 
                         "Anaerotignum lactatifermentans",
                         "Bacteroides caccae",
                         "Bacteroides faecis",
                         "Bacteroides fragilis",
                         "Bacteroides thetaiotaomicron" ,
                         "Bacteroides uniformis" ,
                         "Phocaeicola vulgatus" , 
                         "Bifidobacterium adolescentis"  , 
                         "Bifidobacterium longum" ,
                         "Blautia hansenii" ,
                         "Blautia hydrogenotrophica",
                         "Blautia obeum"  , "Blautia producta" ,
                         "Enterococcus faecalis"  ,
                         "Escherichia coli" , 
                         "Flavonifractor sp.", 
                         "Lactiplantibacillus plantarum" , 
                         "Parabacteroides merdae"  ,
                         "Phascolarctobacterium faecium", 
                         "Ruminococcus gnavus",
                         "Bacteroides faecis", 
                         "Bacteroides ovatus", 
                         "Eubacterium eligens",
                         "Prevotella copri",
                         "Bacteroides xylanisolvens",
                         "Bacteroides fragilis",
                         "Clostridium sporogenes",
                         "Collinsella aerofaciens" ,
                         "Bacteroides intestinalis",
                         "Ruminococcus bromi",
                         "Marvinbryantia formatixigens",
                         "Anaerobutyricum hallii")) %>%
           subset(condition %in% c("NIC"))->sub_nic


ratio%>% 
  subset(Bacteria %in% c("Eubacterium limosum",
                         "Roseburia intestinalis")) %>%
           subset(condition %in% c("NIC-1"))->sub_nic_1


selection<-rbind(sub_nic, sub_nic_1)
#unique(ratio$Bacteria)

```


```{r}
selection$p_corrected[selection$p_corrected=="ns"]<-NA

selectionn<-selection[!is.na(selection$ratio),]

max<-max(selectionn$ratio)
min<-min(selectionn$ratio)

ggplot(selectionn, aes(x=metabolite, y=Bacteria, fill=ratio))+
  geom_tile()+
  scale_fill_gradientn(colors = c("coral", "white", "cyan4"), na.value= "grey",values = rescale(c(min,0,max)))+
  facet_grid(rows=vars(Class), space="free", scale="free")+
  theme(strip.text.y = element_text(angle=0, size = 15),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, size=15, vjust=0.2),
        axis.text.y = element_text(size=15, face="italic"))+
  labs(fill="Δrel. metabolites")+
  ggtitle(expression(Metabolic~changes~as~a~response~to~H[2]*O[2]))+
   geom_text(aes(label=p_corrected))


 ggsave(filename = "H2O2_heatplot_selected.jpeg",   
         path = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/output", 
         width = 22,
         height = 20,
         units = c("cm"))
```

# 4. Calculate means of absolute values: to filter for metabolites that are of low concentration

```{r}
mean_metabolites_absolute <- aggregate(HPLC_absolute_long$concentration, list(HPLC_absolute_long$assay, HPLC_absolute_long$metabolite, HPLC_absolute_long$condition), FUN=mean)


names(mean_metabolites_absolute)<-c("Bacteria", "metabolite", "condition", "mean_conc")

mean_metabolites_absolute%>%
  subset(condition %in% c("NIC"))->NIC_mean_metabolites_absolute
names(NIC_mean_metabolites_absolute)[names(NIC_mean_metabolites_absolute) == "mean_conc"] <- "mean_conc_h2o2"
##remove condition colum: otherwise we cannot merge the df's later

mean_metabolites_absolute%>%
  subset(condition %in% c("NIC-1"))->NIC_1_mean_metabolites_absolute
names(NIC_1_mean_metabolites_absolute)[names(NIC_1_mean_metabolites_absolute) == "mean_conc"] <- "mean_conc_h2o2"
##remove condition colum: otherwise we cannot merge the df's later


mean_metabolites_absolute%>%
  subset(condition %in% c("control"))->control_mean_metabolites_absolute
names(control_mean_metabolites_absolute)[names(control_mean_metabolites_absolute) == "mean_conc"] <- "mean_conc_control"
control_mean_metabolites_absolute <- control_mean_metabolites_absolute[, !colnames(control_mean_metabolites_absolute) %in% c("condition")]


means_absolute_NIC<-left_join(control_mean_metabolites_absolute, NIC_mean_metabolites_absolute)
means_absolute_NIC_1<-left_join(control_mean_metabolites_absolute, NIC_1_mean_metabolites_absolute)

means_absolute<-rbind(means_absolute_NIC, means_absolute_NIC_1)
```


```{r}
stats_absolute<-merge(ratio, means_absolute)

stats_absolute$p_corrected_low_conc<-stats_absolute$p_corrected

for (i in 1:nrow(stats_absolute)){
  if (
    !is.na(stats_absolute[i,"mean_conc_control"]) && 
      !is.na(stats_absolute[i,"mean_conc_h2o2"]) &&
      (
        (stats_absolute[i,"mean_conc_control"] > -1 && 
         stats_absolute[i,"mean_conc_control"] < 0 && 
         stats_absolute[i,"mean_conc_h2o2"] > -1 && 
         stats_absolute[i,"mean_conc_h2o2"] < 0) 
        || 
        (stats_absolute[i,"mean_conc_control"] > 0 && 
         stats_absolute[i,"mean_conc_control"] < 1 &&
         stats_absolute[i,"mean_conc_h2o2"] > 0 && 
         stats_absolute[i,"mean_conc_h2o2"] < 1)
        ||
        (stats_absolute[i,"mean_conc_control"] < 0 && 
         stats_absolute[i,"mean_conc_control"] > -1 &&
         stats_absolute[i,"mean_conc_h2o2"] > 0 && 
         stats_absolute[i,"mean_conc_h2o2"] < 1)
        ||
        (stats_absolute[i,"mean_conc_control"] > 0 && 
         stats_absolute[i,"mean_conc_control"] < 1 &&
         stats_absolute[i,"mean_conc_h2o2"] < 0 && 
         stats_absolute[i,"mean_conc_h2o2"] > -1)
      )
    ) {
    stats_absolute[i,"p_corrected_low_conc"] = "ns"
}}


```


```{r}
stats_absolute%>%
  subset(Bacteria %in% c("Anaerostipes caccae", 
                         "Anaerostipes hadrus", 
                         "Anaerotignum lactatifermentans",
                         "Bacteroides caccae",
                         "Bacteroides faecis",
                         "Bacteroides fragilis",
                         "Bacteroides thetaiotaomicron" ,
                         "Bacteroides uniformis" ,
                         "Phocaeicola vulgatus" , 
                         "Bifidobacterium adolescentis"  , 
                         "Bifidobacterium longum" ,
                         "Blautia hansenii" ,
                         "Blautia hydrogenotrophica",
                         "Blautia obeum"  , "Blautia producta" ,
                         "Enterococcus faecalis"  ,
                         "Escherichia coli" , 
                         "Flavonifractor sp.", 
                         "Lactiplantibacillus plantarum" , 
                         "Parabacteroides merdae"  ,
                         "Phascolarctobacterium faecium", 
                         "Ruminococcus gnavus",
                         "Bacteroides faecis", 
                         "Bacteroides ovatus", 
                         "Eubacterium eligens",
                         "Prevotella copri",
                         "Bacteroides xylanisolvens",
                         "Bacteroides fragilis",
                         "Clostridium sporogenes",
                         "Collinsella aerofaciens" ,
                         "Bacteroides intestinalis",
                         "Ruminococcus bromi",
                         "Marvinbryantia formatixigens",
                         "Anaerobutyricum hallii")) %>%
           subset(condition %in% c("NIC"))->sub_nic


stats_absolute%>% 
  subset(Bacteria %in% c("Eubacterium limosum",
                         "Roseburia intestinalis")) %>%
           subset(condition %in% c("NIC-1"))->sub_nic_1


selection<-rbind(sub_nic, sub_nic_1)
#unique(ratio$Bacteria)

```


```{r}
selection$p_corrected_low_conc[selection$p_corrected_low_conc=="ns"]<-NA

selectionn<-selection[!is.na(selection$ratio),]

max<-max(selectionn$ratio)
min<-min(selectionn$ratio)

ggplot(selectionn, aes(x=metabolite, y=Bacteria, fill=ratio))+
  geom_tile()+
  scale_fill_gradientn(colors = c("coral", "white", "cyan4"), na.value= "grey",values = rescale(c(min,0,max)))+
  facet_grid(rows=vars(Class), space="free", scale="free")+
  theme(strip.text.y = element_text(angle=0, size = 15),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, size=15, vjust=0.2),
        axis.text.y = element_text(size=15, face="italic"))+
  labs(fill="Δrel. metabolites")+
  ggtitle(expression(Metabolic~changes~as~a~response~to~H[2]*O[2]))+
   geom_text(aes(label=p_corrected_low_conc))


 ggsave(filename = "H2O2_heatplot_selected_low_conc.jpeg",   
         path = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/output", 
         width = 22,
         height = 20,
         units = c("cm"))
```



