---
title: "Relative metabolites O2"
output: html_document
date: '2022-12-29'
---
```{r setup, echo =FALSE}
.libPaths('C:\\Users\\zuendj\\Documents\\R4Libs')
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/zuendj/Desktop/03_data R/oxidative stress", cho = TRUE, warning = FALSE, message = FALSE)

# Install Packages:
# install.packages("rlang")
# install.packages("tidyr")
# install.packages("ggbreak")

library(ggpmisc)
library(tidyverse)
library(ggplot2)
library(tidyr)
# install.packages("devtools")
# devtools::install_github("slowkow/ggrepel")
library(ggrepel)
library(ggbreak)
library(openxlsx)
library(readxl)
library(ggh4x)
library(ggpubr)
library(ggpubr)
library(rstatix)
library(dplyr)
```

# 1. Read all the data into one df
```{r}
setwd("P:/Shared_documents/Janina Zünd/08_phd/09_student thesis/Marina/03_experimentation/03_MIC_MBC/02_hplc results/Single_bacteria_HPLC_blank_corrected - Copy")
bacteria <- list.files(path = "P:/Shared_documents/Janina Zünd/08_phd/09_student thesis/Marina/03_experimentation/03_MIC_MBC/02_hplc results/Single_bacteria_HPLC_blank_corrected - Copy")
```
## aerobic and anaerobic values go next to each other in order to do correlation: blank correction
```{r}
# setwd("P:/Shared_documents/Janina Zünd/08_phd/09_student thesis/Marina/03_experimentation/03_MIC_MBC/02_hplc results/Single_bacteria_HPLC_blank_corrected - Copy")
# 
# #########################################
# #blank correction
# ########################################
# 
# #define empty matrices, no values added so far 
# HPLC_values_blank_corrected<- matrix(0, 0, 5)
# HPLC_values_wide<- matrix(0, 0, 13)
# 
# #room for bacteria, phyla, agar condition,growth condition, aerobic/anaerobic and 10 metabolites, each species has 21 data point which are 7 triplicates 
# #HPLC_values_relative_mean<-matrix(0, 6, 15)#same matrix as above but with mean of the triplicates 
# 
# #column names for matrix 
# col_names_HPLC <- list("sample_ID", "condition_agar", "condition", "succinate", "lactate", "formate", "acetate","propionate", "isobutyrate", "butyrate", "ethanol", "isovalerate", "valerate")
# 
# 
# col_names<- list("sample_ID", "condition_agar", "oxygen_condition", "metabolite", "concentration")
# colnames(HPLC_values_blank_corrected)<- col_names
# colnames(HPLC_values_wide)<- col_names_HPLC
# 
# for (l in 1:length(bacteria)) {
#   #generate MICwide format
#   
#   HPLCwide = read_excel(bacteria[l])
#   # print(bacteria[l])
#   aerobe<-as.data.frame(HPLCwide[c(2:25), c(7,3,5,17:26)])
#   colnames(aerobe)<- col_names_HPLC
#   # aerobe[is.na(aerobe)] <- 0
# 
#   
#   for (i in 1:3) {
# 
#   for (z in 4:12) {
#     aerobe[i, z] = as.numeric(aerobe[i, z]) - as.numeric(aerobe[4, z])
#   }}
#   for (i in 5:7) {
#     for (z in 4:12) {
#       aerobe[i, z] = as.numeric(aerobe[i, z]) - as.numeric(aerobe[8, z])
#     }}
#   for (i in 9:11) {
#     for (z in 4:12) {
#       aerobe[i, z] = as.numeric(aerobe[i, z]) - as.numeric(aerobe[12, z])
#     }}
#   
#   for (i in 13:15) {
#     for (z in 4:12) {
#       aerobe[i, z] = as.numeric(aerobe[i, z]) - as.numeric(aerobe[16, z])
#     }}
#   for (i in 17:19) {
#     for (z in 4:12) {
#       aerobe[i, z] = as.numeric(aerobe[i, z]) - as.numeric(aerobe[20, z])
#     }}
#   for (i in 21:23) {
#     for (z in 4:12) {
#       aerobe[i, z] = as.numeric(aerobe[i, z]) - as.numeric(aerobe[24, z])
#     }}
#   
#   aerobe<-aerobe[-c(4,8,12,16,20,24),]
#   
#   aerobe_long <- aerobe%>%
#     gather(metabolite, concentration, "succinate":"valerate", factor_key = T )
#   
#   HPLC_values_wide<-rbind(HPLC_values_wide, aerobe, stringsAsFactors = FALSE)
#   HPLC_values_wide$condition_agar[is.na(HPLC_values_wide$condition_agar)] <- "agar with highest growth"
#   HPLC_values_blank_corrected<- rbind(HPLC_values_blank_corrected, aerobe_long)}
```

### get row names & recode some variables
```{r}
# HPLC_values_blank_corrected$condition_agar[is.na(HPLC_values_blank_corrected$condition_agar)] <- "agar with highest growth"
# HPLC_values_blank_corrected$condition_agar[HPLC_values_blank_corrected$condition_agar == 'highest'] <- "0.15% agar"
# HPLC_values_blank_corrected$condition_agar[HPLC_values_blank_corrected$condition_agar == 'lowest'] <- "minimal agar (''MIC'')"
# HPLC_values_blank_corrected$condition_agar <-factor(HPLC_values_blank_corrected$condition_agar, levels = c("0.15% agar", "agar with highest growth", "minimal agar (''MIC'')"))
# 
# 
 bacteria_names <- strsplit(bacteria, ".xls")
# bacteria_names <- as.character(bacteria_names)
# row_names<-matrix(0,length(bacteria)*180,1)
# position = 0
# for (i in 1:length(bacteria)) {
#   for (j in 1:180) {
#     row_names[position+j,1] = bacteria_names[i]
#   }
#   position = position+180
# }
# 
# 
# HPLC_values_blank_corrected_names<-cbind(row_names, HPLC_values_blank_corrected)
# HPLC_values_blank_corrected_names$concentration<-as.numeric(HPLC_values_blank_corrected_names$concentration)
```

```{r}
# write.csv(HPLC_values_blank_corrected_names, file = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/metabolites_O2.csv")
# write.csv(HPLC_values_wide, file = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/wide_metabolites_O2.csv")
```

# 2. Load the data from excel

```{r}
HPLC_blank_corrected_O2<- read.csv("C:/Users/zuendj/Desktop/03_data R/oxidative stress/metabolites_O2.csv", sep=",", header=T, check.names =F)
```

# 3. Plots: absolute data

```{r}
#loop for plots for all bacteria
for (i in 1:length(bacteria_names)) {
##create subset
  subset<- HPLC_blank_corrected_O2$row_names == bacteria_names[i]
  bacteria_subset<- HPLC_blank_corrected_O2[subset,]

  t<-paste0(bacteria_names[i])

ggplot(subset(bacteria_subset, metabolite %in% c("succinate", "lactate", "formate", "acetate", "propionate", "butyrate")), aes(x=metabolite, y=concentration, fill=condition)) +
    stat_summary(fun = mean, geom = "bar", position = position_dodge(1), color = "black") +
    stat_summary(fun.data = mean_se, geom = "errorbar" , width = .5, position = position_dodge(1)) +
    #scale_fill_manual(values=c('darkgoldenrod1', "cyan4"))+
    facet_grid( rows=vars(condition_agar), cols= vars(metabolite), scales = "free", switch="y")+
    #geom_hline(yintercept=0)+
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank() )+
    ggtitle(t)+
    ylab("metabolites produced [mM]")->p
  #scale_y_continuous(limits = c(-0.5, 1))+

  
show(p)

  x<-paste0(bacteria[i], ".jpg")

  # ggsave(filename = x, path = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/output/absolute_O2")
}
```

# 4. Calculate relative values
```{r}
##starting again with the wide table
HPLC_values_wide<- read.csv("C:/Users/zuendj/Desktop/03_data R/oxidative stress/wide_metabolites_O2.csv", sep=",", header=T, check.names =F)

##values have to be numeric
HPLC_values_wide$succinate<-as.numeric(HPLC_values_wide$succinate)
HPLC_values_wide$lactate<-as.numeric(HPLC_values_wide$lactate)
HPLC_values_wide$formate<-as.numeric(HPLC_values_wide$formate)
HPLC_values_wide$propionate<-as.numeric(HPLC_values_wide$propionate)
HPLC_values_wide$butyrate<-as.numeric(HPLC_values_wide$butyrate)
HPLC_values_wide$acetate<-as.numeric(HPLC_values_wide$acetate)
HPLC_values_wide$valerate<-as.numeric(HPLC_values_wide$valerate)
HPLC_values_wide$isobutyrate<-as.numeric(HPLC_values_wide$isobutyrate)
HPLC_values_wide$isovalerate<-as.numeric(HPLC_values_wide$isovalerate)

#drop ethanol etc.
HPLC_values<-HPLC_values_wide[,-c(10,12,13,14)]

#define new df for relative values
HPLC_values_relative_wide <- HPLC_values


#calculate relative values; first calculate total SCFA
HPLC_values$total_SCFA<-0

for (i in 1:nrow(HPLC_values)){
  x<-HPLC_values[i,c(5:10)]
  HPLC_values[i,11]<- sum(x[!is.na(x) & x > 0], na.rm=T)
  
  for (j in 5:10){
    HPLC_values_relative_wide[i,j] = HPLC_values[i,j] / HPLC_values[i, 11]
  }
}


```



```{r}
#re-name
##recode some variables
HPLC_values_relative_wide$condition_agar[HPLC_values_relative_wide$condition_agar == 'highest'] <- "0.15% agar"
HPLC_values_relative_wide$condition_agar[HPLC_values_relative_wide$condition_agar == 'lowest'] <- "minimal agar (''MIC'')"
HPLC_values_relative_wide$condition_agar <-factor(HPLC_values_relative_wide$condition_agar, levels = c("0.15% agar", "agar with highest growth", "minimal agar (''MIC'')"))
```


```{r}
#transform into long
HPLC_relative_long <-  gather(HPLC_values_relative_wide, metabolite, concentration, "succinate":"butyrate", factor_key = T ) 


##generate row names
bacteria_names <- strsplit(bacteria, ".xls")
bacteria_names <- as.character(bacteria_names)

row_names_relative<-matrix(0,length(bacteria_names)*18,1)

position = 0
for (i in 1:length(bacteria_names)) {
  for (j in 1:18) {
    row_names_relative[position+j,1] = bacteria_names[i]
  }
  position = position+18
}

row_names_relative_2<- rbind(row_names_relative, row_names_relative, row_names_relative, row_names_relative, row_names_relative, row_names_relative )
HPLC_relative_long<-cbind(row_names_relative_2, HPLC_relative_long)

```


# 5. relative plots & Statistics: t-test

```{r}
stats<-matrix(0,0,5)
bacteria_names<- unique(HPLC_relative_long$row_names_relative_2)
metabol<- unique(HPLC_relative_long$metabolite)
cond<- unique(HPLC_relative_long$condition_agar)


HPLC_relative_long$condition<-factor(HPLC_relative_long$condition, levels=c("anaerobe", "aerobe"))


#loop for plots for all bacteria
for (i in 1:length(bacteria_names)) {
##create subset
  subset(HPLC_relative_long, row_names_relative_2 == bacteria_names[i])->bacteria_subset
  

rm(wil) 
  
try(
  
  
  bacteria_subset %>%
  filter(!is.na(concentration)) %>%
  group_by(metabolite, condition_agar) %>%
  t_test(data= .,
  formula = concentration ~ condition, ref.group = 
    "anaerobe")%>%
  add_significance()%>% 
  add_xy_position(x="metabolite") -> wil
  
  
  , silent = T)
    
    
  t<-paste0(bacteria_names[i])
  
  
  p_individual<-ggplot(subset(bacteria_subset, metabolite %in% c("succinate", "lactate", "formate", "acetate", "propionate", "butyrate")), aes(x=metabolite, y=concentration)) +
    stat_summary(fun = mean, geom = "bar", position = position_dodge(1), color = "black", aes(fill=condition)) +
    stat_summary(fun.data = mean_se, geom = "errorbar" , width = .5, position = position_dodge(1), aes(group=condition)) +
    #scale_fill_manual(values=c('darkgoldenrod1', "cyan4"))+
    facet_grid( rows=vars(condition_agar), scales = "free", switch="y")+
    #geom_hline(yintercept=0)+
    theme(axis.title.x = element_blank(),
          # axis.text.x = element_blank(),
          axis.ticks.x = element_blank() )+
    ggtitle(t)+
    ylab("metabolites produced [mM]")
  
 

try(show(p_individual+stat_pvalue_manual(wil, hide.ns = F, label = "p.signif", tip.length = 0)), silent=T)
try(wil$Bacteria<- bacteria_names[i], silent=T)
try(out<-wil[, c("condition_agar", "metabolite", "p", "p.signif", "Bacteria")], silent=T)


stats<-rbind(stats, out)

  
  x<-paste0(bacteria_names[i], ".jpg")

  ggsave(filename = x, path = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/output/relative_O2")
  }
```
# 6. Perform T test & Shapiro Wilk test to check for normality


```{r}
stats<-matrix(0,0,5)
stats_shapi<-matrix(0,0,5)
bacteria_names<- unique(HPLC_relative_long$row_names_relative_2)
metabol<- unique(HPLC_relative_long$metabolite)
cond<- unique(HPLC_relative_long$condition_agar)
ana<- unique(HPLC_relative_long$condition)


#loop for plots for all bacteria
for (i in 1:length(bacteria_names)) {
##create subset
  subset(HPLC_relative_long, row_names_relative_2 == bacteria_names[i])->bacteria_subset
  
  for (j in 1:length(metabol)){
     subset(bacteria_subset, metabolite == metabol[j])->metabol_subset
    
    for (x in 1:length(cond)){
    
     subset(metabol_subset, condition_agar == cond[x])->subset

rm(wil) 
rm(out)
rm(shapi_out)
rm(shapi)

  
try(
  subset %>%
  filter(!is.na(concentration)) %>%
  t_test(data= .,
  formula = concentration ~ condition, ref.group = 
    "anaerobe")%>%
  add_significance()%>% 
  add_xy_position(x="metabolite") -> wil
  , 
  silent = T)
    
    
try(wil$Bacteria<- bacteria_names[i], silent=T)
try(wil$agar_condition<- cond[x], silent=T)
try(wil$metabolite<- metabol[j], silent=T)
try(out<-wil[, c("agar_condition", "metabolite", "p", "p.signif", "Bacteria")], silent=T)



for (z in 1:length(ana)){
try(
  subset%>%
    subset(condition %in% ana[z])%>%
     shapiro_test(concentration) -> shapi_out
  , silent = T)
  
try(shapi<-shapi_out[, c("p")], silent=T)
try(shapi$Bacteria<- bacteria_names[i], silent=T)
try(shapi$agar_condition<- cond[x], silent=T)
try(shapi$metabolite<- metabol[j], silent=T)
try(shapi$condition<- ana[z], silent=T)


try(stats_shapi<-rbind(stats_shapi, shapi), silent=T)
try(stats<-rbind(stats, out), silent=T)

}}}}

stats%>%
  filter(p != "NaN")->stats

```


```{r}
shapiro_wide<-pivot_wider(stats_shapi, values_from = p, names_from = condition)
names(shapiro_wide)<-c(    "Bacteria", "agar_condition", "metabolite", "p_shapiro_aerobe", "p_shapiro_anaerobe")
```


# 4. Calculate means & SD -> replicates do not correspont to each other -> better display mean

```{r}
mean_metabolites <- aggregate(HPLC_relative_long$concentration, list(HPLC_relative_long$row_names_relative, HPLC_relative_long$condition_agar, HPLC_relative_long$metabolite, HPLC_relative_long$condition), FUN=mean)

sdv <- aggregate(HPLC_relative_long$concentration, list(HPLC_relative_long$row_names_relative, HPLC_relative_long$condition_agar, HPLC_relative_long$metabolite, HPLC_relative_long$condition), FUN=sd)

names(mean_metabolites)<-c("Bacteria",  "agar_condition", "metabolite", "condition", "mean_conc")
names(sdv)<-c("Bacteria",  "agar_condition", "metabolite", "condition", "sd")


total_means<-merge(mean_metabolites, sdv)


total_means%>%
  subset(condition %in% c("aerobe"))->aerobe_means

names(aerobe_means)[6]<-"sd_aerobe"
names(aerobe_means)[5]<-"mean_aerobe"

aerobe_means<-aerobe_means[, -c(4)]

total_means%>%
  subset(condition %in% c("anaerobe"))->anaerobe_means

names(anaerobe_means)[6]<-"sd_anaerobe"
names(anaerobe_means)[5]<-"mean_anaerobe"

anaerobe_means<-anaerobe_means[, -c(4)]

means_corr<-merge(anaerobe_means, aerobe_means)

means_corr<-means_corr[!is.na(means_corr$mean_anaerobe),]
```
## Merge mean values with statistics

```{r}
means_stats<-left_join(means_corr, stats, by = c("Bacteria", "agar_condition", "metabolite"))
means_stats_shapiro<-left_join(means_stats, shapiro_wide, by = c("Bacteria", "agar_condition", "metabolite"))

means_stats_shapiro<-means_stats_shapiro[!is.na(means_stats_shapiro$p),]


means_stats_shapiro$shapiro_min<-0

for (i in 1:nrow(means_stats_shapiro)){
  
  
x<-c(means_stats_shapiro[i,"p_shapiro_aerobe"], means_stats_shapiro[i,"p_shapiro_anaerobe"])

x<-x[!is.na(x)]
  
means_stats_shapiro[i, "shapiro_min"] <- min(x)
                                         }
```

## if shapiro is sig. -> non-normal distribution -> replace p.sig with "ns" (wilcox stest is non-sig. for all)

```{r}
means_stats_shapiro$p_corrected<-means_stats_shapiro$p.signif

# means_stats_shapiro$p_shapiro_aerobe[is.na(means_stats_shapiro$p_shapiro_aerobe)]<-0
# means_stats_shapiro$p_shapiro_anaerobe[is.na(means_stats_shapiro$p_shapiro_anaerobe)]<-0


for (i in 1:nrow(means_stats_shapiro)){
  if (is.na(means_stats_shapiro[i,c("p.signif")])){
    
    means_stats_shapiro[i,c("p_shapiro_aerobe")] = 0
    means_stats_shapiro[i,c("p_shapiro_anaerobe")] = 0}}



  
  
  if (!is.na(means_stats_shapiro[i,c("p_shapiro_anaerobe")]) <= 0.05){
    means_stats_shapiro[i,c("p_corrected")] ="ns"}
  
  if (!is.na(means_stats_shapiro[i,c("p_shapiro_aerobe")])<=0.05){
    means_stats_shapiro[i,c("p_corrected")] ="ns"}
  }


means_stats_shapiro%>%
  filter(p_shapiro_anaerobe <= 0.05)
```


# 5. Correlation-Plots with means

```{r, echo=FALSE, include=FALSE}
library(scales)

for (i in 1:length(bacteria_names)) {
  # for (i in 1:1) {
##create subset 
    
    
    means_corr%>%
      subset(Bacteria %in% c(bacteria_names[i]))->bacteria_subset
  
  plot <-ggplot(subset(bacteria_subset, metabolite %in% c("succinate", "lactate", "formate", "butyrate", "propionate", "acetate")), aes(x=mean_aerobe, y=mean_anaerobe))+
    geom_abline(intercept=0, slope=1, color="darkgrey", size=0.75, alpha=0.5, linetype =3)+
    stat_poly_line(color="darkgrey", b_0.constant =T) +
    geom_point(aes(colour=metabolite, shape=agar_condition), size=3, alpha=0.5) +
    geom_errorbar(aes(ymin=mean_anaerobe-sd_anaerobe, ymax=mean_anaerobe+sd_anaerobe, colour=metabolite))+
    geom_errorbarh(aes(xmin=mean_aerobe-sd_aerobe, xmax=mean_aerobe+sd_aerobe, colour=metabolite))+
    coord_fixed()+
    #expand_limits(x=c(0.05, 1), y=c(0.05, 1))+
    # geom_abline(intercept=0.0, slope=1.1111111, color="darkgrey", size=0.75, alpha=0.5, linetype =3)+
    # geom_abline(intercept=-0.0, slope=0.9, color="darkgrey", size=0.75, alpha=0.5, linetype =3 )+
    ggtitle(bacteria_names[i])+
    labs(x="aerobic", y="anaerobic")+
    stat_poly_eq(b_0.constant =T)+
    scale_x_continuous( limits = c(-1, 1), labels = label_number(accuracy = 0.01)) +
    scale_y_continuous( limits = c(-1, 1), labels = label_number(accuracy = 0.01))
  show(plot)
  
  x<-paste0(bacteria[i], ".jpg")
  
  ggsave(filename = x,   path = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/output/correlations_O2", 
         width = 18,
         height = 15,
         units = c("cm"))
  }




```

# 6. assign taxonomy 

```{r}

#########################################
#Assign phyla to species name
########################################
bacteria_pyla<-as.data.frame(unique(means_corr[,1]))

bacteria_pyla$class<-0
bacteria_pyla<-as.data.frame(bacteria_pyla)


names(bacteria_pyla)<-c("Bacteria", "class")

###
bacteroidetes<-grep("Bac", bacteria_pyla$Bacteria)
bacteria_pyla[bacteroidetes, c(2)] ="Bacteroidia"

pre<-grep("Prevo", bacteria_pyla$Bacteria)
bacteria_pyla[pre,2]="Bacteroidia"

parabac<-grep("Parabac", bacteria_pyla$Bacteria)

bacteria_pyla[parabac,2]="Bacteroidia"

blaut<-grep("Blaut", bacteria_pyla$Bacteria)
bacteria_pyla[blaut,2]="Clostridia"

rumi<-grep("Rumi", bacteria_pyla$Bacteria)
bacteria_pyla[rumi,2]="Clostridia"

rose<-grep("Rose", bacteria_pyla$Bacteria)
bacteria_pyla[rose,2]="Clostridia"

clo<-grep("Clo", bacteria_pyla$Bacteria)
bacteria_pyla[clo,2]="Clostridia"

Faec<-grep("Faecalibacterium", bacteria_pyla$Bacteria)

bacteria_pyla[Faec,2]="Clostridia"

flavi<-grep("Flavi", bacteria_pyla$Bacteria)



bacteria_pyla[flavi,2]="Clostridia"

copr<-grep("Copro", bacteria_pyla$Bacteria)
bacteria_pyla[copr,2]="Clostridia"

eu<-grep("Eu", bacteria_pyla$Bacteria)
bacteria_pyla[eu,2]="Clostridia"

flavi<-grep("Flavo", bacteria_pyla$Bacteria)
bacteria_pyla[flavi,2]="Clostridia"

lacto<-grep("Lactobacillus", bacteria_pyla$Bacteria)
bacteria_pyla[lacto,2]="Bacilli"

subdo<-grep("Sub", bacteria_pyla$Bacteria)
bacteria_pyla[subdo,2]="Clostridia"

bif<-grep("Bif", bacteria_pyla$Bacteria)
bacteria_pyla[bif,2]="Actinomycetia"

colin<-grep("Collin", bacteria_pyla$Bacteria)
bacteria_pyla[colin,2]="Coriobacteriia"



anaero<-grep("Anaero", bacteria_pyla$Bacteria)
bacteria_pyla[anaero,2]="Clostridia"

anaro<-grep("Anaro", bacteria_pyla$Bacteria)
bacteria_pyla[anaro,2]="Clostridia"

marvin<-grep("Marvin", bacteria_pyla$Bacteria)
bacteria_pyla[marvin,2]="Clostridia"

phasc<-grep("Phasco", bacteria_pyla$Bacteria)
bacteria_pyla[phasc,2]="Negativicutes"

entero<-grep("Entero", bacteria_pyla$Bacteria)
bacteria_pyla[entero,2]="Bacilli"

esch<-grep("Esch", bacteria_pyla$Bacteria)
bacteria_pyla[esch,2]="Gammaproteobacteria"
#########################

names(bacteria_pyla)<-c("Bacteria", "Class")
```


```{r}
ratio_O2<-means_stats_shapiro

min(ratio_O2$mean_aerobe)

ratio_O2<-join(ratio_O2, bacteria_pyla, by= "Bacteria")

# ratio_O2$mean_anaerobe[ratio_O2$mean_anaerobe<0.001]<-0
# ratio_O2$mean_aerobe[ratio_O2$mean_aerobe<0.001]<-0

ratio_O2%>%
  subset(Bacteria != "Ruminococcus_bromii")->ratio_O2


ratio_O2$ratio <-  ratio_O2$mean_aerobe - ratio_O2$mean_anaerobe 

ratio_O2$agar_condition<-as.character(ratio_O2$agar_condition)


ggplot(subset(ratio_O2, agar_condition %in% "minimal agar (''MIC'')"), aes(x=metabolite, y=Bacteria, fill=ratio))+
  geom_tile()+
  scale_fill_gradientn(colors = c("coral", "coral", "white", "cyan4"),  values = rescale(c(-3, -0.5,0,0.5)))+
  facet_grid(rows=vars(Class), space="free", scale="free")+
    theme(strip.text.y = element_text(angle=0, size = 15),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, size=15, vjust=0.2),
        axis.text.y = element_text(size=15 , face="italic"))+
  labs(fill="Δrel. metabolites")



 ggsave(filename = "O2_heatplot.jpeg",   
         path = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/output", 
         width = 22,
         height = 20,
         units = c("cm"))

```



```{r}
ratio_O2%>%
  subset(Bacteria %in% c("Anaerostipes_caccae", 
                         "Anaerostipes_hadrus", 
                         "Anaerotignum_lactatifermentans",
                         "Bacteroides_caccae",
                         "Bacteroides_thetaiotaomicron" ,
                         "Bacteroides_uniformis" ,
                         "Bacteroides_vulgatus" , 
                         "Bifidobacterium_adolescentis"  , 
                         "Bifidobacterium_longum" ,
                         "Blautia_hansenii" ,
                         "Blautia_hydrogenotrophica",
                         "Blautia_obeum"  , "Blautia_producta" , 
                         "Coprococcus_sp" ,
                         "Enterococcus_faecalis"  ,
                         "Escherichia_coli" , 
                         "Eubacterium_limosum" ,
                         "Faecalibacterium_prausnitzii" ,
                         "Flavonifractor_sp", 
                         "Lactobacillus_plantarum" , 
                         "Parabacterium_merdae"  ,
                         "Phascolarctobacterium_faecium", 
                         "Roseburia_intestinalis"   ,
                         "Ruminococcus_gnavus" )) %>%
           subset(agar_condition %in% c("minimal agar (''MIC'')"))->sub_nic


ratio_O2%>% 
  subset(Bacteria %in% c("Bacteroides_faecis", "Bacteroides_ovatus", "Eubacterium_eligens", "Prevotella_copri", "Bacteroides_xylanisolvens", "Bacteroides_fragilis", "Clostridium_sporogenes", "Collinsella_aerofaciens"  )) %>%
           subset(agar_condition %in% c("agar with highest growth"))->sub_nic_1


selection<-rbind(sub_nic, sub_nic_1)


selection$p.signif[selection$p.signif=="ns"]<-NA
selection$p_corrected[selection$p_corrected=="ns"]<-NA



```

```{r}

ggplot(selection, aes(x=metabolite, y=Bacteria, fill=ratio))+
  geom_tile()+
  scale_fill_gradientn(colors = c("coral", "coral", "white", "cyan4"),  values = rescale(c(-0.6, -0.5,0,0.5)))+
  facet_grid(rows=vars(Class), space="free", scale="free")+
    theme(strip.text.y = element_text(angle=0, size = 15),
        axis.title = element_blank(),
        axis.text.x = element_text(angle = 90, size=15, vjust=0.2),
        axis.text.y = element_text(size=15 , face="italic"))+
  labs(fill="Δrel. metabolites")+
  ggtitle(expression(Metabolic~changes~as~a~response~to~O[2]))+
  geom_text(aes(label=p_corrected))

ggsave(filename = "O2_heatplot_selection.jpeg",   
         path = "C:/Users/zuendj/Desktop/03_data R/oxidative stress/output", 
         width = 22,
         height = 20,
         units = c("cm"))



selection %>%
  subset(Bacteria %in% c("Escherichia_coli"))

```


```{r}
HPLC_values[, c("assay","condition", "total_SCFA")]%>%
  ggplot(aes(y=assay, x=total_SCFA))+
  geom_boxplot(aes(colour=condition) )+
  geom_point(aes(colour=condition))



HPLC_values[, c("assay","condition", "acetate")]%>%
  ggplot(aes(y=assay, x=acetate))+
  geom_boxplot(aes(colour=condition) )+
  geom_point(aes(colour=condition))



HPLC_values[, c("assay","condition", "propionate")]%>%
  ggplot(aes(y=assay, x=propionate))+
  geom_boxplot(aes(colour=condition) )+
  geom_point(aes(colour=condition))

HPLC_values[, c("assay","condition", "butyrate")]%>%
  ggplot(aes(y=assay, x=butyrate))+
  geom_boxplot(aes(colour=condition) )+
  geom_point(aes(colour=condition))

HPLC_values[, c("assay","condition", "succinate")]%>%
  ggplot(aes(y=assay, x=succinate))+
  geom_boxplot(aes(colour=condition) )+
  geom_point(aes(colour=condition))


HPLC_values[, c("assay","condition", "lactate")]%>%
  ggplot(aes(y=assay, x=lactate))+
  geom_boxplot(aes(colour=condition) )+
  geom_point(aes(colour=condition))


HPLC_values[, c("assay","condition", "formate")]%>%
  ggplot(aes(y=assay, x=Formate))+
  geom_boxplot(aes(colour=condition) )+
  geom_point(aes(colour=condition))

```




