---
title: "Alpha and Beta diveristy under Ox. stress"
---

```{r setup, echo =FALSE}
.libPaths('C:\\Users\\zuendj\\Documents\\R4Libs')
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "P:/Shared_documents/Serafina Plüss/20230320_sequencing_results/Janina", cho = TRUE, warning = FALSE, message = FALSE)
# install.packages('devtools')
install.packages(
  "microViz",
  repos = c(davidbarnett = "https://david-barnett.r-universe.dev", getOption("repos"))
)
devtools::install_github("tidyverse/tidyverse") 

# library(tidyverse)
# library(openxlsx)
# library(rstatix)
# library(microViz)
# library(ggpubr)
# library(plyr)
# library(metagMisc)
# library(picante)
# library(phyloseq)
# library(ggh4x)
# library(cowplot)

```

```{r, echo =FALSE}
# source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/phyloseq_taxa_tests.R") 
# source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/phyloseq_normalisation.R") 
# source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/phyloseq_alpha.R") 
# source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/phyloseq_beta.R") 
# source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/phyloseq_heatmap.R")
# source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/phyloseq_taxa_tests.R")
# source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/phyloseq_varia.R")
# source("C:/Users/zuendj/Desktop/03_data R/divcom_functions/R/phyloseq_beta.R")
#  source("https://raw.githubusercontent.com/fconstancias/metabarcodingRpipeline/dev/scripts/functions_export_simplified.R")

#get metadata file: load this function
sample.data.frame <- function(ps) {
  return(as(phyloseq::sample_data(ps), "data.frame"))}

facet_labeller <- function(variable,value){
  return(names[value])}
```

##create manifest
```{r}
# meta<-left_join(read.xlsx("P:/Shared_documents/Janina_Zuend/08_phd/01_sequencing_data/Ox_stress/meta.xlsx"), hplc[,c(2:12)])
# 
# names<-setNames(as.data.frame(list.files("P:/Shared_documents/Janina_Zuend/08_phd/01_sequencing_data/Ox_stress/raw/230630_M06272_0140_000000000-L3NKW_v2")), c("filename"))
# 
# setNames(as.data.frame(str_split_fixed(names$filename, pattern="_", 4)), c("sample_name", "x", "y", "z"))%>%
#   cbind(., names)%>%
#   pivot_wider(values_from = "filename", names_from = "z")->sample_names
# 
# write.xlsx(left_join(meta, setNames(sample_names[,c("sample_name", "R1_001.fastq.gz", "R2_001.fastq.gz")],c("sample_name", "forward-absolute-filepath", "reverse-absolute-filepath" ))),
# "P:/Shared_documents/Janina_Zuend/08_phd/01_sequencing_data/Ox_stress/manifest.xlsx")
```

```{r}
# tax<-read_tsv("P:/Shared_documents/Janina_Zuend/08_phd/01_sequencing_data/Ox_stress/taxonomy.tsv")
# 
# biom<-import_biom("P:/Shared_documents/Janina_Zuend/08_phd/01_sequencing_data/Ox_stress/feature-table.biom")
```



## 1. Get the phyloseq object 
```{r}
phyloseq <-readRDS("P:/Shared_documents/Janina_Zuend/08_phd/01_sequencing_data/Ox_stress/phyloseq_phylo.RDS")%>%
  phyloseq_get_strains()%>%
  subset_samples(!donor_name %in% "D1" )
```
#Update the metadata
```{r}
#get the hplc data
hplc<-read.xlsx("C:/Users/zuendj/Desktop/03_data R/wp2_complex/input/20230613_hplc_list_results_stress.xlsx")%>%
  subset(!sample_name %in% NA)

#get the sample list and join with hplc
meta<-left_join(read.xlsx("P:/Shared_documents/Janina_Zuend/08_phd/01_sequencing_data/Ox_stress/meta.xlsx"), hplc[,c(2:12)])

meta %>%
  mutate(donor_name = ifelse(donor_name == "D1", "D1_exclude", donor_name))%>%
  mutate(donor_name = ifelse(donor_name == "D8", "D1", donor_name))->meta

#export
write.xlsx(meta[,-c(13)], "P:/Shared_documents/Janina_Zuend/08_phd/01_sequencing_data/Ox_stress/meta_new.xlsx")

#load the new metadata file to the phyloseq
phyloseq%>%
  physeq_add_metadata(physeq = .,
                      metadata = "P:/Shared_documents/Janina_Zuend/08_phd/01_sequencing_data/Ox_stress/meta_new.xlsx" %>%
                        readxl::read_xlsx(),
                      sample_column = "sample_name") -> phyloseq

#store meta data in df to acess & annalyze hplc and OD
sample.data.frame(sample_data(phyloseq))->meta
```

#Use tax fix to correct for empty values in tax table
###Filtering
```{r}
ps_strain_filt <- phyloseq
# threshold in %
threshold = 0.1

# filter per sample to remove all ASV lower in abundance than threshold (%)
otu_table(ps_strain_filt) <- otu_table(ps_strain_filt) %>%
  as.data.frame() %>%
  dplyr::mutate(across(everything(), ~ ifelse(. <= sum(.) * (threshold/100), 0, .))) %>% otu_table(., taxa_are_rows = TRUE)

# remove all ASVs that are not present in any sample
ps_strain_filt <- ps_strain_filt %>%
  filter_taxa(function(x) sum(x > 0) > 0, TRUE)

noNA = !is.na(tax_table(ps_strain_filt)[,"Genus"]) & !is.na(tax_table(ps_strain_filt)[,"Species"])
tax_table(ps_strain_filt)[noNA][,"Species"] = paste(tax_table(ps_strain_filt)[noNA][,"Genus"], tax_table(ps_strain_filt)[noNA][,"Species"])
ps_strain_filt%>%
  tax_fix()->ps_strain_filt
```

```{r}
ps_strain_filt %>%
  phyloseq_rarefaction_curves(stepsize = 500, 
                              color_data = "time", 
                              facet_data = NULL) -> p 


p + geom_vline(xintercept = 3121,
               color = "red",
               linetype = "dashed", size=0.25)+
  facet_grid(rows=vars(donor_name)) -> plot

plot
```


```{r}
ps_strain_filt%>% 
  phyloseq_check_lib_size(data_color = "stress",
                          data_facet = "time",
                          nreads_display = 5543,
                          first_n = nsamples(ps_strain_filt)) -> lib



min_lib = c(3122)

ps_strain_filt %>%
  rarefy_even_depth(rngseed = 123,
                    sample.size = min_lib) -> phyloseq_rare
```


#Export phyloseq for subsequent analysis 
```{r}
# saveRDS(ps_strain_filt, file= "C:/Users/zuendj/Desktop/03_data R/wp2_complex/input/phyloseq_nonrare.RDS")
# saveRDS(phyloseq_rare, file= "C:/Users/zuendj/Desktop/03_data R/wp2_complex/input/phyloseq_rare.RDS")
```

#Export ref seq table
```{r}
library(seqinr)
write.fasta(names= rownames(as.data.frame(refseq(phyloseq_rare))), sequences=as.list(as.data.frame(refseq(phyloseq_rare))$x), file.out = "P:/Shared_documents/Janina_Zuend/08_phd/12_working_package_2/08_blast/ox_stress.fasta")
```


## Alpha diversity analysis
# passages of controls 
```{r}
names<-list("diversity_shannon"= "Shannon", 
            "observed"= "Observed")

phyloseq_rare%>%
  phyloseq_alphas(phylo = TRUE)%>%
  subset(!time %in% "post_stress_2")%>%
  subset(time %in% "pre-stress" | 
           ((stress_level %in% c("control") & (stress %in% c("H2O2")) | 
               (stress_level %in% "max" & stress %in% "O2" & incubation %in% "anaerobe"))) |
                    Sample_id_compete %in% "feces"  #with or without feces
                 ) %>%
  mutate(time = factor(time, levels=c("feces","pre-stress", "stress", "post_stress_1")))-> sub


left_join(left_join(setNames(aggregate(sub$observed, by = list(sub$donor_name, sub$time), FUN="mean"), c("donor_name", "time", "mean")), 
setNames(aggregate(sub$observed, by = list(sub$donor_name, sub$time), FUN="min"), c("donor_name", "time", "min"))),
setNames(aggregate(sub$observed, by = list(sub$donor_name, sub$time), FUN="max"), c("donor_name", "time", "max")))%>%
  ggplot(aes(x=time, y=mean))+
  theme(text = element_text(size  = 15),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle=90, hjust=1, vjust=0.5),
        legend.text = element_text(vjust=1, hjust=0),
        strip.background = element_rect(fill="white"),
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), 
        panel.spacing = unit(0.31, "cm"),
        legend.text.align = 0)+
  geom_line(aes(group = donor_name))+
  geom_point(aes(group = donor_name, y=max), alpha=0.6)+
  geom_point(aes(group = donor_name, y=min), alpha=0.6)+
  ylab("ASVs")+
  # labs(colour="Donor")+
  scale_x_discrete(labels = c("Feces","Pre-culture", "Stress", "Post-stress"))+
  facet_grid(cols=vars(donor_name))+
   scale_y_continuous(breaks=c(70, 100, 130)) ->alpha_controls
# 
# aggregate(sub$value, by=list(sub$index), FUN=mean)
# aggregate(sub$value, by=list(sub$index), FUN=sd)


ggsave(filename = "alpha_passages_control.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final",
       width = 15,
       height = 5,
       units = c("cm"))

left_join(setNames(aggregate(sub$observed, by=list( sub$time), FUN="mean"), c( "time", "mean")),
setNames(aggregate(sub$observed, by=list( sub$time), FUN="sd"), c("time", "sd")))%>%
   mutate(mean = round(mean, digits=0))%>%
  mutate(sd = round(sd, digits=0))
```

##beta dist feces

```{r}


phyloseq_rare %>%
    subset_samples(incubation %in% c("feces", "anaerobe"))%>%
  subset_samples(!time %in% c("post_stress_2"))%>%
  subset_samples(time %in% "pre-stress" | time %in% "feces" |
           ((stress_level %in% c("control") & (stress %in% c("H2O2")) | 
               (stress_level %in% "max" & stress %in% "O2" & incubation %in% "anaerobe"))))%>%
  
  subset_samples(!Sample_id_compete %in% c("feces_dil_1"))%>%
  subset_samples(!donor_name %in% NA)-> sub

sub%>%
  phyloseq::distance(method = "bray") -> bc

sub %>%
  ordinate(method = "PCoA",
           distance = bc) -> ord

sub %>% 
  phyloseq_distance_boxplot(p = ., 
                            dist = bc, 
                            d = "time") -> out # column in metadata for comparaisons

setNames(as.data.frame(out$matrix),c("sample_name", "sample_name_comp", "value", "condition_comp", "comparison")) -> distances

distances%>%
  subset(comparison %in% c("feces", "pre-stress", "stress"))-> distances

joined<-left_join(distances, sample.data.frame(sub)[,c(1,8:13)])

left_join(joined, setNames(meta[, c("sample_name", "donor_name", "stress")], c("sample_name_comp", "donor_name_comp", "stress_comp")))->joined

for (i in 1:nrow(joined)){
  if (joined[i,c("donor_name")] == joined[i,c("donor_name_comp")]       )     {joined[i,c("match")]= "true"}
}


joined%>%
   subset(! (stress %in% "O2" & stress_comp %in% "H2O2"))%>%
  subset(! (stress %in% "H2O2" & stress_comp %in% "O2"))%>%
  subset(match %in% "true")->match
```
#Taxa retained analysis needed?
## Analysis shared taxa
## number of ASV culture vs. feces
```{r}


donor_count<-c(unique(sample.data.frame(sub)$donor_name))
maintain<-as.data.frame(matrix(5,0,0))
loss <- as.data.frame(matrix(3,0,0))
gain <- as.data.frame(matrix(3,0,0))

for (j in 1:length(donor_count)){
  
sub %>% 
    subset_samples(donor_name %in% c(donor_count[j]))%>%
     tax_glom(taxrank = "Genus")%>%
    transform_sample_counts(function(x) {x/sum(x)}) %>%
    filter_taxa(., function(x){mean(x) > 0}, TRUE)%>%
    psmelt()->df_abund
  


condition_count <- c("feces" ,  "pre-stress" ,  "stress"  ,    "post_stress_1")

abundance = 0.01
n=nrow(maintain)+1

for(i in 1:length(condition_count)){

  
df_abund%>%
  subset(time %in% c(condition_count[i]))%>%
  subset(Abundance > 0.00001)-> culture

df_abund%>%
  subset(time %in% c("feces"))%>%
  subset(Abundance > 0.00001)-> feces

rbind(culture, feces)-> all

otu_all<-setNames(as.data.frame(unique(all$Genus)), c("Genus"))




diff_feces<-setNames(as.data.frame(setdiff(otu_all$Genus, culture$Genus)), c("Genus"))
diff_culture<-setNames(as.data.frame(setdiff(otu_all$Genus, feces$Genus)), c("Genus"))

intersec<-setNames(as.data.frame(intersect(culture$Genus, feces$Genus)), c("Genus"))

maintain[n,"donor_name"] = donor_count[j]
maintain[n,"condition"] = condition_count[i]

maintain[n,c("Shared")]<-nrow(intersec)/(nrow(diff_culture)+nrow(diff_feces)+nrow(intersec)) 
maintain[n,c("Cultures")]<-nrow(diff_culture)/(nrow(diff_culture)+nrow(diff_feces)+nrow(intersec)) 
maintain[n,c("Feces")]<-nrow(diff_feces)/(nrow(diff_culture)+nrow(diff_feces)+nrow(intersec)) 
maintain[n,c("Shared_number")]<-nrow(intersec)
maintain[n,c("Cultures_number")]<-nrow(diff_culture) 

try(lol<-data.frame(Feces = diff_feces,
           donor_name = rep(donor_count[j], length(diff_feces)),
           time = rep(condition_count[i], length(diff_feces))), silent = T)
# print(head(lol))

loss<-rbind(loss, lol)

try(lol<-data.frame(Feces = diff_culture,
           donor_name = rep(donor_count[j], length(diff_feces)),
           time = rep(condition_count[i], length(diff_feces))), silent = T)
# print(head(lol))

gain<-rbind(gain, lol)

n=n+1
  }}

```

```{r}
maintain%>%
  mutate(condition = factor (condition , levels = c("feces", "pre-stress", "stress", "post_stress_1")))%>%
  gather(sample, value, "Shared":"Feces")%>%
  ggplot(aes(x=condition, y=value))+
   geom_bar(stat = "identity", width=0.9143, colour="black", size=0.1, aes(fill=sample))+
   scale_fill_manual(values=c("grey","burlywood4", "cyan4"))+
  facet_grid(cols=vars(donor_name))

maintain%>%
   mutate(condition = factor (condition , levels = c("feces", "pre-stress", "stress", "post_stress_1")))%>%
  gather(sample, value, "Shared_number":"Cultures_number")%>%
   group_by( condition, donor_name, sample) %>%
  dplyr::summarize(value = sum(value)) %>%
  ggplot(aes(x=condition, y=value, fill = sample, stratum = sample, alluvium = sample, label = sample))+
  facet_grid(cols=vars(donor_name))+
  geom_flow(stat = "alluvium", lode.guidance = "frontback", reverse = FALSE) +
  geom_stratum(reverse = FALSE, width = 0.7, color = "transparent") +
  # geom_bar(stat = "identity", width=0.9143, colour="black", size=0.1, aes(fill=sample))+
   scale_fill_manual(values=c("grey","burlywood4", "cyan4"), labels=c("Culture", "Feces"), name="")+
   scale_x_discrete(labels = c("Feces","Pre-culture", "Stress", "Post-stress"))+
 theme(text = element_text(size  = 15),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle=90, hjust=1, vjust=0.5),
        legend.text = element_text(vjust=1, hjust=0),
        strip.background = element_rect(fill="white"),
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), 
        panel.spacing = unit(0.31, "cm"),
        legend.text.align = 0)+
  ylab("Number of Genera")->shared
shared

mean(C$Shared)

left_join(setNames(aggregate(maintain$Shared, by = list(c(maintain$condition)), FUN="mean"), c("condition", "mean"))%>%
  mutate(mean = round(mean, digits=2)), 
  setNames(aggregate(maintain$Shared, by = list(c(maintain$condition)), FUN="sd"), c("condition", "sd"))%>%
  mutate(sd= round(sd, digits=2)) )
```

```{r}


sub %>% 
    tax_glom(taxrank = "Genus")%>%
    transform_sample_counts(function(x) {x/sum(x)}) %>%
    filter_taxa(., function(x){mean(x) > 0}, TRUE)%>%
    psmelt() %>%
  subset(time %in% "feces")->df_abund


rbind(left_join(loss%>%
            mutate(gainorloss = "loss"), df_abund[, c("donor_name", "Genus",  "Abundance")]),
          left_join(gain%>%
            mutate(gainorloss = "gain"), df_abund[, c("donor_name", "Genus",  "Abundance")]))->hoi
hoi%>%
  mutate(Genus= factor(Genus))%>%
 subset(Abundance > 0.01)%>% 
  mutate(Genus = reorder(Genus, Abundance))%>%
  ggplot(aes(y=Genus, x=time, fill = Abundance))+
  geom_tile()+
  facet_grid(cols=vars(donor_name), rows= vars(gainorloss), scales="free")

unique(ab$Genus)
```

##Beta dist. over all passages (over all controls)
```{r}
match%>%
  subset(incubation %in% c("feces", "anaerobe"))%>%
  subset(!time %in% c("post_stress_2"))%>%
  subset(time %in% "pre-stress" | 
           ((stress_level %in% c("control") & (stress %in% c("H2O2")) | 
               (stress_level %in% "max" & stress %in% "O2" & incubation %in% "anaerobe"))))%>%
  subset((comparison %in% "feces" & time %in% c( "pre-stress","stress", "post_stress_1"))|
           (comparison %in% "pre-stress" & time %in% c( "stress", "post_stress_1"))|
           (comparison %in% "stress" & time %in% c( "post_stress_1")))%>%
   mutate(time = factor(time, levels=c("pre-stress","stress", "post_stress_1")))->sub



left_join(left_join(setNames(aggregate(sub$value, by = list(sub$donor_name, sub$time, sub$comparison), FUN="mean"), c("donor_name", "time", "comparison", "mean")), 
setNames(aggregate(sub$value, by = list(sub$donor_name, sub$time, sub$comparison), FUN="min"), c("donor_name", "time","comparison", "min"))),
setNames(aggregate(sub$value, by = list(sub$donor_name, sub$time, sub$comparison), FUN="max"), c("donor_name", "time", "comparison","max")))%>%
  
rbind(., data.frame(donor_name = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
                     time = rep("feces", 7),
                     mean = rep(0, 7),
                     min = rep(0, 7),
                     max = rep(0, 7),
                    comparison ="feces"
                     ))%>%
  rbind(., data.frame(donor_name = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
                     time = rep("pre-stress", 7),
                     mean = rep(0, 7),
                     min = rep(0, 7),
                     max = rep(0, 7),
                    comparison ="pre-stress"
                     ))%>%
    rbind(., data.frame(donor_name = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
                     time = rep("stress", 7),
                     mean = rep(0, 7),
                     min = rep(0, 7),
                     max = rep(0, 7),
                    comparison ="stress"
                     ))%>%
  mutate(time = factor(time, levels=c("feces","pre-stress","stress", "post_stress_1")))->beta_plots

beta_plots%>%
  ggplot(aes(x=time, y=mean))+
  geom_hline(yintercept=0.25, alpha=0.2)+
   geom_hline(yintercept=0.5, alpha=0.2)+
   geom_hline(yintercept=0.75, alpha=0.2)+
  theme(text = element_text(size  = 15),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle=90, hjust=1, vjust=0.5),
        legend.text = element_text(vjust=1, hjust=0),
        strip.background = element_rect(fill="white"),
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), 
        panel.spacing = unit(0.31, "cm"),
        legend.text.align = 0)+
  geom_line(aes(group = comparison, colour=comparison))+
  geom_point(aes(colour = comparison, y=max), alpha=0.6)+
  geom_point(aes(colour = comparison, y=min), alpha=0.6)+
  ylab("Bray")+
  # labs(colour="Donor")+
  scale_x_discrete(labels = c("Feces","Pre-culture", "Stress", "Post-stress"))+
  facet_grid(cols=vars(donor_name))+
   scale_y_continuous(breaks=c(0, 0.5 ,0.75))->beta_controls

beta_controls


ggsave(filename = paste("beta_controls.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final",
       width = 7.5,
       height = 7.5,
       units = c("cm"))
sub%>%
  subset(time %in% "post_stress_1" & comparison %in% "stress")

```
```{r}
beta_plots%>%
  gather(beta_dist, value_all, "min":"max")%>%
  ggplot(aes(x=time, y=value_all))+
  theme(text = element_text(size  = 15),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle=90, hjust=1, vjust=0.5),
        legend.text = element_text(vjust=1, hjust=0),
        strip.background = element_rect(fill="white"),
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), 
        panel.spacing = unit(0.31, "cm"),
        legend.text.align = 0)+
  geom_boxplot(aes(colour = comparison))+
  geom_point(aes(colour = comparison), alpha=0.6, position=position_dodge(0.9))+
  ylab("Bray")+
  # labs(colour="Donor")+
  scale_x_discrete(labels = c("Feces","Pre-culture", "Stress", "Post-stress"))+
   scale_y_continuous(breaks=c(0, 0.5 ,0.75))


beta_plots%>%
  gather(beta_dist, value_all, "min":"max")->long

left_join(setNames(aggregate(long$value_all, by=list(long$comparison, long$time), FUN="mean"), c("comparison", "time", "mean")),
setNames(aggregate(long$value_all, by=list(long$comparison, long$time), FUN="sd"), c("comparison", "time", "sd")))%>%
   mutate(mean = round(mean, digits=2))%>%
  mutate(sd = round(sd, digits=2))
```

```{r}
plot_grid(alpha_controls+theme(axis.text.x =  element_blank(),
                               axis.ticks.x =  element_blank()),
          shared+theme(axis.text.x =  element_blank(),
                               axis.ticks.x =  element_blank(),
                       strip.text = element_blank(), 
                       legend.position = "none"),
          beta_controls+theme(axis.text.x =  element_blank(),
                               axis.ticks.x =  element_blank(), 
                              strip.text = element_blank(), 
                              legend.position = "none")+
            scale_y_continuous(breaks=c(0, 0.5, 1), limits=c(0,1)), 
          controls + theme(strip.text = element_blank(), 
                           axis.title.x = element_blank()),
          nrow=4, 
        align= "v"
        # , 
        # rel_heights = c(0.165, 0.105, 0.68)
          )


ggsave(filename = paste( "controls.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final",
       width = 24,
       height = 40,
       units = c("cm"))
```


# PCOa - controls
```{r}
# library(plyr)


phyloseq_rare %>%
   subset_samples( time %in% c("pre-stress", "feces") | (stress_level %in% c("control") & stress %in% c("H2O2", "O2") & incubation %in% "anaerobe"))%>%
  subset_samples(!time %in% "post_stress_2")-> sub


  dist = "bray"
  ord_meths = c("PCoA")


  plist = llply(as.list(ord_meths), function(i, sub, dist){
          ordi = ordinate(sub, method=i, distance=dist)
          plot_ordination(physeq = sub, 
                          ordination = ordi,
                          color = "donor_name", 
                          shape = "stress",
                          title = "PCoA Bray-Curtis", 
                          label= "time")},
          sub, 
          dist)


  names(plist) <- ord_meths

  pdataframe = ldply(plist, function(x){
      df = x$data[, 1:2]
      colnames(df) = c("Axis_1", "Axis_2")
      return(cbind(df, x$data))})


  
  names(pdataframe)[1] = "method"


  pdataframe%>%  mutate(time = factor(time, levels = c("feces","pre-stress", "stress", "post_stress_1")))%>%
    mutate(time = factor(time, levels=c("feces","pre-stress", "stress", "post_stress_1")))%>%
      ggplot(aes(Axis_1, Axis_2))+ 
      # scale_color_manual(values=c('darkgoldenrod1', "cyan4"))+
    stat_ellipse(aes(group=donor_name, fill=donor_name), geom = "polygon", alpha=0.1)+
    stat_ellipse(aes(group=donor_name, colour=donor_name))+
      theme(text= element_text(size = 15), 
          legend.text.align = 0, 
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))  +
    scale_shape_manual(values=c(  21, 15, 16, 17), labels=c( "Feces", "Pre-culture", "P1", "P2"), name="Passage")+
    # geom_line(aes(group=donor_name, colour= donor_name))+
    labs(colour="Donor")+
    xlab("PCoA1 (20.5%)")+
    ylab("PCoA2 (16.2%)")+
      geom_point(aes(shape=time, colour=donor_name), stroke=1.5, alpha=0.8)+
     scale_fill_manual(values=c("lightblue", "steelblue", "yellowgreen", "forestgreen", "lightpink1", "red3", "sienna1", "orange"), guide="none")+
    scale_colour_manual(values=c("lightblue", "steelblue", "yellowgreen", "forestgreen", "lightpink1", "red3", "sienna1", "orange"))+
    force_panelsizes(rows = unit(8, "cm"), col = unit(8, "cm"))
  
  ggsave(filename =  "Beta_div.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final")
```



##Agar
##beta dist 

```{r}
phyloseq_rare %>%
 subset_samples(time %in% "stress" & 
                  stress %in% "O2" &
                  incubation %in% "anaerobe")-> sub

sub%>%
  phyloseq::distance(method = "bray") -> bc

sub %>%
  ordinate(method = "PCoA",
           distance = bc) -> ord

sub %>% 
  phyloseq_distance_boxplot(p = ., 
                            dist = bc, 
                            d = "stress_level") -> out # column in metadata for comparaisons

setNames(as.data.frame(out$matrix),c("sample_name", "sample_name_comp", "value", "condition_comp", "comparison")) -> distances


distances%>%
  subset(comparison %in% c("max"))-> distances

joined<-left_join(distances, sample.data.frame(sub)[,c(1,8:13)])

joined%>%
  mutate(sample_name_comp = recode(sample_name_comp, 
                                   "M-126" = 'D2',
                                   "M-127" = 'D3',
                                   "M-128" = 'D4',
                                   "M-129" = 'D5',
                                   "M-130" = 'D6',
                                   "M-131" = 'D7',
                                   "M-132" = 'D1',))->joined




for (i in 1:nrow(joined)){
  if (joined[i,c("donor_name")] == joined[i,c("sample_name_comp")]){joined[i,c("match")]= "true"}
}

joined%>%
  subset(match %in% "true")->match
```

##Beta dist. over all passages (over all controls) - agar
```{r}
match%>%
  mutate(stress_level = factor(stress_level, levels =c("high","median",  "low", "control")))%>%
  wilcox_test(value~condition_comp)%>%
  add_significance()%>%
  add_y_position()->stat
  
match%>%
  mutate(stress_level = factor(stress_level, levels =c("high","median",  "low", "control")))%>%
  ggplot(aes(x=stress_level, y=value))+
  theme(text = element_text(size  = 15),
        axis.text.x = element_text(angle=90, hjust=1, vjust=0.5),
        legend.text = element_text(vjust=1, hjust=0),
        strip.background = element_rect(fill="white"),
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), 
        panel.spacing = unit(0.31, "cm"),
        legend.text.align = 0)+
  geom_boxplot()+
  geom_point(aes(colour=donor_name) , alpha=0.6)+
  ylab("Bray-Curtis")+
  xlab("Agar [%]")+
  labs(colour="Donor")+
  stat_pvalue_manual(stat, label = "p.adj.signif", hide.ns = T, size=6)+
  scale_colour_manual(values=c("lightblue", "steelblue", "yellowgreen", "forestgreen", "lightpink1", "red3", "sienna1", "orange"))+
  scale_x_discrete(labels = c("control" = "0.12", "low"= "0.09", "median"="0.07", "high"="0.04"))+
    force_panelsizes(rows = unit(5, "cm"), col = unit(6, "cm"))+
  ylim(0, 0.6)


ggsave(filename = paste("agar_betadist.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final",
       width = 12,
       height = 7,
       units = c("cm"))

```



HO2 alpha diversity absolute
```{r}
# library(stats)
names<-list("diversity_shannon"= "Shannon", 
            "observed"= "Observed")

phyloseq_rare%>% 
  phyloseq_alphas(phylo = TRUE)%>%
  subset(stress %in% "H2O2" & time %in% "stress")%>%
  left_join(., pd(t(as.data.frame(otu_table(phyloseq_rare))), phy_tree(phyloseq_rare))%>%
              rownames_to_column("sample_name"))%>%
  mutate(stress_level = factor(stress_level, levels=c("control", "low", "median", "high")))%>%
  gather(index, value, "observed":"PD")-> sub

sub%>%
  subset(index %in% c("observed", "diversity_shannon"))%>%
  group_by(index)%>%
  wilcox_test(data= .,
  formula = value ~ stress_level, ref.group = "control" , p.adjust.method = "BH"
  )%>%
  add_significance()%>%
  subset(!p.adj.signif %in% "ns")%>%
  add_y_position()%>%
  mutate(y.position = ifelse(index == "diversity_shannon", 4, y.position ))%>%
  mutate(y.position = ifelse(group2 == "median", 4.5 , y.position ))%>%
  mutate(y.position = ifelse(index == "observed", 150 , y.position ))%>%
  mutate(y.position = ifelse(index == "PD", 15 , y.position ))->stat_test


sub %>%
  subset(index %in% c("observed", "diversity_shannon"))%>%
  ggplot(aes(x=stress_level, y=value))+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 90, vjust= 0.5, hjust=1),
        # strip.background = element_blank(),
        axis.ticks.y = element_blank(),
        plot.title = element_text(hjust=0.25),
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single"))+
  geom_jitter(aes(colour=donor_name),  width = 0.1, alpha=0.8)+
  ylab("Alpha diversity")+
  labs(colour="Donor", fill="Stress level")+
  facet_grid(rows=vars(index), scales= "free", labeller=facet_labeller) +
  stat_pvalue_manual(stat_test, label = "p.adj.signif", hide.ns = T)+
  scale_x_discrete(labels = c("0", "0.22", "0.71", "2.3"))+
  xlab(expression(H[2]*O[2]~`[mM]`))


# show(p + theme (legend.position = "none"))

ggsave(filename = paste( "alpha_h2o2.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final",
       width = 7.5,
       height = 7.5,
       units = c("cm"))


```

#Alpha compared to control ==>Distances
```{r}
phyloseq_rare%>% 
  phyloseq_alphas(phylo = TRUE)%>%
  subset(stress %in% "H2O2" )%>%
    left_join(., pd(t(as.data.frame(otu_table(phyloseq_rare))), phy_tree(phyloseq_rare))%>%
              rownames_to_column("sample_name"))%>%
  mutate(stress_level = factor(stress_level, levels=c("control", "low", "median", "high")))%>%
  gather(index, value, "observed":"PD")->sub

sub[, c("donor_name", "time", "stress_level", "index", "value")]%>%
  subset(!time %in% "pre-stress")%>%
  subset(index %in% c("observed", "evenness_pielou"))%>%
  pivot_wider(values_from = value, names_from = stress_level)%>%
  mutate(diff_low = low - control)%>%
  mutate(diff_median = median - control)%>%
  mutate(diff_high = high - control)%>%
  gather(stress_level, value, "diff_low":"diff_high")%>%
  mutate(stress_level = factor(stress_level, levels=c( "diff_low", "diff_median", "diff_high")),
         time = factor(time, levels=c( "stress", "post_stress_1"))) %>%
  ggplot(aes(x=stress_level, y=value))+
  theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=20, angle = 90, vjust= 0.5, hjust=1),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_boxplot( aes(fill=time),outlier.shape = NA)+
  geom_jitter(aes(colour=donor_name, group=time), position=position_jitterdodge(0.1))+
  ylab("Diversity")+
  labs(colour="Donor", fill="Stress level")+
  facet_grid( rows=vars(index), scales= "free") +
  geom_hline(yintercept=0, linetype=2)+
  # ylim(-1, 0.5)+
  scale_x_discrete(labels=c( "low", "median", "high"))



ggsave(filename = paste( "alpha_o2_delta.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/O2/alpha",
       width = 25,
       height = 10,
       units = c("cm"))
  

```

#Alpha: upon oxygen stress
```{r}
names<-list("diversity_shannon"= "Shannon", 
            "observed"= "Observed")
#O2: aerobic
phyloseq_rare%>% 
  subset_samples( stress %in% "control" | (stress %in% "O2" ))%>%
  phyloseq_alphas(phylo = TRUE) %>%
  subset(time %in% "stress")%>%
  gather(index, value, "observed":"diversity_shannon")-> sub

###results are all non-significant
sub%>%
  group_by(index, stress_level)%>%
  t_test(data= .,
  formula = value ~ incubation)%>%
  add_significance()%>%
  mutate(stress_level = factor(stress_level, levels=c("control", "low", "median", "high", "max"))) %>%
  add_xy_position()%>%
  mutate(y.position = 4, xmin = 4.75, xmax=5.25)->stat_test



left_join( setNames(aggregate(sub$value, by=list(sub$index, sub$stress_level, sub$time, sub$incubation), FUN= "mean"), c("index", "stress_level", "time", "incubation","value")), setNames(aggregate(sub$value, by=list(sub$index, sub$stress_level, sub$time, sub$incubation), FUN= "sd"), c("index", "stress_level", "time", "incubation","sd")))%>%
     mutate(incubation = factor(incubation, levels =c("anaerobe", "aerobe")))%>%
   mutate(stress_level = factor(stress_level, levels=c("control", "low", "median", "high", "max"))) %>%
    subset(index %in% c("observed", "diversity_shannon")) ->means

sub %>%
  mutate(incubation = factor(incubation, levels =c("anaerobe", "aerobe")))%>%
  mutate(stress_level = factor(stress_level, levels=c("control", "low", "median", "high", "max"))) %>%
  subset(index %in% c("observed", "diversity_shannon"))%>%
  ggplot(aes(x=stress_level, y=value))+
   scale_fill_manual(values=c("skyblue","pink3"), labels=c("Control", "Stress"))+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 90, vjust= 0.5, hjust=1),
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
   geom_boxplot(aes( fill=incubation), outlier.shape = NA, position = position_dodge2(preserve = "single"))+
  # geom_boxplot(data = means, aes(lower = value-sd, upper = value+sd, middle = value, ymin = value-3*sd, ymax = value+3*sd, fill=incubation), outlier.shape = NA, position = 
  stat_pvalue_manual(stat_test, label = "p.signif", hide.ns = T)+
  geom_point(aes(colour=donor_name, group=incubation), position = position_jitterdodge(0.1))+
  scale_x_discrete(labels=c("0.12", "0.09", "0.07", "0.04", "0"))+
  labs(fill="Condition", colour="Donor")+
  ylab("Alpha diversity")+
  xlab("Agar [%]")+
  facet_grid(rows=vars(index), scales="free",
              labeller=facet_labeller
             )+
  guides(alpha = guide_legend(override.aes = list(fill = "darkgrey"))) 


aggregate(sub$value, by = list(sub$incubation, sub$stress_level, sub$time, sub$index), FUN=mean)%>%
  subset(Group.2 %in% "max")%>%
  subset(Group.3 %in% "stress")%>%
  subset(Group.4 %in% "observed")
  
ggsave(filename =  "alpha_o2.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final",
       width = 10,
       height = 7.5,
       units = c("cm"))
```


#Alpha compared to control ==>Distances
```{r}
phyloseq_rare%>% 
  phyloseq_alphas(phylo = TRUE)%>%
  subset(stress %in% "O2" )%>%
    left_join(., pd(t(as.data.frame(otu_table(phyloseq_rare))), phy_tree(phyloseq_rare))%>%
              rownames_to_column("sample_name"))%>%
  gather(index, value, "observed":"PD")->sub



sub[, c("donor_name", "time", "stress_level", "index", "value", "incubation")]%>%
  pivot_wider(values_from = value, names_from = incubation)%>%
  mutate(diff = aerobe - anaerobe)%>%
  mutate(stress_level = factor(stress_level, levels=c( "control","low", "median", "high", "max")),
         time = factor(time, levels=c( "stress", "post_stress_1"))) -> diff

time_cond<-c( "stress", "post_stress_1")
stress_cond<-c("control","low", "median", "high", "max")
index_cond <- c("observed", "diversity_shannon", "evenness_pielou")

results<-setNames(as.data.frame(matrix(data = NA, nrow = 1, ncol = 4)), c("time", "stress_level", "index", "p"))

for (i in 1:length(time_cond)){
  for (j in 1:length(stress_cond)){
    for(x in 1: length(index_cond)){
    subset(diff, time %in% time_cond[i] & stress_level %in% stress_cond[j] & index %in% index_cond[x])-> sub_test
    p<-(wilcox.test(sub_test$diff, mu = 0, paired = FALSE, exact = FALSE))$p.value
    
    results <- rbind(results, c( time_cond[i],stress_cond[j],index_cond[x] ,p ))
    
    }}}

results%>%
  subset(index %in% c("observed"))%>%
  subset(p <0.05)
```

```{r}
diff%>%
  subset(index %in% c("observed")
  )%>%
  ggplot(aes(x=stress_level, y=diff))+
  theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=20, angle = 90, vjust= 0.5, hjust=1),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_boxplot(aes(fill=time), outlier.shape = NA, position = position_dodge2(preserve = "single"))+
  geom_point(aes(colour=donor_name, group=time),  position=position_jitterdodge(0.1))+
  ggtitle(expression('??Alpha'~'diversity:'~O[2]~vs.~control))+
  ylab("??Shannon-Index")+
  labs(colour="Donor", fill="Stress level")+
  facet_grid(rows=vars(index), scales= "free") +
  geom_hline(yintercept=0, linetype=2)+
  # ylim(-1, 0.5)+
  scale_x_discrete(labels=c("minimal", "low", "median", "high", "max"))

aggregate(diff$diff, by=list(diff$time, diff$stress_level, diff$index), FUN=sd)

ggsave(filename = paste( "alpha_o2_delta.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/O2/alpha",
       width = 25,
       height = 10,
       units = c("cm"))
  

```


# Beta diversity
#Calculations of distances between pre culture and subsequent passages
```{r}
phyloseq_rare %>%
  subset_samples(!Sample_id_compete %in% c("feces_dil_1", "feces"))%>%
  subset_samples(!donor_name %in% NA)-> sub

sub%>%
  phyloseq::distance(method = "bray") -> bc

sub %>%
  ordinate(method = "PCoA",
           distance = bc) -> ord

sub %>% 
  phyloseq_distance_boxplot(p = ., 
                            dist = bc, 
                            d = "time") -> out # column in metadata for comparaisons

setNames(as.data.frame(out$matrix),c("sample_name", "sample_name_comp", "value", "condition_comp", "comparison")) -> distances

distances%>%
  subset(comparison %in% c("pre-stress"))-> distances

joined<-left_join(distances, sample.data.frame(sub)[,c(1,8:13)])

joined%>%
  mutate(sample_name_comp = recode(sample_name_comp, 
                                   "M-09" = 'D1',
                                   "M-10" = 'D2',
                                   "M-11" = 'D3',
                                   "M-12" = 'D4',
                                   "M-13" = 'D5',
                                   "M-14" = 'D6',
                                   "M-15" = 'D7',
                                   "M-16" = 'D8',))->joined

for (i in 1:nrow(joined)){
  if (joined[i,c("donor_name")] == joined[i,c("sample_name_comp")]){joined[i,c("match")]= "true"}
}

joined%>%
  subset(match %in% "true")->match_pre_culture
```
##Beta dist. over all passages (over all controls)
```{r}
match_pre_culture%>%
  subset(stress_level %in% c("control") & stress %in% c("H2O2", "O2", "control") & incubation %in% c("anaerobe"))%>%
  mutate(time = factor(time, levels=c("stress", "post_stress_1", "post_stress_2")))%>%
  ggplot(aes(x=time, y=value))+
  theme(text = element_text(size  = 17),
        axis.title.x = element_blank(),
        axis.text.x = element_text( angle=90),
        legend.text = element_text( vjust=1, hjust=0),
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_boxplot(outlier.shape=NA)+
  geom_jitter(aes(colour=donor_name), width=0.1)+
  ylab("Bray distance")+
    scale_x_discrete(labels=c("P1", "P2", "P3"))+
  labs(colour="Donor")

ggsave(filename = paste("beta_controls.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final",
       width = 7.5,
       height = 7.5,
       units = c("cm"))

```
##beta dist feces

```{r}
phyloseq_rare %>%
  subset_samples(!Sample_id_compete %in% c("feces_dil_1"))%>%
  subset_samples(!donor_name %in% NA)-> sub

sub%>%
  phyloseq::distance(method = "bray") -> bc

sub %>%
  ordinate(method = "PCoA",
           distance = bc) -> ord

sub %>% 
  phyloseq_distance_boxplot(p = ., 
                            dist = bc, 
                            d = "time") -> out # column in metadata for comparaisons

setNames(as.data.frame(out$matrix),c("sample_name", "sample_name_comp", "value", "condition_comp", "comparison")) -> distances

distances%>%
  subset(comparison %in% c("feces"))-> distances

joined<-left_join(distances, sample.data.frame(sub)[,c(1,8:13)])

joined%>%
  mutate(sample_name_comp = recode(sample_name_comp, 
                                   "M-02" = 'D2',
                                   "M-03" = 'D3',
                                   "M-04" = 'D4',
                                   "M-05" = 'D5',
                                   "M-06" = 'D6',
                                   "M-07" = 'D7',
                                   "M-08" = 'D1',))->joined

for (i in 1:nrow(joined)){
  if (joined[i,c("donor_name")] == joined[i,c("sample_name_comp")]){joined[i,c("match")]= "true"}
}

joined%>%
  subset(match %in% "true")->match
```

##Beta dist. over all passages (over all controls)
```{r}
match%>%
  subset(!time %in% c("post_stress_2"))%>%
  subset(stress_level %in% c("control") & stress %in% c("H2O2", "O2", "control") & incubation %in% c("anaerobe"))%>%
   mutate(time = factor(time, levels=c("pre-stress","stress", "post_stress_1")))->sub

sub%>%
  ggplot(aes(x=time, y=value))+
  theme(text = element_text(size  = 17),
        axis.title.x = element_blank(),
        axis.text.x = element_text( angle=90),
        legend.text = element_text( vjust=1, hjust=0),
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_boxplot(outlier.shape=NA)+
  geom_jitter(aes(colour=donor_name), width=0.1)+
  ylab("Bray distance")+
    scale_x_discrete(labels=c("P1", "P2", "P3", "P4"))+
  labs(colour="Donor")


left_join(left_join(setNames(aggregate(sub$value, by = list(sub$donor_name, sub$time), FUN="mean"), c("donor_name", "time", "mean")), 
setNames(aggregate(sub$value, by = list(sub$donor_name, sub$time), FUN="min"), c("donor_name", "time", "min"))),
setNames(aggregate(sub$value, by = list(sub$donor_name, sub$time), FUN="max"), c("donor_name", "time", "max")))%>%
rbind(., data.frame(donor_name = c("D1", "D2", "D3", "D4", "D5", "D6", "D7"),
                     time = rep("feces", 7),
                     mean = rep(0, 7),
                     min = rep(0, 7),
                     max = rep(0, 7)
                     ))%>%
  plyr::mutate(time = factor(time, levels=c("feces","pre-stress","stress", "post_stress_1")))%>%
  ggplot(aes(x=time, y=mean))+
  theme(text = element_text(size  = 15),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle=90, hjust=1, vjust=0.5),
        legend.text = element_text(vjust=1, hjust=0),
        strip.background = element_rect(fill="white"),
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), 
        panel.spacing = unit(0.31, "cm"),
        legend.text.align = 0)+
  geom_line(aes(group = donor_name))+
  geom_point(aes(group = donor_name, y=max), alpha=0.6)+
  geom_point(aes(group = donor_name, y=min), alpha=0.6)+
  ylab("Bray-Curtis")+
  # labs(colour="Donor")+
  scale_x_discrete(labels = c("Feces","Pre-culture", "Stress", "Post-stress"))+
  facet_grid(cols=vars(donor_name)) ->beta_controls




ggsave(filename = paste("beta_controls.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final",
       width = 7.5,
       height = 7.5,
       units = c("cm"))

```



#Beta distances H2O2
```{r}
names<-list("pre-stress" = "Pre", 
            "stress"= "Stress", 
            "post_stress_1"= "Post 1", 
            "post_stress_2"= "Post 2")

match%>%
  subset(stress %in% "H2O2" )%>%
  subset(time %in% c("stress", "post_stress_1"))->sub

sub%>%
  mutate(stress_level =as.character(stress_level))%>%
  wilcox_test(data= .,
  formula = value ~ stress_level, ref.group = "control")%>%
  mutate(group2 = factor(group2, levels=c("low", "median", "high")))%>%
  add_xy_position()%>%
   mutate(y.position = ifelse(group2 == "median", 0.8, y.position))%>%
   mutate(y.position = ifelse(group2 == "high", 1.1, y.position))%>%
   mutate(xmax = ifelse(group2 == "median", 3, xmax))%>%
   mutate(xmax = ifelse(group2 == "high", 4, xmax))->stat_test

sub %>%
  mutate(stress_level = factor(stress_level, c("control", "low", "median", "high")))%>%
  ggplot(aes(x=stress_level, y=value))+
  # scale_fill_manual(values=c("steelblue4", "steelblue3",  "coral2", "coral3"))+
 theme(text = element_text(size  = 15),
        axis.text.x = element_text(angle = 90, vjust= 0.5, hjust=1),
        legend.text = element_text(vjust=1, hjust=0),
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), 
       legend.text.align = 0)+
   geom_boxplot(aes(fill=time), outlier.shape = NA, position = position_dodge2(preserve = "single"))+
  geom_point(aes(colour=donor_name, group=time), position = position_jitterdodge(0.2), alpha =.9)+
  ylab("Bray distance")+
  labs(colour="Donor", fill="Stress level") +
   stat_pvalue_manual(stat_test, label = "p.adj.signif", hide.ns = T)+
  scale_x_discrete(labels = c("0", "0.22", "0.71", "2.3"))+
  xlab(expression(H[2]*O[2]~`[mM]`))

ggsave(filename = paste("beta_h2o2.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final",
       width = 7.5,
       height = 7.5,
       units = c("cm"))
```


#O2
```{r}
# match$time<-factor(match$time, levels=c("pre-stress", "stress", "post_stress_1", "post_stress_2"))
# match$stress_level<-factor(match$stress_level, levels=c("control", "low", "median", "high", "max"))

match%>%
  
  subset(stress %in% "O2" )%>%
  subset(time %in% "stress")->sub

sub%>%
  group_by(stress_level)%>%
  wilcox_test(data= .,
  formula = value ~ incubation)%>%
  add_significance()%>%
  add_xy_position()%>%
  mutate(xmax = 5.2,
         xmin = 4.8)->stat_test

sub %>%
  left_join(., setNames(aggregate(sub$value, by=list( sub$stress_level, sub$time, sub$incubation), FUN= "mean"), c( "stress_level", "time", "incubation","mean")))%>%
              left_join(.,   setNames(aggregate(sub$value, by=list( sub$stress_level, sub$time, sub$incubation), FUN= "mean"), c( "stress_level", "time", "incubation","sd")))%>%
  mutate(incubation = factor(incubation, levels =c("anaerobe", "aerobe")))%>%
    mutate(stress_level = factor(stress_level, c("control", "low", "median", "high", "max")))%>%
  ggplot(aes(x=stress_level, y=value))+
  theme(text = element_text(size  = 15),
        axis.text.x = element_text(angle=90, vjust=0.5, hjust=1),
        legend.text = element_text(vjust=0.5, hjust=0),
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), 
        legend.text.align = 0)+
  geom_boxplot(aes(fill=incubation), outlier.shape = NA, position = position_dodge2(preserve = "single"), alpha=0.9)+
  scale_fill_manual(values=c("skyblue", "lightpink3"), labels=c("Control", "Stress"), name="Condition")+
  geom_point(aes(colour=donor_name, group=incubation), position = position_jitterdodge(0.02), alpha=0.9)+
  ylab("Bray distance")+
  labs(colour = "Donor", fill= "Incubation")+
  scale_x_discrete(labels=c("0.12", "0.09", "0.07", "0.04", "0"))+
  labs(fill="Condition", colour="Donor")+
  xlab("Agar [%]")+
  stat_pvalue_manual(stat_test, label = "p.signif", hide.ns = T)


aggregate(sub$value, by = list(sub$incubation, sub$stress_level, sub$time), FUN=mean)%>%
  subset(Group.2 %in% "max")%>%
  subset(Group.3 %in% "stress")

ggsave(filename = "Beta_o2.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/final",
       width =10,
       height = 7.5,
       units = c("cm"))

```




#Distance between control conditions: H2O2
```{r}

phyloseq_rare %>%
  # subset_samples(time %in% "stress")%>%
  subset_samples(stress %in% c("H2O2"))-> sub

sub %>%
  phyloseq::distance(method = "bray") -> bc

sub %>%
  ordinate(method = "PCoA",
           distance = bc) -> ord
sub %>% 
  phyloseq_distance_boxplot(p = ., 
                            dist = bc, 
                            d = "stress_level") -> out# column in metadata for comparaisons

setNames(as.data.frame(out$matrix),c("sample_name", "sample_name_comp", "value", "condition_comp", "comparison")) -> distances

distances%>%
  subset(comparison %in% c("control"))-> distances

joined<-left_join(distances, sample.data.frame(sub)[,c(1,7:12)])

sample_name<-setNames(sample.data.frame(sub)[,c("sample_name", "donor_name", "time")], c("sample_name_comp", "donor_comp", "time_comp"))


joined_2<-left_join(joined, sample_name)
 

for (i in 1:nrow(joined_2)){
  if (joined_2[i,c("donor_name")] == joined_2[i,c("donor_comp")] & joined_2[i,c("time")] == joined_2[i,c("time_comp")]){joined_2[i,c("match")]= "true"}
}

joined_2%>%
  subset(match == c("true"))-> match_h2o2
```

```{r}
match_h2o2$stress_level<-factor(match_h2o2$stress_level, levels=c("low", "median", "high"))
match_h2o2$time<-factor(match_h2o2$time, levels=c("stress", "post_stress_1"))

names<-list("stress"= "Stress", 
            "post_stress_1"= "Post 1")

facet_labeller <- function(variable,value){
  return(names[value])
}


match_h2o2%>%
 subset(stress %in% "H2O2" )%>%
  ggplot(aes(x=time, y=value))+
  # scale_fill_manual(values=c("steelblue3",  "coral2", "coral3"))+
 theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_text(size=20, angle = 90, vjust= 0.5, hjust=1),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single"))+
  geom_point(aes(colour=donor_name, group=stress_level), position = position_jitterdodge(0.05))+
  ggtitle(expression('Beta'~distances~between~H[2]*O[2]~and~control))+
   scale_x_discrete(labels=c("Stress", "Post-stress"))+
  ylab("Bray distance")+
  labs(colour="Donor", fill="Stress level") +
   facet_grid(cols=vars(stress_level), scale="free")
   # force_panelsizes(rows = unit(8, "cm"),
   #                 cols = unit(11, "cm"))

ggsave(filename = paste("beta_h2o2.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/H2O2",
       width = 25,
       height = 10,
       units = c("cm"))
```

```{r}

time <- c("stress", "post_stress_1", "post_stress_2")
stress_level_cond <-c( "low", "median", "high")

for (i in 1:length(time)){
  for (j in 1:length(stress_level_cond)){
  phyloseq_rare %>%
  subset_samples(!Sample_id_compete %in% c("feces_dil_1", "feces"))%>%
  subset_samples(stress %in% "H2O2") %>%
    subset_samples(time %in% time[i]) %>%
    subset_samples(stress_level %in% c("control", stress_level_cond[j]))->sub
  
  sub %>%
  microbiome::transform("clr") -> ps_transformed

ps_transformed%>%
  phyloseq::distance(method = "euclidean") -> bc

ps_transformed %>% 
  physeq_betadisper(dm = bc,
                    variable = "stress_level") -> betadisp    
    
print(length(get_variable(sub,"stress_level")))  
    
  print(c(time[i], stress_level_cond[j]) )
  print(betadisp)
  }}


```

```{r}
# source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_beta.R")
time <- c("stress", "post_stress_1")
stress_level_cond <-c("low", "median", "high")

for (i in 1:length(time)){

  phyloseq_rare %>%
  subset_samples(!Sample_id_compete %in% c("feces_dil_1", "feces"))%>%
  subset_samples(stress %in% "O2" & incubation %in% "aerobe") %>%
    subset_samples(time %in% time[i] ) ->sub
  
sub %>%
  phyloseq::distance(method = "bray") -> bc
  
  sub%>%
    physeq_pairwise_permanovas_adonis2(dm = bc,
                                     compare_header = "stress_level") %>% 
  arrange(pvalFDR)->p
    
    
  print(c(time[i]))
  print(p)
  }

```



#Distance between control conditions: O2 - comparison of the different beta diversity measures
```{r}
phyloseq_rare %>%
  subset_samples(stress %in% c("O2"))-> sub

sub%>%
  phyloseq::distance(method = "bray") -> bc

ps_transformed %>%
  ordinate(method = "PCoA",
           distance = bc) -> ord
sub %>% 
  phyloseq_distance_boxplot(p = ., 
                            dist = bc, 
                            d = "incubation") -> out -> out # column in metadata for comparaisons

setNames(as.data.frame(out$matrix),c("sample_name", "sample_name_comp", "value", "condition_comp", "comparison")) -> distances

distances%>%
  subset(comparison %in% c("anaerobe"))-> distances

joined<-left_join(distances, sample.data.frame(sub)[,c(1,7:12)])

sample_name<-setNames(sample.data.frame(sub)[,c("sample_name", "incubation", "donor_name", "time", "stress_level")], c("sample_name_comp", "incubation_comp","donor_comp", "time_comp", "stress_level_comp"))


joined_2<-left_join(joined, sample_name)
 

for (i in 1:nrow(joined_2)){
  if ((joined_2[i,c("condition_comp")] != joined_2[i,c("incubation_comp")]) & 
      (joined_2[i,c("donor_name")] == joined_2[i,c("donor_comp")]) & 
      (joined_2[i,c("time")] == joined_2[i,c("time_comp")]) &
      (joined_2[i,c("stress_level")] == joined_2[i,c("stress_level_comp")])) {joined_2[i,c("match")]= "true"}
}

joined_2%>%
  subset(match == c("true"))%>%
  mutate(index = "bray")-> match_o2_bray
```

```{r}
match_o2_bray%>%
  group_by( time)%>%
  t_test(data= .,
  formula = value ~ stress_level)%>%
  add_significance()%>%
  add_xy_position()->stat_test

match_o2_bray %>%
  mutate(time = factor(time, levels = c("stress", "post_stress_1")))%>%
  mutate(stress_level = factor(stress_level, c("control", "low", "median", "high", "max")))%>%
  ggplot(aes(x=time, y=value))+
  # scale_fill_manual(values=c("steelblue4", "steelblue3",  "coral2", "coral3"))+
 theme(text = element_text(size  = 15),
        axis.text.x = element_text(angle = 90, vjust= 0.5, hjust=1),
        legend.text = element_text(vjust=1, hjust=0),
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), 
       legend.text.align = 0)+
   geom_boxplot( outlier.shape = NA, position = position_dodge2(preserve = "single"))+
  geom_point(aes(colour=donor_name, group=time), position = position_jitterdodge(0.2), alpha =.9)+
  ylab("Bray distance")+
  labs(colour="Donor", fill="Stress level") +
  #  stat_pvalue_manual(stat_test, label = "p.adj.signif", hide.ns = T)+
  scale_x_discrete(labels = c("Stress", "Post-stress"))+
  xlab(expression(O[2]~`[mM]`))+
  facet_grid(cols=vars(stress_level))

aggregate(match_o2_bray$value, by=list(match_o2_bray$time, match_o2_bray$stress_level), FUN=mean)
aggregate(match_o2_bray$value, by=list(match_o2_bray$time, match_o2_bray$stress_level), FUN=sd)
```


```{r}
phyloseq_rare %>%
  # subset_samples(time %in% "stress")%>%
  subset_samples(stress %in% c("O2"))-> sub

# sub %>%
#   microbiome::transform("clr") -> ps_transformed

sub%>%
  phyloseq::distance(method = "unifrac") -> bc

ps_transformed %>%
  ordinate(method = "PCoA",
           distance = bc) -> ord
sub %>% 
  phyloseq_distance_boxplot(p = ., 
                            dist = bc, 
                            d = "incubation") -> out -> out # column in metadata for comparaisons

setNames(as.data.frame(out$matrix),c("sample_name", "sample_name_comp", "value", "condition_comp", "comparison")) -> distances

distances%>%
  subset(comparison %in% c("anaerobe"))-> distances

joined<-left_join(distances, sample.data.frame(sub)[,c(1,7:12)])

sample_name<-setNames(sample.data.frame(sub)[,c("sample_name", "incubation", "donor_name", "time", "stress_level")], c("sample_name_comp", "incubation_comp","donor_comp", "time_comp", "stress_level_comp"))


joined_2<-left_join(joined, sample_name)
 

for (i in 1:nrow(joined_2)){
  if ((joined_2[i,c("condition_comp")] != joined_2[i,c("incubation_comp")]) & 
      (joined_2[i,c("donor_name")] == joined_2[i,c("donor_comp")]) & 
      (joined_2[i,c("time")] == joined_2[i,c("time_comp")]) &
      (joined_2[i,c("stress_level")] == joined_2[i,c("stress_level_comp")])) {joined_2[i,c("match")]= "true"}
}

joined_2%>%
  subset(match == c("true"))%>%
  mutate(index = "unifrac")-> match_o2_unifrac
```



```{r}
phyloseq_rare %>%
  # subset_samples(time %in% "stress")%>%
  subset_samples(stress %in% c("O2"))-> sub

# sub %>%
#   microbiome::transform("clr") -> ps_transformed

sub%>%
  phyloseq::distance(method = "jaccard", binary=F) -> bc

ps_transformed %>%
  ordinate(method = "PCoA",
           distance = bc) -> ord
sub %>% 
  phyloseq_distance_boxplot(p = ., 
                            dist = bc, 
                            d = "incubation") -> out -> out # column in metadata for comparaisons

setNames(as.data.frame(out$matrix),c("sample_name", "sample_name_comp", "value", "condition_comp", "comparison")) -> distances

distances%>%
  subset(comparison %in% c("anaerobe"))-> distances

joined<-left_join(distances, sample.data.frame(sub)[,c(1,7:12)])

sample_name<-setNames(sample.data.frame(sub)[,c("sample_name", "incubation", "donor_name", "time", "stress_level")], c("sample_name_comp", "incubation_comp","donor_comp", "time_comp", "stress_level_comp"))


joined_2<-left_join(joined, sample_name)
 

for (i in 1:nrow(joined_2)){
  if ((joined_2[i,c("condition_comp")] != joined_2[i,c("incubation_comp")]) & 
      (joined_2[i,c("donor_name")] == joined_2[i,c("donor_comp")]) & 
      (joined_2[i,c("time")] == joined_2[i,c("time_comp")]) &
      (joined_2[i,c("stress_level")] == joined_2[i,c("stress_level_comp")])) {joined_2[i,c("match")]= "true"}
}

joined_2%>%
  subset(match == c("true"))%>%
  mutate(index = "wjaccard")-> match_o2_wjaccad
```

```{r}
phyloseq_rare %>%
  # subset_samples(time %in% "stress")%>%
  subset_samples(stress %in% c("O2"))-> sub

# sub %>%
#   microbiome::transform("clr") -> ps_transformed

sub%>%
  phyloseq::distance(method = "jaccard", binary =T) -> bc

ps_transformed %>%
  ordinate(method = "PCoA",
           distance = bc) -> ord
sub %>% 
  phyloseq_distance_boxplot(p = ., 
                            dist = bc, 
                            d = "incubation") -> out -> out # column in metadata for comparaisons

setNames(as.data.frame(out$matrix),c("sample_name", "sample_name_comp", "value", "condition_comp", "comparison")) -> distances

distances%>%
  subset(comparison %in% c("anaerobe"))-> distances

joined<-left_join(distances, sample.data.frame(sub)[,c(1,7:12)])

sample_name<-setNames(sample.data.frame(sub)[,c("sample_name", "incubation", "donor_name", "time", "stress_level")], c("sample_name_comp", "incubation_comp","donor_comp", "time_comp", "stress_level_comp"))


joined_2<-left_join(joined, sample_name)
 

for (i in 1:nrow(joined_2)){
  if ((joined_2[i,c("condition_comp")] != joined_2[i,c("incubation_comp")]) & 
      (joined_2[i,c("donor_name")] == joined_2[i,c("donor_comp")]) & 
      (joined_2[i,c("time")] == joined_2[i,c("time_comp")]) &
      (joined_2[i,c("stress_level")] == joined_2[i,c("stress_level_comp")])) {joined_2[i,c("match")]= "true"}
}

joined_2%>%
  subset(match == c("true"))%>%
  mutate(index = "jaccard")-> match_o2_jaccard
```

```{r}
phyloseq_rare %>%
  # subset_samples(time %in% "stress")%>%
  subset_samples(stress %in% c("O2"))-> sub

# sub %>%
#   microbiome::transform("clr") -> ps_transformed

sub%>%
  phyloseq::distance(method = "wunifrac") -> bc

ps_transformed %>%
  ordinate(method = "PCoA",
           distance = bc) -> ord
sub %>% 
  phyloseq_distance_boxplot(p = ., 
                            dist = bc, 
                            d = "incubation") -> out -> out # column in metadata for comparaisons

setNames(as.data.frame(out$matrix),c("sample_name", "sample_name_comp", "value", "condition_comp", "comparison")) -> distances

distances%>%
  subset(comparison %in% c("anaerobe"))-> distances

joined<-left_join(distances, sample.data.frame(sub)[,c(1,7:12)])

sample_name<-setNames(sample.data.frame(sub)[,c("sample_name", "incubation", "donor_name", "time", "stress_level")], c("sample_name_comp", "incubation_comp","donor_comp", "time_comp", "stress_level_comp"))


joined_2<-left_join(joined, sample_name)
 

for (i in 1:nrow(joined_2)){
  if ((joined_2[i,c("condition_comp")] != joined_2[i,c("incubation_comp")]) & 
      (joined_2[i,c("donor_name")] == joined_2[i,c("donor_comp")]) & 
      (joined_2[i,c("time")] == joined_2[i,c("time_comp")]) &
      (joined_2[i,c("stress_level")] == joined_2[i,c("stress_level_comp")])) {joined_2[i,c("match")]= "true"}
}

joined_2%>%
  subset(match == c("true"))%>%
  mutate(index = "wunifraq")-> match_o2_wunifraq
```

```{r}
rbind(match_o2_bray, match_o2_unifrac)%>%
  rbind(match_o2_wunifraq)%>%
  rbind(match_o2_jaccard)%>%
  rbind(match_o2_wjaccad)->match_idex_o2
```

```{r}

match_idex_o2%>%
  mutate(stress_level = factor(stress_level, levels=c("control","low", "median", "high", "max")))%>%
  # mutate(time = factor(time, c("stress", "post_stress_1", "post_stress_2")))%>%
  # subset(time %in% c("stress"))%>%
  ggplot(aes(x=stress_level, y=value))+
  theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_text(size=20, angle = 90, vjust= 0.5, hjust=1),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
   geom_boxplot( outlier.shape = NA, aes(fill=index))+
  geom_point(aes(colour=donor_name, group=index), position = position_dodge(0.9))+
   ggtitle(expression('Beta'~distances~between~O[2]~and~control))+
  scale_x_discrete(labels=c("minimal", "low", "medium", "high", "max"))+
  ylab("Distance")+
  labs(colour="Donor", fill="Stress level") +
  facet_grid(cols=vars(index), scale="free", rows=vars(time))->p

p
  


ggsave(filename = paste(x, "O2_alpha_o2.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/O2/beta",
       width = 25,
       height = 10,
       units = c("cm"))


```

```{r}

time <- c("stress", "post_stress_1")
stress_level_cond <-c("control", "low", "median", "high", "max")

for (i in 1:length(time)){
  for (j in 1:length(stress_level_cond)){
  phyloseq_rare %>%
  subset_samples(!Sample_id_compete %in% c("feces_dil_1", "feces"))%>%
  subset_samples(stress %in% "O2") %>%
    subset_samples(time %in% time[i]) %>%
    subset_samples(stress_level %in% c(stress_level_cond[j]))->sub
  


sub%>%
  phyloseq::distance(method = "bray") -> bc

sub %>% 
  physeq_betadisper(dm = bc,
                    variable = "incubation") -> betadisp    
    
print(length(get_variable(sub,"incubation")))  
    
  print(c(time[i], stress_level_cond[j]) )
  print(betadisp)
  }}


```


```{r}
time <- c("stress")
stress_level_cond <-c("median","high","max")

for (i in 1:length(time)){
  for (j in 1:length(stress_level_cond)){
  phyloseq_rare %>%
  subset_samples(!Sample_id_compete %in% c("feces_dil_1", "feces"))%>%
  subset_samples(stress %in% "O2") %>%
    subset_samples(time %in% time[i]) %>%
    subset_samples(stress_level %in% c(stress_level_cond[j]))->sub
  
  sub %>%
  phyloseq::distance(method = "bray") -> bc
    
    vegan::adonis2(bc ~ get_variable(sub,"incubation")+get_variable(sub,"donor_name"))->adonis
    
print(length(get_variable(sub,"incubation")))  
    
  print(c(time[i], stress_level_cond[j]) )
  print(adonis)
  }}
```

```{r}
phyloseq_rare %>%
  subset_samples(!Sample_id_compete %in% c("feces_dil_1", "feces"))%>%
  subset_samples(stress %in% "O2" ) %>%
  subset_samples(time %in% "stress")%>%
  # subset_samples(stress_level %in% c("high", "max"))%>%
    subset_samples(incubation %in% "anaerobe")->sub
  
 sub %>%
  microbiome::transform("clr") -> ps_transformed

ps_transformed%>%
  phyloseq::distance(method = "euclidean") -> bc
    
    vegan::adonis2(bc ~ get_variable(sub,"stress_level")  , p.adjust.methods= "BH")->adonis
    
    
    
  print(c(time[i], stress_level_cond[j]))
  print(adonis)

```


# PCoA -02 aerobic incubation anaerobic incubation
```{r}
library(plyr)


phyloseq_rare %>%
  subset_samples(time %in% c("stress"))%>%
   # subset_samples(time %in% c("post_stress_2", "post_stress_1"))%>%
   # subset_samples(!donor_name %in% c("D1", "D5"))%>%
  # subset_samples(stress_level %in% c("max"))%>%
  subset_samples(stress %in% c("O2"))-> sub


  dist = "bray"
  ord_meths = c("PCoA")


  plist = llply(as.list(ord_meths), function(i, sub, dist){
          ordi = ordinate(sub, method=i, distance=dist)
          plot_ordination(physeq = sub, 
                          ordination = ordi,
                          color = "donor_name", 
                          shape = "stress",
                          title = "PCoA Bray-Curtis", 
                          label= "time")},
          sub, 
          dist)


  names(plist) <- ord_meths

  pdataframe = ldply(plist, function(x){
      df = x$data[, 1:2]
      colnames(df) = c("Axis_1", "Axis_2")
      return(cbind(df, x$data))})


  
  names(pdataframe)[1] = "method"


  pdataframe%>%
      ggplot(aes(Axis_1, Axis_2))+ 
        scale_colour_manual(values=c( "lightpink3","skyblue3"), labels=c("Control", "Stress"), name="Condition")+
    scale_shape_manual(values= c(15,  16, 17, 0, 1, 2, 3, 4), name="Donor")+
      geom_point(aes(colour=incubation, shape=donor_name), size=4, stroke=1.5, alpha=.85)+
    stat_ellipse(aes(group=incubation, colour=incubation))+
      theme(axis.text.y= element_text(size = 20), 
          axis.title.y = element_text(size=20), 
          axis.title.x = element_text(size=20),
          axis.text.x = element_text(size=20),
          legend.text = element_text(size = 20),
          legend.title = element_text(size= 20),
          plot.title = element_text(size=20), 
          legend.text.align = 0, 
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
    xlab("PCoA1 (27.1%)")+
    ylab("PCoA2 (19.3%)")+
    ggtitle("Stress")
  
  ggsave(filename = paste("O2_beta_stress.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/O2/beta",
       width = 16,
       height = 11,
       units = c("cm"))  
  
  
  phyloseq_rare %>%
  # subset_samples(time %in% c("stress"))%>%
    subset_samples(time %in% c("post_stress_2", "post_stress_1"))%>%
    # subset_samples(!donor_name %in% c( "D1", "D5"))%>%
  # subset_samples(stress_level %in% c("max"))%>%
  subset_samples(stress %in% c("O2"))-> sub


  dist = "bray"
  ord_meths = c("PCoA")


  plist = llply(as.list(ord_meths), function(i, sub, dist){
          ordi = ordinate(sub, method=i, distance=dist)
          plot_ordination(physeq = sub, 
                          ordination = ordi,
                          color = "donor_name", 
                          shape = "stress",
                          title = "PCoA Bray-Curtis", 
                          label= "time")},
          sub, 
          dist)


  names(plist) <- ord_meths

  pdataframe = ldply(plist, function(x){
      df = x$data[, 1:2]
      colnames(df) = c("Axis_1", "Axis_2")
      return(cbind(df, x$data))})


  
  names(pdataframe)[1] = "method"


  pdataframe%>%
      ggplot(aes(Axis_1, Axis_2))+ 
        scale_colour_manual(values=c( "lightpink3","skyblue3"), labels=c("Control", "Stress"), name="Condition")+
    scale_shape_manual(values= c(15,  16, 17, 0, 1, 2, 3, 4), name="Donor")+
      geom_point(aes(colour=incubation, shape=donor_name), size=4, stroke=1.5, alpha=.85)+
    stat_ellipse(aes(group=incubation, colour=incubation))+
      theme(axis.text.y= element_text(size = 20), 
          axis.title.y = element_text(size=20), 
          axis.title.x = element_text(size=20),
          axis.text.x = element_text(size=20),
          legend.text = element_text(size = 20),
          legend.title = element_text(size= 20),
          plot.title = element_text(size=20), 
          legend.text.align = 0, 
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
    xlab("PCoA1 (20.6%)")+
    ylab("PCoA2 (18.6%)")+
    ggtitle("Post-stress")
  
  
    ggsave(filename = paste("O2_beta_post.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/O2/beta",
       width = 16,
       height = 11,
       units = c("cm"))
  
  
```

# PCoA -02 stress levels

```{r}
library(plyr)


phyloseq_rare %>%
  subset_samples(time %in% c("stress") & incubation %in% "aerobe")%>%
   # subset_samples(time %in% c("post_stress_2", "post_stress_1"))%>%
   # subset_samples(!donor_name %in% c("D1", "D5"))%>%
  # subset_samples(stress_level %in% c("max"))%>%
  subset_samples(stress %in% c("O2"))-> sub


  dist = "bray"
  ord_meths = c("PCoA")


  plist = llply(as.list(ord_meths), function(i, sub, dist){
          ordi = ordinate(sub, method=i, distance=dist)
          plot_ordination(physeq = sub, 
                          ordination = ordi,
                          color = "donor_name", 
                          shape = "stress",
                          title = "PCoA Bray-Curtis", 
                          label= "time")},
          sub, 
          dist)


  names(plist) <- ord_meths

  pdataframe = ldply(plist, function(x){
      df = x$data[, 1:2]
      colnames(df) = c("Axis_1", "Axis_2")
      return(cbind(df, x$data))})


  
  names(pdataframe)[1] = "method"


  pdataframe%>%
      ggplot(aes(Axis_1, Axis_2))+ 
        # scale_colour_manual(values=c("skyblue3", "lightpink3"), labels=c("Control", "Stress"), name="Condition")+
    scale_shape_manual(values= c(15,  16, 17, 0, 1, 2, 3, 4), name="Donor")+
      geom_point(aes(colour=stress_level, shape=donor_name), size=4, stroke=1.5, alpha=.85)+
    stat_ellipse(aes(group=stress_level, colour=incubation))+
      theme(axis.text.y= element_text(size = 20), 
          axis.title.y = element_text(size=20), 
          axis.title.x = element_text(size=20),
          axis.text.x = element_text(size=20),
          legend.text = element_text(size = 20),
          legend.title = element_text(size= 20),
          plot.title = element_text(size=20), 
          legend.text.align = 0, 
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
    xlab("PCoA1 (27.1%)")+
    ylab("PCoA2 (19.3%)")
  
  ggsave(filename = paste("O2_beta_stress.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/O2/beta",
       width = 16,
       height = 11,
       units = c("cm"))  
  
  
  phyloseq_rare %>%
  # subset_samples(time %in% c("stress"))%>%
    subset_samples(time %in% c("post_stress_2", "post_stress_1"))%>%
    # subset_samples(!donor_name %in% c( "D1", "D5"))%>%
  # subset_samples(stress_level %in% c("max"))%>%
  subset_samples(stress %in% c("O2"))-> sub


  dist = "bray"
  ord_meths = c("PCoA")


  plist = llply(as.list(ord_meths), function(i, sub, dist){
          ordi = ordinate(sub, method=i, distance=dist)
          plot_ordination(physeq = sub, 
                          ordination = ordi,
                          color = "donor_name", 
                          shape = "stress",
                          title = "PCoA Bray-Curtis", 
                          label= "time")},
          sub, 
          dist)


  names(plist) <- ord_meths

  pdataframe = ldply(plist, function(x){
      df = x$data[, 1:2]
      colnames(df) = c("Axis_1", "Axis_2")
      return(cbind(df, x$data))})


  
  names(pdataframe)[1] = "method"


  pdataframe%>%
      ggplot(aes(Axis_1, Axis_2))+ 
        scale_colour_manual(values=c( "lightpink3","skyblue3"), labels=c("Control", "Stress"), name="Condition")+
    scale_shape_manual(values= c(15,  16, 17, 0, 1, 2, 3, 4), name="Donor")+
      geom_point(aes(colour=incubation, shape=donor_name), size=4, stroke=1.5, alpha=.85)+
    stat_ellipse(aes(group=incubation, colour=incubation))+
      theme(axis.text.y= element_text(size = 20), 
          axis.title.y = element_text(size=20), 
          axis.title.x = element_text(size=20),
          axis.text.x = element_text(size=20),
          legend.text = element_text(size = 20),
          legend.title = element_text(size= 20),
          plot.title = element_text(size=20), 
          legend.text.align = 0, 
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
    xlab("PCoA1 (20.6%)")+
    ylab("PCoA2 (18.6%)")
  
  
    ggsave(filename = paste("O2_beta_post.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/O2/beta",
       width = 16,
       height = 11,
       units = c("cm"))
  
  
```


# PCoA -H202
```{r}
# library(plyr)


phyloseq_rare %>%
  # subset_samples(time %in% "stress")%>%
  # subset_samples(!donor_name %in% c("D1", "D5"))%>%
   subset_samples(stress_level %in% c("control","low","median","high"))%>%
  subset_samples(stress %in% c("H2O2") & time %in% "stress")-> sub


  dist = "bray"
  ord_meths = c("PCoA")


  plist = llply(as.list(ord_meths), function(i, sub, dist){
          ordi = ordinate(sub, method=i, distance=dist)
          plot_ordination(physeq = sub, 
                          ordination = ordi,
                          color = "donor_name", 
                          shape = "stress",
                          title = "PCoA Bray-Curtis", 
                          label= "time")},
          sub, 
          dist)


  names(plist) <- ord_meths

  pdataframe = ldply(plist, function(x){
      df = x$data[, 1:2]
      colnames(df) = c("Axis_1", "Axis_2")
      return(cbind(df, x$data))})


  
  names(pdataframe)[1] = "method"


  pdataframe%>%
    mutate(stress_level = factor(stress_level, levels=c("control", "low", "median", "high")))->sub
      ggplot(data=sub, aes(Axis_1, Axis_2))+ 
        stat_ellipse(data = subset(sub, time %in% "stress"), aes(group=stress_level, fill=stress_level), geom="polygon", alpha=0.2, lty=1, size=1.5)+
         # stat_ellipse(data = subset(sub, time %in% "post_stress_2"), aes(group=stress_level, colour=stress_level), lty=2, size=1, alpha=.8)+
      geom_point(aes(colour=stress_level, shape=donor_name), size=4, alpha=0.8)+
      theme(axis.text.y= element_text(size = 20), 
          axis.title.y = element_text(size=20), 
          axis.title.x = element_text(size=20),
          axis.text.x = element_text(size=20),
          legend.text = element_text(size = 20),
          legend.title = element_text(size= 20),
          plot.title = element_text(size=20), 
          legend.text.align = 0, 
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
    scale_colour_manual(values=c("skyblue2","skyblue3", "lightpink3", "lightpink4"), 
                        # labels=c("control", expression("222"~mu*M), expression("710"~mu*M),  "2.3 mM"))+
                        labels=c("control", "low", "median", "high"), name="Stress level")+
     scale_fill_manual(values=c("skyblue2","skyblue3", "lightpink3", "lightpink4"), name="Stress level")+
     scale_shape_manual(values= c(15,  16, 17, 0, 1, 2, 3, 4), name="Donor")+
  xlab("PCoA1 (30.6 %)")+
    ylab("PCoA2 (14 %)")+
    scale_alpha_manual(values=c(1, 0.7, 0.6))+
        ggtitle("Stress")
      
      ggsave(filename = paste("H2O2_beta_ocoa.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/H2O2/beta",
       width = 16,
       height = 18,
       units = c("cm"))
      
      
      phyloseq_rare %>%
  # subset_samples(time %in% "stress")%>%
  # subset_samples(!donor_name %in% c("D1", "D5"))%>%
   subset_samples(stress_level %in% c("control","low","median","high"))%>%
  subset_samples(stress %in% c("H2O2") & !time %in% "stress")-> sub


  dist = "bray"
  ord_meths = c("PCoA")


  plist = llply(as.list(ord_meths), function(i, sub, dist){
          ordi = ordinate(sub, method=i, distance=dist)
          plot_ordination(physeq = sub, 
                          ordination = ordi,
                          color = "donor_name", 
                          shape = "stress",
                          title = "PCoA Bray-Curtis", 
                          label= "time")},
          sub, 
          dist)


  names(plist) <- ord_meths

  pdataframe = ldply(plist, function(x){
      df = x$data[, 1:2]
      colnames(df) = c("Axis_1", "Axis_2")
      return(cbind(df, x$data))})


  
  names(pdataframe)[1] = "method"


  pdataframe%>%
    mutate(stress_level = factor(stress_level, levels=c("control", "low", "median", "high")))->sub
      ggplot(data=sub, aes(Axis_1, Axis_2))+ 
        stat_ellipse(aes(group=stress_level, fill=stress_level), geom="polygon", alpha=0.2, lty=1, size=1.5)+
      geom_point(aes(colour=stress_level, shape=donor_name), size=4, alpha=0.8)+
      theme(axis.text.y= element_text(size = 20), 
          axis.title.y = element_text(size=20), 
          axis.title.x = element_text(size=20),
          axis.text.x = element_text(size=20),
          legend.text = element_text(size = 20),
          legend.title = element_text(size= 20),
          plot.title = element_text(size=20), 
          legend.text.align = 0, 
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
    scale_colour_manual(values=c("skyblue2","skyblue3", "lightpink3", "lightpink4"), 
                        # labels=c("control", expression("222"~mu*M), expression("710"~mu*M),  "2.3 mM"))+
                        labels=c("control", "low", "median", "high"), name="Stress level")+
     scale_fill_manual(values=c("skyblue2","skyblue3", "lightpink3", "lightpink4"), name="Stress level")+
     scale_shape_manual(values= c(15,  16, 17, 0, 1, 2, 3, 4), name="Donor")+
  xlab("PCoA1 (29.9 %)")+
    ylab("PCoA2 (14.4 %)")+
    scale_alpha_manual(values=c(1, 0.7, 0.6))+
        ggtitle("Post-stress")
      
      ggsave(filename = paste("H2O2_beta_post.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/H2O2/beta",
       width = 16,
       height = 10,
       units = c("cm"))

  
  
```




# ppcoa with phylogeny

```{r}
phyloseq_rare %>%
  subset_samples(time %in% c("stress"))%>%
  subset_samples(stress %in% c("O2") )%>%
  subset_samples(donor_name %in% c("D8", "D7"))-> sub

ordu=ordinate(sub, "PCoA", "unifrac", weighted = T)
plot_ordination(sub, ordu, color="donor_name", shape="stress_level") +
  scale_shape_manual(values= c(15,  16, 17, 0, 1, 2, 3, 4), name="Donor")+
   stat_ellipse(aes(group=incubation, colour=incubation))

  
```
#### ODLD code
#alpha comapred to control ==>Distances
```{r}

phyloseq_rare%>% 
  phyloseq_alphas(phylo = TRUE)-> alphas

###ad control value to all stress-level
alphas%>%
  subset(stress %in% "H2O2" & stress_level %in% "control")->low
low$stress_level <- "low"
low$stress <- "control"

alphas%>%
  subset(stress %in% "H2O2" & stress_level %in% "control")->median
median$stress_level <- "median"
median$stress <- "control"


alphas%>%
  subset(stress %in% "H2O2" & stress_level %in% "control")->high
high$stress_level <- "high"
high$stress <- "control"

rbind(subset(alphas, stress %in% "H2O2"), low, median, high)->alphas_comparison

names<-list("stress"= "Stress", 
            "post_stress_1"= "Post 1", 
            "post_stress_2"= "Post 2")


alphas_comparison[, c("donor_name", "time", "stress_level", "stress", "diversity_shannon")]%>%
  subset(!stress_level %in% "control" & !time %in% "pre-stress")%>%
  pivot_wider(values_from = diversity_shannon, names_from = stress)%>%
  mutate(diff = H2O2-control)%>%
  mutate(stress_level = factor(stress_level, levels=c( "low", "median", "high")),
         time = factor(time, levels=c( "stress", "post_stress_1", "post_stress_2"))) %>%
  ggplot(aes(x=stress_level, y=diff))+
  theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=20, angle = 90, vjust= 0.5, hjust=1),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_boxplot( outlier.shape = NA, position = position_dodge2(preserve = "single"))+
  geom_jitter(aes(colour=donor_name),  width = 0.1)+
  ggtitle(expression('??Alpha'~'diversity:'~H[2]*O[2]~vs.~control))+
  ylab("??Shannon-Index")+
  labs(colour="Donor", fill="Stress level")+
  facet_grid(cols=vars(time),scales= "free", labeller=facet_labeller) +
  geom_hline(yintercept=0, linetype=2)->p

p



ggsave(filename = paste( "alpha_h2o2_delta.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/H2O2/alpha",
       width = 25,
       height = 10,
       units = c("cm"))

```






#stats
```{r}
alphas%>%
  # subset(!donor_name %in% "D3")%>%
  subset(!time %in% c("pre-stress"))%>%
  group_by(time, stress_level) %>%
  t_test(data= .,
  formula = evenness_pielou ~ incubation ,
  )%>%
  add_significance()
```

#line plots
```{r}
match%>%
  subset(stress %in% "H2O2" )%>%
  ggplot(aes(x=time, y=value))+
  theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=20, angle=90),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20),  panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_line(aes(colour=stress_level, group=stress_level), size=1)+
  facet_grid( cols=vars(donor_name))+
  geom_point(aes(colour=stress_level, group=stress_level))+
  ggtitle("Beta dist.: pre-stress vs. culures:H2O2")

ggsave(filename = "Beta_dist_H2O2.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/H2O2",
       width = 30,
       height = 17,
       units = c("cm"))

match$stress_level<-factor(match$stress_level, levels=c("control", "low", "median", "high", "max"))

```

