---
title: "C-sources & 8 donors"
---

Link to Florentin's github page: https://htmlpreview.github.io/?https://github.com/fconstancias/DivComAnalyses/blob/master/SOP/community_analyses/Microbiome_R_tutorial.html#alpha-diversity

1. Getting ready: load all packages
-newest version of R
-rtools required 


```{r setup, echo =FALSE}
.libPaths('C:\\Users\\zuendj\\Documents\\R4Libs')
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "P:/Shared_documents/Serafina Plüss/20230320_sequencing_results/Janina", cho = TRUE, warning = FALSE, message = FALSE)

# library("plyr")
# install.packages("vctrs")
# install.packages("vegan")
# install.packages("cli")

# library(vctrs)
# library(vegan)
# install.packages("ggplot2")
#library(ggplot2)
#install.packages('tinytex')
#install.packages('rmarkdown')
#install.packages("rlang")
#install.packages("nlme")
#install.packages("mgcv")
#install.packages("ape")
#library(nlme)
# install.packages("remotes")
# remotes::install_github("mikemc/speedyseq")
# install.packages("BiocManager")
# BiocManager::install("DESeq2")
#install.packages("fs")
# install.packages('devtools')
#install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
#unload("mia")
# install.packages("GUniFrac")

# install.packages(
#   "microViz",
#   repos = c(davidbarnett = "https://david-barnett.r-universe.dev", getOption("repos"))
# )


 # devtools::install_github("tidyverse/tidyverse")
 # library(tidyverse)
  # library(openxlsx)
# # install.packages("ggalluvial")
# library(ggalluvial)
 # library(ggh4x)
# library(GUniFrac)
 # library(rstatix)
# library(tinytex)
# library(rmarkdown)
# library(rlang) 
# library(tidyverse)
 # library(microViz)
# library(dplyr)
# library(mgcv)
# library(ape)
# library(phyloseq)
# library(speedyseq)
# library(DESeq2)
# library(plyr)
# library(fs)
# library(devtools)
# library(pairwiseAdonis)
# library(ggpubr)

# install.packages("dplyr")
# install.packages("data.table")
# devtools::install_github("vmikk/metagMisc")
#  
# library(metagMisc)


```

```{r, echo =FALSE}
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_taxa_tests.R") 
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_normalisation.R") 
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_alpha.R") 
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_beta.R") 
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_heatmap.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_taxa_tests.R")
source("https://raw.githubusercontent.com/fconstancias/metabaRpipe-source/master/Rscripts/functions.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_varia.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_beta.R")

#get metadata file: load this function
sample.data.frame <- function(ps) {
  return(as(phyloseq::sample_data(ps), "data.frame"))}
```
## 1. Get the phyloseq object 
```{r}
# unloadNamespace("miaViz")
# unloadNamespace("mia")

phyloseq <-readRDS("P:/Shared_documents/Janina_Zuend/08_phd/01_sequencing_data/Ox_stress/phyloseq.RDS")
phyloseq%>%
  phyloseq_get_strains()->phyloseq
```

```{r}
hplc<-read.xlsx("C:/Users/zuendj/Desktop/03_data R/wp2_complex/input/20230613_hplc_list_results_stress.xlsx")
hplc%>%
  subset(!sample_name %in% NA)->hplc

seq<-read.xlsx("P:/Shared_documents/Janina_Zuend/08_phd/01_sequencing_data/Ox_stress/mapping_file.xlsx")

meta<-left_join(seq, hplc[,c(1:10)])
write.xlsx(meta, "P:/Shared_documents/Janina_Zuend/08_phd/01_sequencing_data/Ox_stress/meta_new.xlsx")
```
### Get and update metadata file
```{r, echo=FALSE}

phyloseq%>%
  physeq_add_metadata(physeq = .,
                      metadata = "P:/Shared_documents/Janina_Zuend/08_phd/01_sequencing_data/Ox_stress/meta.xlsx" %>%
                        readxl::read_xlsx(),
                      sample_column = "sample_name") -> phyloseq

as.data.frame(sample_data(phyloseq))->meta
## remove donor 2 from data as very low sample size
# phyloseq %>%
#   subset_samples(!donor_name %in% "D2")->phyloseq



```
#Rarefying
```{r}
phyloseq%>%
  phyloseq_check_lib_size(data_color = "donor_name",
                          data_facet = NULL,
                          nreads_display = 3165,
                       first_n = nsamples(phyloseq)) -> lib

lib$df %>%
  subset(sample_name %in% "M-99")

phyloseq %>%
  rarefy_even_depth(rngseed = 123,
                    sample.size = 3165
                    # sample.size = 5000
                    ) -> phyloseq_rare
```

#Fix tax table
```{r}
# phylo_C@otu_table
# phylo_C@tax_table
noNA = !is.na(tax_table(phyloseq_rare)[,"Genus"]) & !is.na(tax_table(phyloseq_rare)[,"Species"])
tax_table(phyloseq_rare)[noNA][,"Species"] = paste(tax_table(phyloseq_rare)[noNA][,"Genus"], tax_table(phyloseq_rare)[noNA][,"Species"])
phyloseq_rare%>%
  tax_fix()->phyloseq_rare
noNA = !is.na(tax_table(phyloseq)[,"Genus"]) & !is.na(tax_table(phyloseq)[,"Species"])
tax_table(phyloseq)[noNA][,"Species"] = paste(tax_table(phyloseq)[noNA][,"Genus"], tax_table(phyloseq)[noNA][,"Species"])
phyloseq%>%
  tax_fix()->phyloseq
```

#A) Alpha diversity
#Alpha: From feces to cultures 
```{r}
###Feces to culture
phyloseq_rare%>% 
     subset_samples(Sample_id_compete %in% c("feces_dil_1", "feces") | stress %in% "control")%>%
  phyloseq_alphas(phylo = TRUE) -> alphas

unique(alphas$time)

alphas$time<-factor(alphas$time, levels=c("feces", "feces_dil", "pre-stress"))

aggregate(alphas$observed, by=list(alphas$time), FUN=median)

alphas%>%
  ggplot(aes(x=time, y=observed))+
  theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=20, angle=90),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), strip.text.x = element_blank(), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_boxplot(aes(group=time), size=1)+
  geom_point(aes(colour=donor_name), size=3)+
  ggtitle("Alpha div. in feces & culures")+
  ylim(80, 300)

ggsave(filename = "Alpha_feces to culture.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output",
       width = 17,
       height = 17,
       units = c("cm"))
```
#Alpha: Upon H2O2 stress
```{r}
#H2O2
###Line plot########################################################################################
phyloseq_rare%>% 
     subset_samples(stress %in% c("H2O2", "control"))%>%
  phyloseq_alphas(phylo = TRUE) -> alphas

###ad control value to all stress-level
alphas%>%
  subset(stress %in% "control")->low
low$stress_level <- "low"
alphas%>%
  subset(stress %in% "control")->median
median$stress_level <- "median"
alphas%>%
  subset(stress %in% "control")->high
high$stress_level <- "high"
rbind(alphas, low, median, high)->alphas
#####

alphas$time<-factor(alphas$time, levels=c("pre-stress", "stress", "post_stress_1", "post_stress_2"))
alphas$stress_level<-factor(alphas$stress_level, levels=c("control", "low", "median", "high"))


alphas%>%
  ggplot(aes(x=time, y=observed))+
 theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=20, angle=90),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_line(aes(colour=donor_name, group=donor_name), size=1)+
  facet_grid(cols=vars(stress_level))+
  ggtitle("Alpha diversity")

ggsave(filename = "alpha_h2o2.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/H2O2/alpha/output",
       width = 20,
       height = 14,
       units = c("cm"))
# library(ggh4x)
alphas%>%
  subset(!time %in% c("pre-stress"))%>%
     subset(!(stress_level %in% c("high", "low", "median") & time %in% "pre-stress"))%>%
  ggplot(aes(x=time, y=observed))+
  scale_fill_manual(values=c("steelblue4", "steelblue3",  "coral2", "coral3"))+
 theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_text(size=20),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_boxplot(aes(fill=stress_level), outlier.shape = NA, position = position_dodge2(preserve = "single"))+
  geom_point(aes(colour=donor_name, group=stress_level), position = position_jitterdodge(0.05))+
  ggtitle("Alpha diversity")+
  scale_x_discrete(labels=c("Stress", "Post 1", "Post 2"))+
  ylab("Observed")+
  labs(colour="Donor", fill="Stress level")
   # force_panelsizes(rows = unit(8, "cm"),
   #                 cols = unit(11, "cm"))
  
  # x<-paste(stress_level_cond[i])

ggsave(filename = paste( "alpha_h2o2.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/H2O2/alpha",
       width = 20,
       height = 12,
       units = c("cm"))

#########################################################################################################
##Box plots
#########################################################################################################
stress_level_cond<-c("low"  ,  "median" , "high")
for (i in 1: length(stress_level_cond)){
alphas%>%
    subset(time %in% c("pre-stress"))%>%
    subset(stress_level %in% c(stress_level_cond[i], "control"))%>%
  ggplot(aes(x=time, y=observed))+
 theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20),
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=20, angle=90),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_boxplot(aes(fill=stress_level))+
  geom_point(aes(colour=donor_name, group=stress_level), position = position_jitterdodge(0.05))+
  ggtitle("Alpha diversity")
  
  x<-paste(stress_level_cond[i])

ggsave(filename = paste(x, "alpha_h2o2.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/H2O2/alpha",
       width = 20,
       height = 14,
       units = c("cm"))}

```
#Alpha: upon oxygen stress
```{r}
#O2: aerobic
phyloseq_rare%>% 
     subset_samples( stress %in% "control" | (stress %in% "O2" ))%>%
  phyloseq_alphas(phylo = TRUE) -> alphas

###ad control value to all stress-level (anaerobe)
alphas%>%
  subset(stress %in% "control")->low
low$stress_level <- "low"

alphas%>%
  subset(stress %in% "control")->median
median$stress_level <- "median"

alphas%>%
  subset(stress %in% "control")->high
high$stress_level <- "high"

alphas%>%
  subset(stress %in% "control")->max
max$stress_level <- "max"
###ad control value to all stress-level (aerobe)
alphas%>%
  subset(stress %in% "control")->control
aerobe<-rbind(control, low, median, high, max)
aerobe$incubation<-"aerobe"
#####

alphas <-rbind(alphas, aerobe, low, control, median, high, max)

alphas$time<-factor(alphas$time, levels=c("pre-stress", "stress", "post_stress_1", "post_stress_2"))

alphas$stress_level<-factor(alphas$stress_level, levels=c("control", "low", "median", "high", "max"))

alphas%>%
  ggplot(aes(x=time, y=observed))+
  theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=20, angle=90),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_line(aes(colour=donor_name, group=donor_name), size=1)+
  facet_grid(cols=vars(stress_level, incubation))+
  ggtitle("Alpha div.")+
  ylim(80, 200)

ggsave(filename = "alpha_o2.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/O2/alpha",
       width = 40,
       height = 14,
       units = c("cm"))

stress_level_cond<-c("control", "low"  ,  "median" , "high", "max")

# for (i in 1: length(stress_level_cond)){

alphas$incubation<-factor(alphas$incubation, levels=c("anaerobe", "aerobe"))

names<-list("stress"= "Stress", 
            "post_stress_1"= "Post 1", 
            "post_stress_2"= "Post 2")

facet_labeller <- function(variable,value){
  return(names[value])
}


alphas%>%
  subset(!time %in% "pre-stress")->sub
sub$time <-factor(sub$time, levels=c("stress", "post_stress_1", "post_stress_2"))


sub%>%
  subset(stress_level %in% c("control", "median", "max"))%>%
  # subset(!(incubation %in% c("aerobe") & time %in% "pre-stress"))%>%
    # subset(stress_level %in% c(stress_level_cond[i]))%>%
  ggplot(aes(x=stress_level, y=observed))+
  scale_alpha_manual(values=c(1,0.5), name="Incubation", labels=c("Control", "Stress"))+
  scale_fill_manual(values=c("steelblue4" ,"coral2"  , "coral4"), labels=c("minimal", "median","max"))+
 theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_boxplot(aes(fill=stress_level, alpha=incubation), outlier.shape = NA, position = position_dodge2(preserve = "single"))+
  geom_point(aes(colour=donor_name, group=incubation), position = position_jitterdodge(0.1))+
  ggtitle("Alpha diversity")+
  scale_x_discrete(labels=c("minimal", "low", "median", "high", "max"))+
  ylab("Observed")+
  labs(colour="Donor", fill="Stress level")+
   # force_panelsizes(rows = unit(c(1,2,2), "cm"),
   #                 cols = unit(11, "cm"))+
  facet_grid(cols=vars(time), scale="free", labeller=facet_labeller)+
  guides(alpha = guide_legend(override.aes = list(fill = "darkgrey"))) +
  ylim(40, 160)
  
  x<-paste(stress_level_cond[i])

ggsave(filename = paste(x, "O2_alpha_o2.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/O2/alpha",
       width = 20,
       height = 20,
       units = c("cm"))
# }
```

#B) Beta diversity
#Calculations of distances: feces vs. cultures
```{r}
# library(dplyr)


phyloseq_rare %>%
  subset_samples(!Sample_id_compete %in% c("feces_dil_1"))%>%
  subset_samples(!donor_name %in% NA)-> sub
sub%>%
  phyloseq::distance(method = "bray") -> bc
sub %>%
  ordinate(method = "PCoA",
           distance = bc) -> ord
sub %>% 
  phyloseq_distance_boxplot(p = ., 
                            dist = bc, 
                            d = "ox_nr") -> out # column in metadata for comparaisons

setNames(as.data.frame(out$matrix),c("sample_name", "sample_name_comp", "value", "condition_comp", "comparison")) -> distances

distances%>%
  subset(comparison %in% c("feces"))-> distances

joined<-left_join(distances, sample.data.frame(sub)[,c(1,7:12)])

joined%>%
  mutate(sample_name_comp = recode(sample_name_comp, "M-01" = 'D1',
                                   "M-02" = 'D2',
                                   "M-03" = 'D3',
                                   "M-04" = 'D4',
                                   "M-05" = 'D5',
                                   "M-06" = 'D6',
                                   "M-07" = 'D7',
                                   "M-08" = 'D8',))->joined

for (i in 1:nrow(joined)){
  if (joined[i,c("donor_name")] == joined[i,c("sample_name_comp")]){joined[i,c("match")]= "true"}
}

joined%>%
  subset(match %in% "true")->match

# match$condition_comp<-factor(match$condition_comp, levels=c("control","protein_YCFA"))
```

#Plots
```{r}

###feces to cultures
match%>%
  subset(stress %in% "control")->sub_dist

aggregate(sub_dist$value, by=list(sub_dist$time), FUN=mean)

sub_dist$time<-factor(sub_dist$time, levels=c("pre-stress"))
sub_dist$stress_level<-factor(sub_dist$stress_level, levels=c("control", "low", "median", "high"))
sub_dist%>%
  ggplot(aes(x=time, y=value))+
  theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=20, angle=90),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20),  panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_boxplot()+ geom_point(aes(colour=donor_name), size=3)+
  ggtitle("Beta dist.: feces vs. culures")+
  ylim(0,1)

ggsave(filename = "Beta_dist_culture_feces.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output",
       width = 14,
       height = 17,
       units = c("cm"))
```


#Calculations of distances: control culture vs. subsequent passages
```{r}
phyloseq_rare %>%
  subset_samples(!Sample_id_compete %in% c("feces_dil_1", "feces"))%>%
  subset_samples(!donor_name %in% NA)-> sub
sub%>%
  phyloseq::distance(method = "bray") -> bc
sub %>%
  ordinate(method = "PCoA",
           distance = bc) -> ord
sub %>% 
  phyloseq_distance_boxplot(p = ., 
                            dist = bc, 
                            d = "time") -> out # column in metadata for comparaisons

setNames(as.data.frame(out$matrix),c("sample_name", "sample_name_comp", "value", "condition_comp", "comparison")) -> distances

distances%>%
  subset(comparison %in% c("pre-stress"))-> distances

joined<-left_join(distances, sample.data.frame(sub)[,c(1,7:12)])

joined%>%
  mutate(sample_name_comp = recode(sample_name_comp, "M-09" = 'D1',
                                   "M-10" = 'D2',
                                   "M-11" = 'D3',
                                   "M-12" = 'D4',
                                   "M-13" = 'D5',
                                   "M-14" = 'D6',
                                   "M-15" = 'D7',
                                   "M-16" = 'D8',))->joined

for (i in 1:nrow(joined)){
  if (joined[i,c("donor_name")] == joined[i,c("sample_name_comp")]){joined[i,c("match")]= "true"}
}

joined%>%
  subset(match %in% "true")->match

# match$condition_comp<-factor(match$condition_comp, levels=c("control","protein_YCFA"))
```

#H2O2
```{r}
match%>%
  subset(stress %in% "H2O2" )%>%
  ggplot(aes(x=time, y=value))+
  theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=20, angle=90),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20),  panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_line(aes(colour=stress_level, group=stress_level), size=1)+
  facet_grid( cols=vars(donor_name))+
  geom_point(aes(colour=stress_level, group=stress_level))+
  ggtitle("Beta dist.: pre-stress vs. culures:H2O2")

ggsave(filename = "Beta_dist_H2O2.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/H2O2",
       width = 30,
       height = 17,
       units = c("cm"))



match%>%
  subset(stress %in% "H2O2" )%>%
  ggplot(aes(x=time, y=value))+
  scale_fill_manual(values=c("steelblue4", "steelblue3",  "coral2", "coral3"))+
 theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_text(size=20),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_boxplot(aes(fill=stress_level), outlier.shape = NA, position = position_dodge2(preserve = "single"))+
  geom_point(aes(colour=donor_name, group=stress_level), position = position_jitterdodge(0.05))+
  ggtitle("Bray-Curtis distances")+
  scale_x_discrete(labels=c("Stress", "Post 1", "Post 2"))+
  ylab("Distance")+
  labs(colour="Donor", fill="Stress level") 
  #  force_panelsizes(rows = unit(10, "cm"),
  #                  cols = unit(11, "cm"))

ggsave(filename = paste("beta_h2o2.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/H2O2",
       width = 20,
       height = 10,
       units = c("cm"))
#########################################################################################################
##Box plots
#########################################################################################################
stress_level_cond<-c("low"  ,  "median" , "high")
for (i in 1: length(stress_level_cond)){
match%>%
  subset(stress %in% "H2O2" )%>%
    subset(stress_level %in% c(stress_level_cond[i], "control"))%>%
  ggplot(aes(x=time, y=value))+
 theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=20, angle=90),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_boxplot(aes(fill=stress_level))+
  geom_point(aes(colour=donor_name, group=stress_level), position = position_jitterdodge(0.1))+
  ggtitle("Alpha diversity")
  
  x<-paste(stress_level_cond[i])

ggsave(filename = paste(x, "beta_h2o2.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/H2O2",
       width = 20,
       height = 14,
       units = c("cm"))}


```
#O2
```{r}
match$time<-factor(match$time, levels=c("pre-stress", "stress", "post_stress_1", "post_stress_2"))
match$stress_level<-factor(match$stress_level, levels=c("control", "low", "median", "high", "max"))

match%>%
  subset(stress %in% "O2" )%>%
  ggplot(aes(x=time, y=value))+
  theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=20, angle=90),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
   geom_line(aes(colour=stress_level, group=stress_level), size=1)+
  facet_grid( cols=vars(donor_name, incubation))+
  geom_point(aes(colour=stress_level, group=stress_level))+
  ggtitle("Beta dist.: pre-stress vs. culures:anaerobic controls")+
  ylim(0.25, 0.75)

ggsave(filename = "Beta_dist_O2_anaerobe.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/O2",
       width = 40,
       height = 17,
       units = c("cm"))

match%>%
  subset(stress %in% "O2" )%>%
  ggplot(aes(x=time, y=value))+
  theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=20, angle=90),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
   geom_boxplot(aes(fill=incubation), size=1, outlier.shape = NA)+
  facet_grid(cols=vars(stress_level))+
  geom_point(aes(colour=donor_name, group=incubation), position = position_dodge(1), alpha=0.8, size=3)+
  ggtitle("Beta dist.: pre-stress vs. culures:anaerobic controls")+
  ylim(0.25, 0.75)

ggsave(filename = "Beta_dist_box.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/O2",
       width = 40,
       height = 17,
       units = c("cm"))

```

#Distance between control conditions: H2O2
```{r}

phyloseq_rare %>%
  # subset_samples(time %in% "stress")%>%
  subset_samples(stress %in% c("H2O2"))-> sub
sub%>%
  phyloseq::distance(method = "bray") -> bc
sub %>%
  ordinate(method = "PCoA",
           distance = bc) -> ord
sub %>% 
  phyloseq_distance_boxplot(p = ., 
                            dist = bc, 
                            d = "stress_level") -> out # column in metadata for comparaisons

setNames(as.data.frame(out$matrix),c("sample_name", "sample_name_comp", "value", "condition_comp", "comparison")) -> distances

distances%>%
  subset(comparison %in% c("control"))-> distances

joined<-left_join(distances, sample.data.frame(sub)[,c(1,7:12)])

sample_name<-setNames(sample.data.frame(sub)[,c("sample_name", "donor_name", "time")], c("sample_name_comp", "donor_comp", "time_comp"))


joined_2<-left_join(joined, sample_name)
 

for (i in 1:nrow(joined_2)){
  if (joined_2[i,c("donor_name")] == joined_2[i,c("donor_comp")] & joined_2[i,c("time")] == joined_2[i,c("time_comp")]){joined_2[i,c("match")]= "true"}
}

joined_2%>%
  subset(match == c("true"))-> match_h2o2
```

```{r}
match_h2o2$stress_level<-factor(match_h2o2$stress_level, levels=c("low", "median", "high"))
match_h2o2$time<-factor(match_h2o2$time, levels=c("stress", "post_stress_1", "post_stress_2"))

match_h2o2%>%
  ggplot(aes(x=time, y=value))+
  theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=20, angle=90),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
   geom_boxplot(aes(fill=stress_level), size=1, outlier.shape = NA)+
  geom_point(aes(colour=donor_name, group=stress_level), position = position_dodge(0.8), alpha=0.8, size=3)+
  ggtitle("Beta dist.: control vs. stress")+
  ylim(0,1)

ggsave(filename = "Beta_dist_box.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/H2O2",
       width = 17,
       height = 17,
       units = c("cm"))


match_h2o2%>%
 subset(stress %in% "H2O2" )%>%
  ggplot(aes(x=time, y=value))+
  scale_fill_manual(values=c("steelblue3",  "coral2", "coral3"))+
 theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_text(size=20),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
  geom_boxplot(aes(fill=stress_level), outlier.shape = NA, position = position_dodge2(preserve = "single"))+
  geom_point(aes(colour=donor_name, group=stress_level), position = position_jitterdodge(0.05))+
  ggtitle("Bray-Curtis distances to control")+
  scale_x_discrete(labels=c("Stress", "Post 1", "Post 2"))+
  ylab("Distance")+
  labs(colour="Donor", fill="Stress level") 
   # force_panelsizes(rows = unit(8, "cm"),
   #                 cols = unit(11, "cm"))

ggsave(filename = paste("beta_h2o2.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/H2O2",
       width = 20,
       height = 10,
       units = c("cm"))
```


#Distance between control conditions: O2
```{r}


phyloseq_rare %>%
  # subset_samples(time %in% "stress")%>%
  subset_samples(stress %in% c("O2"))-> sub
sub%>%
  phyloseq::distance(method = "bray") -> bc
sub %>%
  ordinate(method = "PCoA",
           distance = bc) -> ord
sub %>% 
  phyloseq_distance_boxplot(p = ., 
                            dist = bc, 
                            d = "incubation") -> out # column in metadata for comparaisons

setNames(as.data.frame(out$matrix),c("sample_name", "sample_name_comp", "value", "condition_comp", "comparison")) -> distances

distances%>%
  subset(comparison %in% c("anaerobe"))-> distances

joined<-left_join(distances, sample.data.frame(sub)[,c(1,7:12)])

sample_name<-setNames(sample.data.frame(sub)[,c("sample_name", "incubation", "donor_name", "time", "stress_level")], c("sample_name_comp", "incubation_comp","donor_comp", "time_comp", "stress_level_comp"))


joined_2<-left_join(joined, sample_name)
 

for (i in 1:nrow(joined_2)){
  if ((joined_2[i,c("condition_comp")] != joined_2[i,c("incubation_comp")]) & 
      (joined_2[i,c("donor_name")] == joined_2[i,c("donor_comp")]) & 
      (joined_2[i,c("time")] == joined_2[i,c("time_comp")]) &
      (joined_2[i,c("stress_level")] == joined_2[i,c("stress_level_comp")])) {joined_2[i,c("match")]= "true"}
}

joined_2%>%
  subset(match == c("true"))-> match_o2
```
```{r}
match_o2$stress_level<-factor(match_o2$stress_level, levels=c("control","low", "median", "high", "max"))
match_o2$time<-factor(match_o2$time, levels=c("stress", "post_stress_1", "post_stress_2"))

names<-list("stress"= "Stress", 
            "post_stress_1"= "Post 1", 
            "post_stress_2"= "Post 2")

facet_labeller <- function(variable,value){
  return(names[value])
}

match_o2%>%
  subset(stress_level %in% c("control", "median", "max"))%>%
  ggplot(aes(x=time, y=value))+
  theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
   geom_boxplot(aes(fill=stress_level), size=1, outlier.shape = NA)+
  geom_jitter(aes(colour=donor_name, group=stress_level), position = position_jitterdodge(0.05))+
  ggtitle("Bray-Curtis distance to control")+
   scale_fill_manual(values=c("steelblue4","coral2"  , "coral4"), labels=c("minimal",  "median","max"))+
  ylim(0,1)+
  ylab("Distance")+
  labs(colour="Donor", fill="Stress level") +
  facet_grid(cols=vars(time), scale="free", labeller=facet_labeller)
  


ggsave(filename = paste(x, "O2_alpha_o2.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/O2/beta",
       width = 20,
       height = 10,
       units = c("cm"))


```

#PCoA Plots H2O2
```{r}
 # library(plyr)
#########
#BETA
#########
dist = "bray"
# dist = "jaccard"
ord_meths = c("PCoA")

stress_level_cond <-c("low", "median", "high")

for (i in 1:length(stress_level_cond)){
  
phyloseq_rare%>% 
    subset_samples(time %in% "stress")%>%
     subset_samples(stress %in% "H2O2" & stress_level %in% c("control", stress_level_cond[i]))->betas

plist = llply(as.list(ord_meths), function(i, betas, dist){
        ordi = ordinate(betas, method=i, distance=dist)
        plot_ordination(physeq = betas, 
                        ordination = ordi,
                        color = "donor_name", 
                        title = "PCoA Bray-Curtis")},
        betas, 
        dist)


names(plist) <- ord_meths

pdataframe = ldply(plist, function(x){
    df = x$data[, 1:2]
    colnames(df) = c("Axis_1", "Axis_2")
    return(cbind(df, x$data))})


names(pdataframe)[1] = "method"

  

  pdataframe%>%
    # subset(donor_name %in% c("D8", "D7"))%>%
    ggplot(aes(Axis_1, Axis_2))+ 
    scale_shape_manual(values = c(0,1,2,3,4,5,6,7))+
    stat_ellipse( aes(group=stress_level, fill=stress_level),geom = "polygon", lwd=0.01, alpha=0.2, level=0.9 )+
    geom_point(aes(colour=stress_level, shape=donor_name), alpha=0.5, size=3, stroke=2)+
    # ggtitle(title)+
    # scale_fill_manual(values=c("#F8766D", "darkgoldenrod1", "#00BFC4"), name="Fibers")+
    theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_text(size=20),
        axis.text.x = element_text(size=20),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20))+
   xlab("PCoA1") + ylab("PCoA2")+labs()->p
  
 show(p)
 x<-stress_level_cond[i]
 
 ggsave(filename = paste0(x, "beta.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/H2O2/PcoAs",
       width = 17,
       height = 17,
       units = c("cm"))}

```
#PCoA Plots H2O2
```{r}
#########
#BETA
#########
dist = "bray"
# dist = "jaccard"
ord_meths = c("PCoA")

stress_level_cond <-c("control", "low", "median", "high", "max")

for (i in 1:length(stress_level_cond)){
  
phyloseq_rare%>% 
    subset_samples(time %in% "post_stress_1")%>%
     subset_samples(stress %in% "O2" & stress_level %in% c(stress_level_cond[i]))->betas

plist = llply(as.list(ord_meths), function(i, betas, dist){
        ordi = ordinate(betas, method=i, distance=dist)
        plot_ordination(physeq = betas, 
                        ordination = ordi,
                        color = "donor_name", 
                        title = "PCoA Bray-Curtis")},
        betas, 
        dist)


names(plist) <- ord_meths

pdataframe = ldply(plist, function(x){
    df = x$data[, 1:2]
    colnames(df) = c("Axis_1", "Axis_2")
    return(cbind(df, x$data))})


names(pdataframe)[1] = "method"

   x<-stress_level_cond[i]

  pdataframe%>%
    # subset(donor_name %in% c("D8", "D7"))%>%
    ggplot(aes(Axis_1, Axis_2))+ 
    scale_shape_manual(values = c(0,1,2,3,4,5,6,7))+
    stat_ellipse( aes(group=incubation, fill=incubation),geom = "polygon", lwd=0.01, alpha=0.2, level=0.9 )+
    geom_point(aes(colour=incubation, shape=donor_name), alpha=0.5, size=3, stroke=2)+
    ggtitle(paste("stress level: ", x))+
    # scale_fill_manual(values=c("#F8766D", "darkgoldenrod1", "#00BFC4"), name="Fibers")+
    theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_text(size=20),
        axis.text.x = element_text(size=20),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20))+
    geom_line(aes(group=donor_name))+
    # geom_text(aes(label=time))+
   xlab("PCoA1") + ylab("PCoA2")+labs()->p
  
 show(p)

 
 ggsave(filename = paste0(x, "beta.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/O2/PcoAs",
       width = 17,
       height = 17,
       units = c("cm"))}

```





############janina here

#Differtial abunance
```{r, echo=FALSE}
#BiocManager::install("microbiome/mia")

 library(mia)
# library(miaViz)
 library(ALDEx2)
```

```{r}
ox <- makeTreeSummarizedExperimentFromPhyloseq(phyloseq) 
```


```{r}
ox_fam <- ox %>% agglomerateByRank(rank = "Family")
ox_gen <- ox %>% agglomerateByRank(rank = "Genus")
ox_species <- ox %>% agglomerateByRank(rank = "Species") 
```

###Aldex: NO Prevalence filter-> filter later with abundance
###Aldex: NO Prevalence filter-> filter later with abundance
```{r}
oxygen <-ox_gen[ , ox_gen$stress %in% c("O2") ]
stress_level_cond <-c("control", "low","median", "high", "max")
time_cond<-c("stress", "post_stress_1", "post_stress_2")
all_oxygen_genus <-data.frame(matrix(ncol=15, nrow=0))

for (i in (1:length(stress_level_cond))){
  ox_sub <-oxygen[ , oxygen$stress_level %in% c(stress_level_cond[i]) & !oxygen$donor_name %in% c("D1")]
    for (j in (1:length(time_cond))){
  ox_sub_sub <-ox_sub[ , ox_sub$time %in% c(time_cond[j])]
    result_name <- paste(stress_level_cond[i])
   
      x <- aldex.clr(
      reads = assay(ox_sub_sub),
     ox_sub_sub$incubation)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out <- data.frame(x_tt, x_effect)
      
      result <-aldex_out
      result[,14] <- result_name
      result[,15] <- paste(time_cond[j])
      names(result)[6]<-paste0("rab.win.aero")
      names(result)[7]<-paste0("rab.win.anaero")
      result$Family<-row.names(result)
      all_oxygen_genus <- rbind(all_oxygen_genus, result)
      }}
all_oxygen_genus %>%
  filter(we.ep < 0.05)->out

out[,c(1,8,10, 14,15)]
```
```{r}
oxygen <-ox_fam[ , ox_fam$stress %in% c("O2") ]
stress_level_cond <-c("control", "low","median", "high", "max")
time_cond<-c("stress", "post_stress_1", "post_stress_2")
all_oxygen <-data.frame(matrix(ncol=15, nrow=0))

for (i in (1:length(stress_level_cond))){
  ox_sub <-oxygen[ , oxygen$stress_level %in% c(stress_level_cond[i])]
    for (j in (1:length(time_cond))){
  ox_sub_sub <-ox_sub[ , ox_sub$time %in% c(time_cond[j])]
    result_name <- paste(stress_level_cond[i])
   
      x <- aldex.clr(
      reads = assay(ox_sub_sub),
     ox_sub_sub$incubation)

      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out <- data.frame(x_tt, x_effect)
      
      result <-aldex_out
      result[,14] <- result_name
      result[,15] <- paste(time_cond[j])
      names(result)[6]<-paste0("rab.win.aero")
      names(result)[7]<-paste0("rab.win.anaero")
      result$Family<-row.names(result)
      all_oxygen <- rbind(all_oxygen, result)
      }}
all_oxygen %>%
  filter(effect < -1 | effect > 1)->out
out[,c(1,8,10, 14,15)] %>%
  subset (V15 %in% "stress")
```

###Aldex: H2O2
###species level
```{r}
h2o2 <-ox_species[ , ox_species$stress %in% c("H2O2")]

stress_level_cond <-c( "low","median", "high")
time_cond<-c("stress", "post_stress_1", "post_stress_2")
all_h2o2_species <-data.frame(matrix(ncol=15, nrow=0))
j=0 

for (i in (1:length(stress_level_cond))){
  ox_sub <-h2o2[ , h2o2$stress_level %in% c(stress_level_cond[i], "control")]

  for (j in (1:length(time_cond))){

  ox_sub_sub <-ox_sub[ , ox_sub$time %in% c(time_cond[j])]
  
     ox_sub_sub <- subsetByPrevalentTaxa(ox_sub_sub, detection = 0, prevalence=0.25, include_lowest= TRUE)
      result_name <- paste(stress_level_cond[i])

      x <- aldex.clr(
      reads = assay(ox_sub_sub),
     ox_sub_sub$stress_level)
      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out <- data.frame(x_tt, x_effect)
      
      result <-aldex_out
      result[,14] <- result_name
      result[,15] <- time_cond[j]

      names(result)[6]<-paste0("rab.win.aero")
      names(result)[7]<-paste0("rab.win.anaero")
      result$Family<-row.names(result)
      all_h2o2_species <- rbind(all_h2o2_species, result)
        }}
```

```{r}
h2o2 <-ox_gen[ , ox_gen$stress %in% c("H2O2")]

stress_level_cond <-c( "low","median", "high")
time_cond<-c("stress", "post_stress_1", "post_stress_2")
all_h2o2_genus <-data.frame(matrix(ncol=15, nrow=0))
j=0 

for (i in (1:length(stress_level_cond))){
  ox_sub <-h2o2[ , h2o2$stress_level %in% c(stress_level_cond[i], "control") & !(h2o2$donor_name %in% "D1")]

  for (j in (1:length(time_cond))){

  ox_sub_sub <-ox_sub[ , ox_sub$time %in% c(time_cond[j])]
  
      ox_sub_sub <- subsetByPrevalentTaxa(ox_sub_sub, detection = 0, prevalence=0.5, include_lowest= TRUE)
      result_name <- paste(stress_level_cond[i])

      x <- aldex.clr(
      reads = assay(ox_sub_sub),
     ox_sub_sub$stress_level)
      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out <- data.frame(x_tt, x_effect)
      
      result <-aldex_out
      result[,14] <- result_name
      result[,15] <- time_cond[j]

      names(result)[6]<-paste0("rab.win.aero")
      names(result)[7]<-paste0("rab.win.anaero")
      result$Family<-row.names(result)
      all_h2o2_genus <- rbind(all_h2o2_genus, result)
        }}
```
####Family level
```{r}
h2o2 <-ox_fam[ , ox_fam$stress %in% c("H2O2")]
stress_level_cond <-c( "low","median", "high")
time_cond<-c("stress", "post_stress_1", "post_stress_2")
all_h2o2 <-data.frame(matrix(ncol=15, nrow=0))

for (i in (1:length(stress_level_cond))){
  ox_sub <-h2o2[ , h2o2$stress_level %in% c(stress_level_cond[i], "control")]
  
  for (j in (1:length(time_cond))){
  ox_sub_sub <-ox_sub[ , ox_sub$time %in% c(time_cond[j])]
      ox_sub_sub <- subsetByPrevalentTaxa(ox_sub_sub, detection = 0, prevalence=0.5, include_lowest= TRUE)
      result_name <- paste(stress_level_cond[i])

      x <- aldex.clr(
      reads = assay(ox_sub_sub),
     ox_sub_sub$stress_level)
      x_tt <- aldex.ttest(
        x, 
        paired.test = FALSE, 
        verbose = FALSE)
      x_effect <- aldex.effect(x, CI = TRUE, verbose = FALSE)
      aldex_out <- data.frame(x_tt, x_effect)
      
      result <-aldex_out
      result[,14] <- result_name
      result[,15] <- paste(time_cond[j])

      names(result)[6]<-paste0("rab.win.aero")
      names(result)[7]<-paste0("rab.win.anaero")
      result$Family<-row.names(result)
      all_h2o2 <- rbind(all_h2o2, result)
        }}

all_h2o2 %>%
  filter(wi.eBH < 0.05)->out
out[,c(1,8,10, 14,15, 16)]

unloadNamespace("miaViz")
unloadNamespace("mia")
```
###CLR differences
#O2
```{r}
phyloseq%>%
  subset_samples(stress %in% "O2")%>%
  physeq_glom_rename(taxrank = "Family", 
                     speedyseq = TRUE)%>%
   microbiome::transform("clr") %>% 
   psmelt() ->df_abund

clr_O2<-left_join(df_abund, as.data.frame(phyloseq@tax_table))

mean_abund<-setNames(aggregate(clr_O2$Abundance, by=list(clr_O2$Family), FUN = "mean"), c("Family", "abund"))

clr_O2<-left_join(clr_O2, mean_abund)

mean_abund$Family<-as.factor(mean_abund$Family)

clr_O2%>%
  mutate(Family = fct_reorder(Family, clr_O2$abund)) ->hoi

ordered<-unique(as.character(hoi$Family))[1:15]
# clr_drug$treatment<-factor(clr_drug$treatment, levels=c("DMSO", "omeprazole",  "ciprofloxacin","b_5-FU","a_control"))

clr_O2%>%
  subset(Family %in% c(ordered))->clr_O2

clr_diff<-setNames(as.data.frame(matrix(0,0,5)), c("diff", "Family", "stress_level", "donor_name", "time"))

stress_level_cond <-c("control", "low","median", "high", "max")
time_cond<-c("stress", "post_stress_1", "post_stress_2")
donor_cond<-unique(clr_O2$donor_name)
donor_fam<-unique(clr_O2$Family)


i=1
for (q in 1: length(time_cond)){
  clr_O2%>%
    subset(time %in% time_cond[q])->sub1a
  
for (m in 1: length(stress_level_cond)){
  sub1a%>%
    subset(stress_level %in% stress_level_cond[m])->sub1
  
  for (j in 1:length(donor_cond)){
    sub1%>%
    subset(donor_name %in% donor_cond[j])->sub2
    
    for (k in 1:length(donor_fam)){
    sub2%>%
    subset(Family %in% donor_fam[k])->sub3
  
      if (nrow(sub3) > 0){
      
      sub3 %>%
          subset(incubation %in% "aerobe")->aerobe
       sub3 %>%
          subset(incubation %in% "anaerobe")->anaerobe
       
      clr_diff[i, c("diff")] <- aerobe$Abundance - anaerobe$Abundance 
      clr_diff[i, c("Family")] <-donor_fam[k]
      clr_diff[i, c("stress_level")] <-stress_level_cond[m]
      clr_diff[i, c("donor_name")] <-donor_cond[j]
      clr_diff[i, c("time")] <-time_cond[q]
  i=i+1}
    }}}}

clr_diff<-left_join(clr_diff, mean_abund)
stats_o2<-setNames(all_oxygen[, c("we.ep", "V14", "V15", "Family")], c("p", "stress_level", "time", "Family"))
stats_o2$Family<-str_remove(stats_o2$Family, "Family:")

for (i in 1:nrow(stats_o2)){
  if (stats_o2[i,c("p")] <0.05 & stats_o2[i,]$p >0.01){stats_o2[i,c("sig")] = "*"}
  if (stats_o2[i,c("p")] <0.01 & stats_o2[i,]$p >0.001){stats_o2[i,c("sig")] = "**"}
  if (stats_o2[i,c("p")] <0.001 ){stats_o2[i,c("sig")] = "***"}
}

diff_stats_o2<-join(clr_diff, stats_o2)
diff_stats_o2$stress_level<-factor(diff_stats_o2$stress_level, levels=c("control", "low","median", "high", "max"))
diff_stats_o2$time<-factor(diff_stats_o2$time, levels=c("stress", "post_stress_1","post_stress_2"))

diff_stats_o2%>%
  mutate(Family = fct_reorder(Family, diff_stats_o2$abund))%>%
  ggplot(aes(x=diff, y=Family))+
geom_vline(xintercept = 0, linetype = "solid", colour = "grey", size=1)+
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey", size=1)+
  geom_vline(xintercept = -1, linetype = "dashed", colour = "grey", size=1)+
   geom_boxplot(aes(fill=stress_level), outlier.shape = NA)+
  geom_jitter(aes(group=stress_level, colour=donor_name), position = position_jitterdodge(0.1) ,size=1, alpha=0.75)+
   # scale_fill_manual(values = c( "cyan4" , "darkgoldenrod1", "#F8766D"),
   #                      labels = c(  "5-FU","Ciprofloxacin", "Omeprazole"), 
   #                     name="Drug")+
   theme(axis.text.y= element_text(size = 20, face="italic"), 
              axis.title.y = element_text(size=20), 
              axis.title.x = element_text(size=20),
              axis.text.x = element_text(size=20),
              legend.text = element_text(size = 20),
              legend.title = element_text(size= 20),
              plot.title = element_text(size=20),
         strip.text = element_text(size=20),
         panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
      ylab("")+xlab("Differences in clr-abundance")+labs(colour="Donor")+
  facet_grid(cols=vars(time))+
  geom_text(data= subset(diff_stats_o2, donor_name %in% c("D1")), aes(label=sig, group=stress_level), position = position_dodge(width=0.75), size=5, colour="red")
      ggtitle("O2-specifically altered families")
ggsave(filename = paste0("clr_dist.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/O2/clr_diff",
       width = 25,
       height = 30,
       units = c("cm"))

51.7/14*27
65.3/17*30
```
##genus level
```{r}
phyloseq%>%
  subset_samples(stress %in% "O2")%>%
  physeq_glom_rename(taxrank = "Genus", 
                     speedyseq = TRUE)%>%
   microbiome::transform("clr") %>% 
   psmelt() ->df_abund

clr_O2<-left_join(df_abund, as.data.frame(phyloseq@tax_table))

mean_abund<-setNames(aggregate(clr_O2$Abundance, by=list(clr_O2$Genus), FUN = "mean"), c("Genus", "abund"))

clr_O2<-left_join(clr_O2, mean_abund)

mean_abund$Genus<-as.factor(mean_abund$Genus)

clr_O2%>%
  mutate(Genus = fct_reorder(Genus, clr_O2$abund)) ->hoi

ordered<-unique(as.character(hoi$Genus))[1:20]
# clr_drug$treatment<-factor(clr_drug$treatment, levels=c("DMSO", "omeprazole",  "ciprofloxacin","b_5-FU","a_control"))

clr_O2%>%
  subset(Genus %in% c(ordered))->clr_O2

clr_diff<-setNames(as.data.frame(matrix(0,0,5)), c("diff", "Genus", "stress_level", "donor_name", "time"))

stress_level_cond <-c("control", "low","median", "high", "max")
time_cond<-c("stress", "post_stress_1", "post_stress_2")
donor_cond<-unique(clr_O2$donor_name)
donor_fam<-unique(clr_O2$Genus)


i=1
for (q in 1: length(time_cond)){
  clr_O2%>%
    subset(time %in% time_cond[q])->sub1a
  
for (m in 1: length(stress_level_cond)){
  sub1a%>%
    subset(stress_level %in% stress_level_cond[m])->sub1
  
  for (j in 1:length(donor_cond)){
    sub1%>%
    subset(donor_name %in% donor_cond[j])->sub2
    
    for (k in 1:length(donor_fam)){
    sub2%>%
    subset(Genus %in% donor_fam[k])->sub3
  
      if (nrow(sub3) > 0){
      
      sub3 %>%
          subset(incubation %in% "aerobe")->aerobe
       sub3 %>%
          subset(incubation %in% "anaerobe")->anaerobe
       
      clr_diff[i, c("diff")] <- aerobe$Abundance - anaerobe$Abundance 
      clr_diff[i, c("Genus")] <-donor_fam[k]
      clr_diff[i, c("stress_level")] <-stress_level_cond[m]
      clr_diff[i, c("donor_name")] <-donor_cond[j]
      clr_diff[i, c("time")] <-time_cond[q]
  i=i+1}
    }}}}

clr_diff<-left_join(clr_diff, mean_abund)
stats_o2<-setNames(all_oxygen_genus[, c("we.ep", "V14", "V15", "Family")], c("p", "stress_level", "time", "Genus"))
stats_o2$Genus<-str_remove(stats_o2$Genus, "Genus:")

for (i in 1:nrow(stats_o2)){
  if (stats_o2[i,c("p")] <0.05 & stats_o2[i,]$p >0.01){stats_o2[i,c("sig")] = "*"}
  if (stats_o2[i,c("p")] <0.01 & stats_o2[i,]$p >0.001){stats_o2[i,c("sig")] = "**"}
  if (stats_o2[i,c("p")] <0.001 ){stats_o2[i,c("sig")] = "***"}
}

diff_stats_o2<-join(clr_diff, stats_o2)
diff_stats_o2$stress_level<-factor(diff_stats_o2$stress_level, levels=c("control", "low","median", "high", "max"))
diff_stats_o2$time<-factor(diff_stats_o2$time, levels=c("stress", "post_stress_1","post_stress_2"))

diff_stats_o2%>%
  mutate(Genus = fct_reorder(Genus, diff_stats_o2$abund))%>%
  ggplot(aes(x=diff, y=Genus))+
geom_vline(xintercept = 0, linetype = "solid", colour = "grey", size=1)+
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey", size=1)+
  geom_vline(xintercept = -1, linetype = "dashed", colour = "grey", size=1)+
   geom_boxplot(aes(fill=stress_level), outlier.shape = NA)+
  geom_jitter(aes(group=stress_level, colour=donor_name), position = position_jitterdodge(0.1) ,size=1, alpha=0.75)+
   # scale_fill_manual(values = c( "cyan4" , "darkgoldenrod1", "#F8766D"),
   #                      labels = c(  "5-FU","Ciprofloxacin", "Omeprazole"), 
   #                     name="Drug")+
   theme(axis.text.y= element_text(size = 20, face="italic"), 
              axis.title.y = element_text(size=20), 
              axis.title.x = element_text(size=20),
              axis.text.x = element_text(size=20),
              legend.text = element_text(size = 20),
              legend.title = element_text(size= 20),
              plot.title = element_text(size=20),
         strip.text = element_text(size=20),
         panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
      ylab("")+xlab("Differences in clr-abundance")+labs(colour="Donor")+
  facet_grid(cols=vars(time))+
  geom_text(data= subset(diff_stats_o2, donor_name %in% c("D1")), aes(label=sig, group=stress_level), position = position_dodge(width=0.75), size=5, colour="red")+
      ggtitle("O2-specifically altered genera")
ggsave(filename = paste0("clr_dist_genus.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/O2/clr_diff",
       width = 25,
       height = 30,
       units = c("cm"))

51.7/14*27
65.3/17*30

names<-list("stress"= "Stress","post_stress_1"= "Post 1", "post_stress_2"= "Post 2")

facet_labeller <- function(variable,value){
  return(names[value])
}



diff_stats_o2%>%
  subset(stress_level %in% c("control", "median","max")) %>%
  subset(!donor_name %in% c("D1"))->sub

sub %>%
   mutate(Genus = fct_reorder(Genus, sub$abund))%>%
  ggplot(aes(x=diff, y=Genus))+
    scale_fill_manual(values=c("steelblue4", "coral2"  , "coral4"), labels=c("minimal", "median","max"))+
  geom_vline(xintercept = 0, linetype = "solid", colour = "grey", size=1)+
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey", size=1)+
  geom_vline(xintercept = -1, linetype = "dashed", colour = "grey", size=1)+
  geom_boxplot(aes(fill=stress_level), outlier.shape = NA)+
  geom_jitter(aes(group=stress_level, colour=donor_name), position = position_jitterdodge(0.1) ,size=1, alpha=0.75)+
   # scale_fill_manual(values = c( "cyan4" , "darkgoldenrod1", "#F8766D"),
   #                      labels = c(  "5-FU","Ciprofloxacin", "Omeprazole"), 
   #                     name="Drug")+
   theme(axis.text.y= element_text(size = 20, face="italic"), 
              axis.title.y = element_text(size=20), 
              axis.title.x = element_text(size=20),
              axis.text.x = element_text(size=20),
              legend.text = element_text(size = 20),
              legend.title = element_text(size= 20),
              plot.title = element_text(size=20), 
         panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), strip.text = element_text(size=20))+
    ylab("")+xlab("Differences in clr-abundance")+
  labs(colour="Donor", fill= "Stress level")+
 geom_text(data= subset(sub, donor_name %in% c("D7") ), aes(x=diff+2, y=Genus, label=sig, group=stress_level), size=8, colour="black", position = position_dodge2(0.8))+ 
  facet_grid(cols=vars(time), labeller= facet_labeller)+
      ggtitle(expression(O[2]*-specifically~altered~genera))+
  scale_x_continuous(breaks=c(-2.5, 2.5))




ggsave(filename = paste0("clr_dist_genus_without_D1.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/O2/clr_diff",
       width = 28,
       height = 40,
       units = c("cm"))
sub%>%
  subset(donor_name %in% "D5") %>%
  subset(!sig %in% NA)

diff_stats_o2 %>%
  subset(Genus %in% "Blautia")
```


#H2O2
```{r}
phyloseq%>%
  subset_samples(stress %in% "H2O2")%>%
  physeq_glom_rename(taxrank = "Family", 
                     speedyseq = TRUE)%>%
   microbiome::transform("clr") %>% 
   psmelt() ->df_abund

clr_H2O2<-left_join(df_abund, as.data.frame(phyloseq@tax_table))

mean_abund<-setNames(aggregate(clr_H2O2$Abundance, by=list(clr_H2O2$Family), FUN = "mean"), c("Family", "abund"))

clr_H2O2<-left_join(clr_H2O2, mean_abund)

mean_abund$Family<-as.factor(mean_abund$Family)

clr_H2O2%>%
  mutate(Family = fct_reorder(Family, clr_H2O2$abund)) ->hoi

ordered<-unique(as.character(hoi$Family))[1:15]
# clr_drug$treatment<-factor(clr_drug$treatment, levels=c("DMSO", "omeprazole",  "ciprofloxacin","b_5-FU","a_control"))

clr_H2O2%>%
  subset(Family %in% c(ordered))->clr_H2O2

clr_diff_H2O2<-setNames(as.data.frame(matrix(0,0,5)), c("diff", "Family", "stress_level", "donor_name", "time"))

stress_level_cond <-c( "low","median", "high")
time_cond<-c("stress", "post_stress_1", "post_stress_2")
donor_cond<-unique(clr_H2O2$donor_name)
donor_fam<-unique(clr_H2O2$Family)


i=1
for (q in 1: length(time_cond)){
  clr_H2O2%>%
    subset(time %in% time_cond[q])->sub1a
  
for (m in 1: length(stress_level_cond)){
  sub1a%>%
    subset(stress_level %in% c(stress_level_cond[m]))->sub1
  
  for (j in 1:length(donor_cond)){
    sub1%>%
    subset(donor_name %in% donor_cond[j])->sub2
    
    for (k in 1:length(donor_fam)){
    sub2%>%
    subset(Family %in% donor_fam[k])->sub3
  
      if (nrow(sub3) > 0){
      
clr_H2O2%>%
          subset(time %in% time_cond[q] & stress_level %in% c("control") & donor_name %in% donor_cond[j] & Family %in% donor_fam[k])->hoi
        
      clr_diff_H2O2[i, c("diff")] <- sub3$Abundance - hoi$Abundance 
   
      
      clr_diff_H2O2[i, c("Family")] <-donor_fam[k]
      clr_diff_H2O2[i, c("stress_level")] <-stress_level_cond[m]
      clr_diff_H2O2[i, c("donor_name")] <-donor_cond[j]
      clr_diff_H2O2[i, c("time")] <-time_cond[q]
     
  
  i=i+1}
    }}}}


clr_diff_H2O2<-left_join(clr_diff_H2O2, mean_abund)


stats<-setNames(all_h2o2[, c("we.ep", "V14", "V15", "Family")], c("p", "stress_level", "time", "Family"))

stats$Family<-str_remove(stats$Family, "Family:")

for (i in 1:nrow(stats)){
  if (stats[i,c("p")] <0.05 & stats[i,]$p >0.01){stats[i,c("sig")] = "*"}
  if (stats[i,c("p")] <0.01 & stats[i,]$p >0.001){stats[i,c("sig")] = "**"}
  if (stats[i,c("p")] <0.001 ){stats[i,c("sig")] = "***"}
}



diff_stats<-join(clr_diff_H2O2, stats)

diff_stats$stress_level<-factor(diff_stats$stress_level, levels=c("control", "low","median", "high", "max"))
diff_stats$time<-factor(diff_stats$time, levels=c("stress", "post_stress_1","post_stress_2"))
diff_stats%>%
  mutate(Family = fct_reorder(Family, diff_stats$abund))%>%
  ggplot(aes(x=diff, y=Family))+
  geom_vline(xintercept = 0, linetype = "solid", colour = "grey", size=1)+
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey", size=1)+
  geom_vline(xintercept = -1, linetype = "dashed", colour = "grey", size=1)+
   geom_boxplot(aes(fill=stress_level), outlier.shape = NA)+
  geom_jitter(aes(group=stress_level, colour=donor_name), position = position_jitterdodge(0.1) ,size=1, alpha=0.75)+
   # scale_fill_manual(values = c( "cyan4" , "darkgoldenrod1", "#F8766D"),
   #                      labels = c(  "5-FU","Ciprofloxacin", "Omeprazole"), 
   #                     name="Drug")+
   theme(axis.text.y= element_text(size = 20, face="italic"), 
              axis.title.y = element_text(size=20), 
              axis.title.x = element_text(size=20),
              axis.text.x = element_text(size=20),
              legend.text = element_text(size = 20),
              legend.title = element_text(size= 20),
              plot.title = element_text(size=20), 
         panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), strip.text = element_text(size=20))+
      ylab("")+xlab("Differences in clr-abundance")+labs(colour="Donor")+
 geom_text(data= subset(diff_stats, donor_name %in% c("D1")), aes(label=sig, group=stress_level), position = position_dodge(width=0.75), size=5, colour="red")+ facet_grid(cols=vars(time))+
  
      ggtitle("Drug-specifically altered families")

ggsave(filename = paste0("clr_dist.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/H2O2/clr_diff",
       width = 25,
       height = 30,
       units = c("cm"))

51.7/14*27
65.3/17*30
```
## genus
```{r}
phyloseq%>%
  subset_samples(stress %in% "H2O2")%>%
  physeq_glom_rename(taxrank = "Genus", 
                     speedyseq = TRUE)%>%
   microbiome::transform("clr") %>% 
   psmelt() ->df_abund

clr_H2O2<-left_join(df_abund, as.data.frame(phyloseq@tax_table))

mean_abund<-setNames(aggregate(clr_H2O2$Abundance, by=list(clr_H2O2$Genus), FUN = "mean"), c("Genus", "abund"))

clr_H2O2<-left_join(clr_H2O2, mean_abund)

mean_abund$Genus<-as.factor(mean_abund$Genus)

clr_H2O2%>%
  mutate(Genus = fct_reorder(Genus, clr_H2O2$abund)) ->hoi

ordered<-unique(as.character(hoi$Genus))[1:20]
# clr_drug$treatment<-factor(clr_drug$treatment, levels=c("DMSO", "omeprazole",  "ciprofloxacin","b_5-FU","a_control"))

clr_H2O2%>%
  subset(Genus %in% c(ordered))->clr_H2O2

clr_diff_H2O2<-setNames(as.data.frame(matrix(0,0,5)), c("diff", "Genus", "stress_level", "donor_name", "time"))

stress_level_cond <-c( "low","median", "high")
time_cond<-c("stress", "post_stress_1", "post_stress_2")
donor_cond<-unique(clr_H2O2$donor_name)
donor_fam<-unique(clr_H2O2$Genus)


i=1
for (q in 1: length(time_cond)){
  clr_H2O2%>%
    subset(time %in% time_cond[q])->sub1a
  
for (m in 1: length(stress_level_cond)){
  sub1a%>%
    subset(stress_level %in% c(stress_level_cond[m]))->sub1
  
  for (j in 1:length(donor_cond)){
    sub1%>%
    subset(donor_name %in% donor_cond[j])->sub2
    
    for (k in 1:length(donor_fam)){
    sub2%>%
    subset(Genus %in% donor_fam[k])->sub3
  
      if (nrow(sub3) > 0){
      
clr_H2O2%>%
          subset(time %in% time_cond[q] & stress_level %in% c("control") & donor_name %in% donor_cond[j] & Genus %in% donor_fam[k])->hoi
        
      clr_diff_H2O2[i, c("diff")] <- sub3$Abundance - hoi$Abundance 
   
      
      clr_diff_H2O2[i, c("Genus")] <-donor_fam[k]
      clr_diff_H2O2[i, c("stress_level")] <-stress_level_cond[m]
      clr_diff_H2O2[i, c("donor_name")] <-donor_cond[j]
      clr_diff_H2O2[i, c("time")] <-time_cond[q]
     
  
  i=i+1}
    }}}}


clr_diff_H2O2<-left_join(clr_diff_H2O2, mean_abund)


stats<-setNames(all_h2o2_genus[, c("we.ep", "V14", "V15", "Family")], c("p", "stress_level", "time", "Genus"))

stats$Genus<-str_remove(stats$Genus, "Genus:")

for (i in 1:nrow(stats)){
  if (stats[i,c("p")] <0.05 & stats[i,]$p >0.01){stats[i,c("sig")] = "*"}
  if (stats[i,c("p")] <0.01 & stats[i,]$p >0.001){stats[i,c("sig")] = "**"}
  if (stats[i,c("p")] <0.001 ){stats[i,c("sig")] = "***"}
}


names<-list("stress"="Stress","post_stress_1"= "Post 1", "post_stress_2"= "Post 2")

facet_labeller <- function(variable,value){
  return(names[value])
}


diff_stats<-left_join(clr_diff_H2O2, stats)
diff_stats$stress_level<-factor(diff_stats$stress_level, levels=c("control", "low","median", "high", "max"))
diff_stats$time<-factor(diff_stats$time, levels=c("stress", "post_stress_1","post_stress_2"))

diff_stats%>%
  subset(!donor_name %in% "D1")->diff_stats

diff_stats%>%
  mutate(Genus = fct_reorder(Genus, diff_stats$abund))%>%
  ggplot(aes(x=diff, y=Genus))+
   scale_fill_manual(values=c("steelblue3",  "coral2", "coral3"))+
  geom_vline(xintercept = 0, linetype = "solid", colour = "grey", size=1)+
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey", size=1)+
  geom_vline(xintercept = -1, linetype = "dashed", colour = "grey", size=1)+
  geom_boxplot(aes(fill=stress_level), outlier.shape = NA)+
  geom_jitter(aes(group=stress_level, colour=donor_name), position = position_jitterdodge(0.1) ,size=1, alpha=0.75)+
   # scale_fill_manual(values = c( "cyan4" , "darkgoldenrod1", "#F8766D"),
   #                      labels = c(  "5-FU","Ciprofloxacin", "Omeprazole"), 
   #                     name="Drug")+
   theme(axis.text.y= element_text(size = 20, face="italic"), 
              axis.title.y = element_text(size=20), 
              axis.title.x = element_text(size=20),
              axis.text.x = element_text(size=20),
              legend.text = element_text(size = 20),
              legend.title = element_text(size= 20),
              plot.title = element_text(size=20), 
         panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), strip.text = element_text(size=20))+
    ylab("")+xlab("Differences in clr-abundance")+
  labs(colour="Donor", fill= "Stress level")+
 geom_text(data= subset(diff_stats, donor_name %in% c("D6")), aes(y=Genus, x=diff+2, label=sig, group=stress_level), position = position_dodge(width=0.75), size=8, colour="black")+ 
  facet_grid(cols=vars(time), labeller= facet_labeller)+
      ggtitle(expression(H[2]*O-specifically~altered~genera))

ggsave(filename = paste0("clr_dist_genus.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/H2O2/clr_diff",
       width = 28,
       height = 40,
       units = c("cm"))

51.7/14*27
65.3/17*30
```
##Species level
```{r}
phyloseq%>%
  subset_samples(stress %in% "H2O2")%>%
  physeq_glom_rename(taxrank = "Species", 
                     speedyseq = TRUE)%>%
   microbiome::transform("clr") %>% 
   psmelt() ->df_abund

clr_H2O2<-left_join(df_abund, as.data.frame(phyloseq@tax_table))

mean_abund<-setNames(aggregate(clr_H2O2$Abundance, by=list(clr_H2O2$Species), FUN = "mean"), c("Species", "abund"))

clr_H2O2<-left_join(clr_H2O2, mean_abund)

mean_abund$Species<-as.factor(mean_abund$Species)

clr_H2O2%>%
  mutate(Species = fct_reorder(Species, clr_H2O2$abund)) ->hoi

ordered<-unique(as.character(hoi$Species))[1:30]
# clr_drug$treatment<-factor(clr_drug$treatment, levels=c("DMSO", "omeprazole",  "ciprofloxacin","b_5-FU","a_control"))

clr_H2O2%>%
  subset(Species %in% c(ordered))->clr_H2O2

clr_diff_H2O2<-setNames(as.data.frame(matrix(0,0,5)), c("diff", "Species", "stress_level", "donor_name", "time"))

stress_level_cond <-c( "low","median", "high")
time_cond<-c("stress", "post_stress_1", "post_stress_2")
donor_cond<-unique(clr_H2O2$donor_name)
donor_fam<-unique(clr_H2O2$Species)


i=1
for (q in 1: length(time_cond)){
  clr_H2O2%>%
    subset(time %in% time_cond[q])->sub1a
  
for (m in 1: length(stress_level_cond)){
  sub1a%>%
    subset(stress_level %in% c(stress_level_cond[m]))->sub1
  
  for (j in 1:length(donor_cond)){
    sub1%>%
    subset(donor_name %in% donor_cond[j])->sub2
    
    for (k in 1:length(donor_fam)){
    sub2%>%
    subset(Species %in% donor_fam[k])->sub3
  
      if (nrow(sub3) > 0){
      
clr_H2O2%>%
          subset(time %in% time_cond[q] & stress_level %in% c("control") & donor_name %in% donor_cond[j] & Species %in% donor_fam[k])->hoi
        
      clr_diff_H2O2[i, c("diff")] <- sub3$Abundance - hoi$Abundance 
   
      
      clr_diff_H2O2[i, c("Species")] <-donor_fam[k]
      clr_diff_H2O2[i, c("stress_level")] <-stress_level_cond[m]
      clr_diff_H2O2[i, c("donor_name")] <-donor_cond[j]
      clr_diff_H2O2[i, c("time")] <-time_cond[q]
     
  
  i=i+1}
    }}}}


clr_diff_H2O2<-left_join(clr_diff_H2O2, mean_abund)


stats<-setNames(all_h2o2_species[, c("we.ep", "V14", "V15", "Family")], c("p", "stress_level", "time", "Species"))

stats$Species<-str_remove(stats$Species, "Species:")

for (i in 1:nrow(stats)){
  if (stats[i,c("p")] <0.05 & stats[i,]$p >0.01){stats[i,c("sig")] = "*"}
  if (stats[i,c("p")] <0.01 & stats[i,]$p >0.001){stats[i,c("sig")] = "**"}
  if (stats[i,c("p")] <0.001 ){stats[i,c("sig")] = "***"}
}



diff_stats<-join(clr_diff_H2O2, stats)

diff_stats$stress_level<-factor(diff_stats$stress_level, levels=c("control", "low","median", "high", "max"))
diff_stats$time<-factor(diff_stats$time, levels=c("stress", "post_stress_1","post_stress_2"))
diff_stats%>%
  mutate(Species = fct_reorder(Species, diff_stats$abund))%>%
  ggplot(aes(x=diff, y=Species))+
  geom_vline(xintercept = 0, linetype = "solid", colour = "grey", size=1)+
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey", size=1)+
  geom_vline(xintercept = -1, linetype = "dashed", colour = "grey", size=1)+
   geom_boxplot(aes(fill=stress_level), outlier.shape = NA)+
  geom_jitter(aes(group=stress_level, colour=donor_name), position = position_jitterdodge(0.1) ,size=1, alpha=0.75)+
   # scale_fill_manual(values = c( "cyan4" , "darkgoldenrod1", "#F8766D"),
   #                      labels = c(  "5-FU","Ciprofloxacin", "Omeprazole"), 
   #                     name="Drug")+
   theme(axis.text.y= element_text(size = 20, face="italic"), 
              axis.title.y = element_text(size=20), 
              axis.title.x = element_text(size=20),
              axis.text.x = element_text(size=20),
              legend.text = element_text(size = 20),
              legend.title = element_text(size= 20),
              plot.title = element_text(size=20), 
         panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), strip.text = element_text(size=20))+
      ylab("")+xlab("Differences in clr-abundance")+labs(colour="Donor")+
 geom_text(data= subset(diff_stats, donor_name %in% c("D1")), aes(label=sig, group=stress_level), position = position_dodge(width=0.75), size=5, colour="red")+ facet_grid(cols=vars(time))+
  
      ggtitle("Drug-specifically altered species")

ggsave(filename = paste0("clr_dist_species.jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/H2O2/clr_diff",
       width = 40,
       height = 60,
       units = c("cm"))

51.7/14*27
65.3/17*30
```


##Species level
```{r}


tax_level<- c("Bacteroides",  "Blautia", "Anaerostipes", "Ruminococcus", "Faecalibacterium",   "Roseburia", "Parabacteroides", "Phascolarctobacterium" , "Prevotella")

for (t in 1:length(tax_level)){
phyloseq%>%
  subset_samples(stress %in% "H2O2")%>%
  physeq_glom_rename(taxrank = "Species", 
                     speedyseq = TRUE)%>%
   microbiome::transform("clr") %>% 
   psmelt() ->df_abund

clr_H2O2<-left_join(df_abund, as.data.frame(phyloseq@tax_table))
clr_H2O2%>%
  subset(Genus %in% c(tax_level[t]))->clr_H2O2

mean_abund<-setNames(aggregate(clr_H2O2$Abundance, by=list(clr_H2O2$Species), FUN = "mean"), c("Species", "abund"))
clr_H2O2<-left_join(clr_H2O2, mean_abund)


mean_abund$Species<-as.factor(mean_abund$Species)

clr_H2O2%>%
  subset(Genus %in% (tax_level[t]))%>%
  mutate(Species = fct_reorder(Species, clr_H2O2$abund)) ->hoi

ordered<-unique(as.character(hoi$Species))[1:30]
# clr_drug$treatment<-factor(clr_drug$treatment, levels=c("DMSO", "omeprazole",  "ciprofloxacin","b_5-FU","a_control"))

clr_H2O2%>%
  subset(Species %in% c(ordered))->clr_H2O2

clr_diff_H2O2<-setNames(as.data.frame(matrix(0,0,5)), c("diff", "Species", "stress_level", "donor_name", "time"))

stress_level_cond <-c( "low","median", "high")
time_cond<-c("stress", "post_stress_1", "post_stress_2")
donor_cond<-unique(clr_H2O2$donor_name)
donor_fam<-unique(clr_H2O2$Species)


i=1
for (q in 1: length(time_cond)){
  clr_H2O2%>%
    subset(time %in% time_cond[q])->sub1a
  
for (m in 1: length(stress_level_cond)){
  sub1a%>%
    subset(stress_level %in% c(stress_level_cond[m]))->sub1
  
  for (j in 1:length(donor_cond)){
    sub1%>%
    subset(donor_name %in% donor_cond[j])->sub2
    
    for (k in 1:length(donor_fam)){
    sub2%>%
    subset(Species %in% donor_fam[k])->sub3
  
      if (nrow(sub3) > 0){
      
clr_H2O2%>%
          subset(time %in% time_cond[q] & stress_level %in% c("control") & donor_name %in% donor_cond[j] & Species %in% donor_fam[k])->hoi
        
      clr_diff_H2O2[i, c("diff")] <- sub3$Abundance - hoi$Abundance 
   
      
      clr_diff_H2O2[i, c("Species")] <-donor_fam[k]
      clr_diff_H2O2[i, c("stress_level")] <-stress_level_cond[m]
      clr_diff_H2O2[i, c("donor_name")] <-donor_cond[j]
      clr_diff_H2O2[i, c("time")] <-time_cond[q]
     
  
  i=i+1}
    }}}}


clr_diff_H2O2<-left_join(clr_diff_H2O2, mean_abund)

clr_diff_H2O2%>%
  subset(!Species %in% paste(tax_level[t], "Genus"))->clr_diff_H2O2

stats<-setNames(all_h2o2_species[, c("we.ep", "V14", "V15", "Family")], c("p", "stress_level", "time", "Species"))

stats$Species<-str_remove(stats$Species, "Species:")

for (i in 1:nrow(stats)){
  if (stats[i,c("p")] <0.05 & stats[i,]$p >0.01){stats[i,c("sig")] = "*"}
  if (stats[i,c("p")] <0.01 & stats[i,]$p >0.001){stats[i,c("sig")] = "**"}
  if (stats[i,c("p")] <0.001 ){stats[i,c("sig")] = "***"}
}

x<-tax_level[t]

diff_stats<-join(clr_diff_H2O2, stats)
diff_stats$stress_level<-factor(diff_stats$stress_level, levels=c("control", "low","median", "high", "max"))
diff_stats$time<-factor(diff_stats$time, levels=c("stress", "post_stress_1","post_stress_2"))


  
diff_stats%>%  
  mutate(Species = fct_reorder(Species, diff_stats$abund))%>%
  ggplot(aes(x=diff, y=Species))+
  geom_vline(xintercept = 0, linetype = "solid", colour = "grey", size=1)+
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey", size=1)+
  geom_vline(xintercept = -1, linetype = "dashed", colour = "grey", size=1)+
   geom_boxplot(aes(fill=stress_level), outlier.shape = NA)+
  geom_jitter(aes(group=stress_level, colour=donor_name), position = position_jitterdodge(0.1) ,size=1, alpha=0.75)+
   # scale_fill_manual(values = c( "cyan4" , "darkgoldenrod1", "#F8766D"),
   #                      labels = c(  "5-FU","Ciprofloxacin", "Omeprazole"), 
   #                     name="Drug")+
   theme(axis.text.y= element_text(size = 20, face="italic"), 
              axis.title.y = element_text(size=20), 
              axis.title.x = element_text(size=20),
              axis.text.x = element_text(size=20),
              legend.text = element_text(size = 20),
              legend.title = element_text(size= 20),
              plot.title = element_text(size=20), 
         panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), strip.text = element_text(size=20))+
      ylab("")+xlab("Differences in clr-abundance")+labs(colour="Donor")+
 geom_text(data= subset(diff_stats, donor_name %in% c("D1")), aes(label=sig, group=stress_level), position = position_dodge(width=0.75), size=5, colour="red")+ facet_grid(cols=vars(time))+
  
      ggtitle(paste("H2O2-specifically altered", x))

ggsave(filename = paste0(x, ".jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/H2O2/clr_diff/genus",
       width = 25,
       height = 30,
       units = c("cm"))}

51.7/14*27
65.3/17*30
```












```{r}
phyloseq_rare %>%
  physeq_glom_rename(taxrank = "Family", 
                     speedyseq = TRUE)%>%
   microbiome::transform("compositional") %>% 
   psmelt() ->df_abund

clr_c<-left_join(df_abund, as.data.frame(phyloseq_rare@tax_table))

clr_c%>%
  subset(Family %in% c("Enterobacteriaceae", "Streptococcaceae"))->clr_entero



clr_entero%>%
  subset(time %in% c("feces", "feces_dil", "pre-stress"))%>%
  ggplot(aes(y=Abundance, x=time))+
  geom_boxplot(aes(fill=Family), outlier.shape = NA)+
  geom_point(aes(colour=donor_name, group=Family), position=position_dodge(1))+ 
  theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=20, angle=90),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), strip.text.x = element_blank(), panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA), legend.text.align = 0)+
 ggtitle("Enterobacteriaceae levels")

ggsave(filename = "aerobes.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output",
       width = 17,
       height = 17,
       units = c("cm"))


clr_c%>%
  subset(Family %in% c("Streptococcaceae"))->clr_entero


clr_entero%>%
  subset(stress %in% "O2")%>%
  ggplot(aes(y=Abundance, x=time))+
  geom_boxplot(aes(fill=stress_level))+
  geom_point(aes(colour=stress_level, group=stress_level), position=position_dodge(1))+
  facet_grid(cols=vars(incubation))+ggtitle("Streptococcaceae")


clr_c%>%
  subset(Family %in% c("Bacteroidaceae"))->clr_entero

clr_entero%>%
  subset(stress %in% "O2")%>%
  ggplot(aes(y=Abundance, x=time))+
  geom_boxplot(aes(fill=stress_level))+
  geom_point(aes(colour=stress_level, group=stress_level), position=position_dodge(1))+
  facet_grid(cols=vars(incubation))+ggtitle("Bacteroidaceae")

phyloseq_rare %>%
  physeq_glom_rename(taxrank = "Genus", 
                     speedyseq = TRUE)%>%
   microbiome::transform("clr") %>% 
   psmelt() ->df_abund

clr_c<-left_join(df_abund, as.data.frame(phyloseq_rare@tax_table))

clr_c%>%
  subset(Genus %in% c("Faecalibacterium"))->clr_entero


clr_entero%>%
  subset(stress %in% "O2")%>%
  ggplot(aes(y=Abundance, x=time))+
  geom_boxplot(aes(fill=stress_level))+
  geom_point(aes(colour=stress_level, group=stress_level), position=position_dodge(1))+
  facet_grid(cols=vars(incubation))+ggtitle("Faecalibacterium")
```

```{r}

donors_loop<-c("D1", "D2", "D3", "D4", "D5", "D6", "D7", "D8")

donor_comp<-c("D1", "D2", "D3", "D4", "D5", "D6", "D7", "D8")


for (i in 1:length(donors_loop)){

phyloseq_rare%>% 
    subset_samples(donor_name %in% donors_loop[i])->sub
  
  
  
  
sub%>%
  # ps_arrange("time")%>%
  microViz::comp_barplot(
    tax_level = "Genus",
    label = "ox_nr",
    n_taxa = 30,
    bar_width = 0.9,
    sample_order = "bray",
    tax_transform_for_plot = "compositional") +
    ylab("Proportion") + xlab( "")-> p_hist


p_hist + 
  facet_grid(cols=vars(stress_level), drop = TRUE, scales = "free_x")+
  theme(axis.text.x = element_text(size=20, angle=90, vjust=0.2, hjust=1, margin=margin(-8,0,0,0)),
        axis.ticks.x = element_blank(),strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_text(size=20),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.spacing = unit(0.2, "cm")) + scale_y_continuous(breaks=c(0.25, 0.5, 0.75))+ ggtitle("Taxonomic distribution and ASV representation in cultures and feces")-> p_hist

p_hist


x<-paste0(donors_loop[i],"H2O2_stress_box.jpeg")
ggsave(filename = x,
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/all_bar",
       width = 80,
       height = 30,
       units = c("cm"))

}
for (i in 1: length(donor_comp)){
  for (j in 1:length(donor_comp)){
  
phyloseq_rare%>% 
  subset_samples(donor_name %in% c(donor_comp[i], donor_comp[j]))->betas

plist = llply(as.list(ord_meths), function(i, betas, dist){
        ordi = ordinate(betas, method=i, distance=dist)
        plot_ordination(physeq = betas, 
                        ordination = ordi,
                        color = "donor_name", 
                        title = "PCoA Bray-Curtis")},
        betas, 
        dist)


names(plist) <- ord_meths

pdataframe = ldply(plist, function(x){
    df = x$data[, 1:2]
    colnames(df) = c("Axis_1", "Axis_2")
    return(cbind(df, x$data))})


names(pdataframe)[1] = "method"

  

  pdataframe%>%
    # subset(donor_name %in% c("D8", "D3", "D1"))%>%
    ggplot(aes(Axis_1, Axis_2))+ 
     scale_shape_manual(values = c(21, 24))+
    stat_ellipse( aes(group=donor_name,  fill=donor_name), geom = "polygon", lwd=0.01, alpha=0.2, level=0.9 )+
    geom_point(aes(fill=donor_name, shape=culture), alpha=1, size=5, stroke=2)+
    ggtitle(title)+
    # scale_fill_manual(values=c("#F8766D", "darkgoldenrod1", "#00BFC4"), name="Fibers")+
    theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_text(size=20),
        axis.text.x = element_text(size=20),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20))+
   xlab("PCoA1") + ylab("PCoA2")+labs(shape="Fibers", colour="Donor")+
    geom_text(aes(label=ox_nr, colour=donor_name))->p
  
 show(p)
 x<-paste(donor_comp[i], donor_comp[j])
 
 ggsave(filename = paste(x,".jpeg"),
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/all_beta",
       width = 80,
       height = 80,
       units = c("cm"))}}

```

```{r}


meta<-sample.data.frame(phyloseq_rare)
# 
# #recode variables: example
# meta$time <-factor(meta$time, levels=c("pre-stress", "stress",
#                                                  "post_stress_1",
#                                                  "post_stress_2" 
#                                                  ))
# 
# 
# 
# #update the metadata file
# sample_data(phyloseq_rare)<-meta


phyloseq_rare%>% 
  subset_samples(donor_name %in% "D7")%>%
    subset_samples(Sample_id_compete %in% c("feces_dil_1", "feces") | stress %in% "control" )%>%
                   # | (stress %in% "H2O2" & stress_level %in% "control") | (stress %in% "O2" & stress_level %in% "control" & incubation %in% "anaerobe")
                   
  ps_arrange(stress_level)%>%
  microViz::comp_barplot(
    tax_level = "Strain",
    label = "time",
    n_taxa =30,
    bar_width = 0.9,
    sample_order = "default",
    tax_transform_for_plot = "compositional") +
    ylab("Proportion") + xlab( "")-> p_hist


p_hist + 
  facet_grid(cols=vars(donor_name), scales = "free", space = "free_x", drop = TRUE)+
  theme(axis.text.x = element_text(size=20, angle=90, vjust=0.2, hjust=1, margin=margin(-8,0,0,0)),
        axis.ticks.x = element_blank(),strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_text(size=20),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.spacing = unit(0.2, "cm")) + scale_y_continuous(breaks=c(0.25, 0.5, 0.75))+ ggtitle("Taxonomic distribution in cultures and feces")-> p_hist

p_hist



ggsave(filename = "feces_cultures.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output",
       width = 45,
       height = 14,
       units = c("cm"))


##########
#ALPHA
#########

phyloseq_rare%>% 
     subset_samples(Sample_id_compete %in% c("feces_dil_1", "feces") | stress %in% "control" | (stress %in% "H2O2" & stress_level %in% "control") | (stress %in% "O2" & stress_level %in% "control" & incubation %in% "anaerobe"))%>%
  phyloseq_alphas(phylo = TRUE) -> alphas

alphas%>%
  # gather(index, alpha_value, "observed":"evenness_pielou")%>%
  # mutate(condition = fct_reorder(factor(condition), mean_value_observed, .desc = T))%>%
  ggplot(aes(x = donor_name,
             y = observed)) +
  # facet_grid(scale="free", space="free_x", cols=vars(sample_type)) +
  geom_boxplot(aes(fill=culture), outlier.shape = NA) +
  geom_point(aes(colour=stress, group=culture), alpha=0.75, position=position_dodge(1), size = 3)+
  ylab("Observed ASVs") + xlab(NULL) +
  # scale_colour_manual(values=c("grey",  'darkgoldenrod1'))+
  theme(axis.text.y= element_text(size = 20), 
            axis.title.y = element_text(size=20), 
            axis.title.x = element_text(size=20),
            axis.text.x = element_text(size=20, angle=90, vjust=0.2, hjust=1),
            legend.text = element_text(size = 20),
            legend.title = element_text(size= 20),
            plot.title = element_text(size=20),
          strip.text.x = element_blank(), 
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
  ggtitle(title)+
  # stat_pvalue_manual(stat.test,  label = "p.adj.signif", label.size = 8)+
  labs(colour="Donor")->alpha

show(alpha)


#########
#BETA
#########
dist = "bray"
# dist = "jaccard"
ord_meths = c("PCoA")


phyloseq_rare%>% 
     subset_samples(Sample_id_compete %in% c("feces_dil_1", "feces") | stress %in% "control"| (stress %in% "H2O2" & stress_level %in% "control") | (stress %in% "O2" & stress_level %in% "control" & incubation %in% "anaerobe"))->betas

plist = llply(as.list(ord_meths), function(i, betas, dist){
        ordi = ordinate(betas, method=i, distance=dist)
        plot_ordination(physeq = betas, 
                        ordination = ordi,
                        color = "donor_name", 
                        title = "PCoA Bray-Curtis")},
        betas, 
        dist)


names(plist) <- ord_meths

pdataframe = ldply(plist, function(x){
    df = x$data[, 1:2]
    colnames(df) = c("Axis_1", "Axis_2")
    return(cbind(df, x$data))})


names(pdataframe)[1] = "method"

  

  pdataframe%>%
    # subset(donor_name %in% c("D8", "D7"))%>%
    ggplot(aes(Axis_1, Axis_2))+ 
     # scale_shape_manual(values = c(21, 24), labels=c(c('feces' = 'feces', '7C' = '6C+Muc', '3C+Muc' = '3C+Muc', "7C-Muc"="6C", "3C"="3C", "H2O"="C-depleted", "Glc"="Glc")))+
    stat_ellipse( aes(group=donor_name, fill=donor_name),geom = "polygon", lwd=0.01, alpha=0.2, level=0.9 )+
    geom_point(aes(colour=donor_name, shape=culture), alpha=0.5, size=3, stroke=2)+
    ggtitle(title)+
    # scale_fill_manual(values=c("#F8766D", "darkgoldenrod1", "#00BFC4"), name="Fibers")+
    theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_text(size=20),
        axis.text.x = element_text(size=20),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20))+
   xlab("PCoA1") + ylab("PCoA2")+labs(shape="Sample type", colour="Donor")->p
  
 show(p)
 
 ggsave(filename = "controls_beta.jpeg",
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output",
       width = 17,
       height = 17,
       units = c("cm"))

```



```{r}

donors_loop<-c("D1", "D2", "D3", "D4", "D5", "D6", "D7", "D8")




for (i in 1:length(donors_loop)){

phyloseq_rare%>% 
    subset_samples(donor_name %in% donors_loop[i])%>%
     subset_samples(stress %in% c("H2O2"))->sub
  
  
  
meta<-sample.data.frame(sub)

#recode variables: example
meta$time <-factor(meta$time, levels=c( "stress",
                                                 "post_stress_1",
                                               "post_stress_2" 
                                                 ))

# print(meta$Sam) 
# sample_data(sub)<-meta


  
sub%>%
  # ps_arrange("time")%>%
  microViz::comp_barplot(
    tax_level = "Genus",
    label = "Sample_id_compete",
    n_taxa = 20,
    bar_width = 0.9,
    sample_order = "default",
    tax_transform_for_plot = "compositional") +
    ylab("Proportion") + xlab( "")-> p_hist


p_hist + 
  facet_grid(cols=vars(stress_level), drop = TRUE, scales = "free_x")+
  theme(axis.text.x = element_text(size=20, angle=90, vjust=0.2, hjust=1, margin=margin(-8,0,0,0)),
        axis.ticks.x = element_blank(),strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_text(size=20),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.spacing = unit(0.2, "cm")) + scale_y_continuous(breaks=c(0.25, 0.5, 0.75))+ ggtitle("Taxonomic distribution and ASV representation in cultures and feces")-> p_hist

p_hist


x<-paste0(donors_loop[i],"H2O2_stress_box.jpeg")
ggsave(filename = x,
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/H2O2_bar",
       width = 50,
       height = 30,
       units = c("cm"))

}



```

```{r}

donors_loop<-c("D1", "D2", "D3", "D4", "D5", "D6", "D7", "D8")



for (i in 1:length(donors_loop)){

phyloseq_rare%>% 
    subset_samples(donor_name %in% donors_loop[i])%>%
  subset_samples(stress %in% c("O2", "control"))%>%
  subset_samples(incubation %in% "aerobe")%>%
  ps_arrange(stress_level)%>%
  microViz::comp_barplot(
    tax_level = "Genus",
    label = "sample_name",
    n_taxa = 20,
    bar_width = 0.9,
    sample_order = "default",
    tax_transform_for_plot = "compositional") +
    ylab("Proportion") + xlab( "")-> p_hist


p_hist + 
  facet_grid(cols=vars(stress_level), scales = "free", space = "free_x", drop = TRUE)+
  theme(axis.text.x = element_text(size=20, angle=90, vjust=0.2, hjust=1, margin=margin(-8,0,0,0)),
        axis.ticks.x = element_blank(),strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_text(size=20),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.spacing = unit(0.2, "cm")) + scale_y_continuous(breaks=c(0.25, 0.5, 0.75))+ ggtitle("Taxonomic distribution and ASV representation in cultures and feces")-> p_hist

p_hist


x<-paste0(donors_loop[i],"O2_stress_box.jpeg")
ggsave(filename = x,
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output",
       width = 50,
       height = 30,
       units = c("cm"))

}


for (i in 1:length(donors_loop)){

phyloseq_rare%>% 
    subset_samples(donor_name %in% donors_loop[i])%>%
  subset_samples(stress %in% c("O2", "control"))%>%
  subset_samples(incubation %in% "anaerobe")%>%
  ps_arrange(stress_level)%>%
  microViz::comp_barplot(
    tax_level = "Genus",
    label = "sample_name",
    n_taxa = 20,
    bar_width = 0.9,
    sample_order = "default",
    tax_transform_for_plot = "compositional") +
    ylab("Proportion") + xlab( "")-> p_hist


p_hist + 
  facet_grid(cols=vars(stress_level), scales = "free", space = "free_x", drop = TRUE)+
  theme(axis.text.x = element_text(size=20, angle=90, vjust=0.2, hjust=1, margin=margin(-8,0,0,0)),
        axis.ticks.x = element_blank(),strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_text(size=20),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.spacing = unit(0.2, "cm")) + scale_y_continuous(breaks=c(0.25, 0.5, 0.75))+ ggtitle("Taxonomic distribution and ASV representation in cultures and feces")-> p_hist

p_hist


x<-paste0(donors_loop[i],"O2_stress_box_ana.jpeg")
ggsave(filename = x,
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output",
       width = 50,
       height = 30,
       units = c("cm"))

}

```

#H2O2

```{r}
phyloseq_rare%>% 
  subset_samples(time %in% c("stress",
                                                 "post_stress_1","post_stress_2"))%>%
     subset_samples( stress %in% c("control", "H2O2"))%>%
  phyloseq_alphas(phylo = TRUE) -> alphas

alphas$time<-factor(alphas$time, levels = c("stress",
                                                 "post_stress_1","post_stress_2" ))
alphas$stress_level<-factor(alphas$stress_level, levels=c("control", "low", "median", "high"))                                              

alphas%>%
  # gather(index, alpha_value, "observed":"evenness_pielou")%>%
  # mutate(condition = fct_reorder(factor(condition), mean_value_observed, .desc = T))%>%
  ggplot(aes(x = time,
             y = observed)) +
 facet_grid(scale="free", space="free_x", cols=vars(stress_level)) +
  geom_boxplot(aes(fill=stress_level), outlier.shape = NA, position = position_dodge(preserve = "single")) +
  geom_point(aes(colour=donor_name, group=stress_level), position = position_dodge(width = 0.75), alpha=0.75, size = 3)+
  ylab("Observed ASVs") + xlab(NULL) +
  # scale_colour_manual(values=c("grey",  'darkgoldenrod1'))+
  theme(axis.text.y= element_text(size = 20), 
            axis.title.y = element_text(size=20), 
            axis.title.x = element_text(size=20),
            axis.text.x = element_text(size=20, angle=90, vjust=0.2, hjust=1),
            legend.text = element_text(size = 20),
            legend.title = element_text(size= 20),
            plot.title = element_text(size=20),
          strip.text.x = element_text(size=20), 
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
  ggtitle(title)+
  # stat_pvalue_manual(stat.test,  label = "p.adj.signif", label.size = 8)+
  labs(colour="Donor")->alpha

show(alpha)

```

```{r}
phyloseq_rare%>% 
  subset_samples(time %in% c("stress","post_stress_1","post_stress_2"))%>%
     subset_samples( stress %in% c("control", "O2"))%>%
  phyloseq_alphas(phylo = TRUE) -> alphas

alphas$time<-factor(alphas$time, levels = c("stress",
                                                 "post_stress_1","post_stress_2" ))
alphas$stress_level<-factor(alphas$stress_level, levels=c("control", "low", "median", "high", "max"))                                              

alphas%>%
  # gather(index, alpha_value, "observed":"evenness_pielou")%>%
  # mutate(condition = fct_reorder(factor(condition), mean_value_observed, .desc = T))%>%
  ggplot(aes(x = time,
             y = observed)) +
  facet_grid(scale="free", space="free_x", cols=vars(stress_level)) +
  geom_boxplot(aes(fill=incubation), outlier.shape = NA, position = position_dodge(preserve = "single")) +
  geom_point(aes(colour=donor_name, group=incubation), position = position_dodge(width = 0.75), alpha=0.75, size = 3)+
  ylab("Observed ASVs") + xlab(NULL) +
  # scale_colour_manual(values=c("grey",  'darkgoldenrod1'))+
  # facet_grid(rows=vars(stress_level))%>%
  theme(axis.text.y= element_text(size = 20), 
            axis.title.y = element_text(size=20), 
            axis.title.x = element_text(size=20),
            axis.text.x = element_text(size=20, angle=90, vjust=0.2, hjust=1),
            legend.text = element_text(size = 20),
            legend.title = element_text(size= 20),
            plot.title = element_text(size=20),
          strip.text.x = element_text(size=20), 
        panel.background = element_rect(fill="white"),
        panel.border=element_rect(fill=NA))+
  ggtitle(title)+
  # stat_pvalue_manual(stat.test,  label = "p.adj.signif", label.size = 8)+
  labs(colour="Donor")->alpha

show(alpha)

```

```{r}
#########
#BETA
#########
dist = "bray"
# dist = "jaccard"
ord_meths = c("PCoA")


phyloseq_rare%>% 
   subset_samples(time %in% c("stress",
                                                 "post_stress_1","post_stress_2"))%>%
     subset_samples( stress %in% c("control", "H2O2"))->betas

plist = llply(as.list(ord_meths), function(i, betas, dist){
        ordi = ordinate(betas, method=i, distance=dist)
        plot_ordination(physeq = betas, 
                        ordination = ordi,
                        color = "donor_name", 
                        title = "PCoA Bray-Curtis")},
        betas, 
        dist)


names(plist) <- ord_meths

pdataframe = ldply(plist, function(x){
    df = x$data[, 1:2]
    colnames(df) = c("Axis_1", "Axis_2")
    return(cbind(df, x$data))})


names(pdataframe)[1] = "method"

pdataframe$stress_level<-factor(pdataframe$stress_level, levels=c("control", "low", "median", "high", "max"))  

  pdataframe%>%
    ggplot(aes(Axis_1, Axis_2))+ 
     scale_shape_manual(values = c(21, 24, 23), labels=c(c('feces' = 'feces', '7C' = '6C+Muc', '3C+Muc' = '3C+Muc', "7C-Muc"="6C", "3C"="3C", "H2O"="C-depleted", "Glc"="Glc")))+
    stat_ellipse( aes(group=stress_level, fill=stress_level),geom = "polygon", lwd=0.01, alpha=0.2, level=0.9 )+
    geom_point(aes(fill=stress_level, shape=time, colour= donor_name), alpha=1, size=5, stroke=2)+
    ggtitle(title)+
    # scale_fill_manual(values=c("#F8766D", "darkgoldenrod1", "#00BFC4"), name="Fibers")+
    theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_text(size=20),
        axis.text.x = element_text(size=20),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20))+
   xlab("PCoA1") + ylab("PCoA2")+labs(shape="Fibers", colour="Donor")+
    geom_text(aes(label=donor_name))->p
  
 show(p)
```
```{r}
#########
#BETA
#########
dist = "bray"
# dist = "jaccard"
ord_meths = c("PCoA")


phyloseq_rare%>% 
   subset_samples(time %in% c("stress",
                                                 "post_stress_1","post_stress_2"))%>%
  subset_samples(stress_level %in% "max")%>%
   subset_samples(time %in% "stress")%>%
     subset_samples( stress %in% c("O2"))->betas

plist = llply(as.list(ord_meths), function(i, betas, dist){
        ordi = ordinate(betas, method=i, distance=dist)
        plot_ordination(physeq = betas, 
                        ordination = ordi,
                        color = "donor_name", 
                        title = "PCoA Bray-Curtis")},
        betas, 
        dist)


names(plist) <- ord_meths

pdataframe = ldply(plist, function(x){
    df = x$data[, 1:2]
    colnames(df) = c("Axis_1", "Axis_2")
    return(cbind(df, x$data))})


names(pdataframe)[1] = "method"

pdataframe$stress_level<-factor(pdataframe$stress_level, levels=c("control", "low", "median", "high", "max"))  

  pdataframe%>%
    ggplot(aes(Axis_1, Axis_2))+ 
     scale_shape_manual(values = c(21, 24, 23), labels=c(c('feces' = 'feces', '7C' = '6C+Muc', '3C+Muc' = '3C+Muc', "7C-Muc"="6C", "3C"="3C", "H2O"="C-depleted", "Glc"="Glc")))+
    stat_ellipse( aes(group=incubation, fill=incubation),geom = "polygon", lwd=0.01, alpha=0.2, level=0.9 )+
    geom_point(aes(fill=incubation, shape=time, colour= donor_name), alpha=1, size=5, stroke=2)+
    ggtitle(title)+
      geom_line(aes(colour=donor_name))+
    # scale_fill_manual(values=c("#F8766D", "darkgoldenrod1", "#00BFC4"), name="Fibers")+
    theme(strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_text(size=20),
        axis.text.x = element_text(size=20),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20))+
   xlab("PCoA1") + ylab("PCoA2")+labs(shape="Fibers", colour="Donor")+
    geom_text(aes(label=donor_name))->p
  
 show(p)
```
```{r}




phyloseq_rare%>% 
  subset_samples(stress %in% c("O2") & time %in% c("stress") & stress_level %in%("max"))%>%
  ps_arrange(stress_level)%>%
  microViz::comp_barplot(
    tax_level = "Genus",
    label = "incubation",
    n_taxa = 20,
    bar_width = 0.9,
    sample_order = "default",
    tax_transform_for_plot = "compositional") +
    ylab("Proportion") + xlab( "")-> p_hist


p_hist + 
  facet_grid(cols=vars(donor_name), scales = "free", space = "free_x", drop = TRUE)+
  theme(axis.text.x = element_text(size=20, angle=90, vjust=0.2, hjust=1, margin=margin(-8,0,0,0)),
        axis.ticks.x = element_blank(),strip.text = element_text(size  = 20),
          axis.text.y= element_text(size = 20), 
        axis.title.y = element_text(size=20), 
        axis.title.x = element_text(size=20),
        legend.text = element_text(size = 20, vjust=1, hjust=0),
        legend.title = element_text(size= 20),
        plot.title = element_text(size=20), panel.spacing = unit(0.2, "cm")) + scale_y_continuous(breaks=c(0.25, 0.5, 0.75))+ ggtitle("Taxonomic distribution and ASV representation in cultures and feces")-> p_hist

p_hist


x<-paste0(donors_loop[i],"O2_stress_box.jpeg")
ggsave(filename = x,
       path = "C:/Users/zuendj/Desktop/03_data R/wp2_complex/output/max_stress_o2",
       width = 50,
       height = 30,
       units = c("cm"))


```

